### 2014-09-02 inst/doc/manuscript.Rmd
incorporated both rmflight and cboettig styles

---
title: "Modeling Lake Trophic State: A Data Mining Approach"
bibliography: LakeTrophicModelling.bib
layout: preprint
csl: ecology.csl
documentclass: elsarticle
output: 
  pdf_document:
    template: elsarticle.latex
    keep_tex: true
    fig_caption: true
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: First Coauthor 
    affiliation: shortcode
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: shortcode
    address: |
      Full address of First Coauthor
abstract: | 
          Write the abstract of the paper here
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Lake Trophic Modelling}
%\usepackage[vmargin=0.5in,hmargin=0.5in]{geometry}
%\usepackage{times}
-->

##Introduction
Cyanobacteria are an important taxonomic group associated with harmful algal blooms in lakes. Understanding the drivers of cyanobacteria presence has important implications for lake management and for
 the protection of human and ecosystem health.  Chlorophyll a concentration, a measure of the biological productivity of a lake, is one such driver and is largely, although not exclusively, determined
 by nutrient inputs.   As nutrient inputs increase, productivity increases and lakes transition from low trophic state (e.g. oligotrophic) to higher trophic states (e.g. hypereutrophic).  These broad
trophic state classifications are associated with ecosystem health and ecosystem services/disservices (e.g. recreation, aesthetics, fisheries, and harmful algal blooms).  Thus, models of trophic state
 might be used to predict things like cyanobacteria.

We have three goals for this preliminary research:

1. Build and assess models of lake trophic state
2. Assess ability to predict trophic state in lakes without available *in situ* water quality data
3. Explore association between cyanobacteria and trophic in order to expand models.

##Data and Modeling Methods

**Data**
We utilize four primary sources of data for this study. These are outlined below and in Table 1.

1. National Lakes Assessment (NLA) 2007: The NLA data were collected during the summer of 2007 and the final data were released in 2009.  With consistent methods and metrics collected at 1056 location
s across the conterminous United States (Map 1), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.  The NLA collected data on biophysical measures of lake wat
er quality and habitat.  For this analysis we primarily examined the water quality measurements from the NLA [@usepa2009national].
2. National Land Cover Dataset (NLCD) 2006: The NLCD is a nationally collected land use land cover dataset.  We collected total land use land cover and total percent impervious surface within a 3 kilo
meter buffer surrounding the lake to examine larger landscape-level effects [@homer2004development;@xian2009updating].
3. Modeled lake morphometry: Various measures of lake morphometry (i.e. depth, volume, fetch, etc.) are important in understanding lake productivity, yet many of these data are difficult to obtain for
 large numbers of lakes over broad regions.  To add this information we modeled lake morphometry [@hollister2010volume;@hollister_predicting_2011;@Lakmorpho2014;@NLMDInPrep].
4. Estimated Cyanobacteria Biovolumes: Cyanobacteria biovolumes is a truer measure of Cyanobacteria dominance than abundance as there is great variability in the size within and between species.  To a
ccount for this, Beaulieu *et al.* [-@beaulieu2013nutrients] used literature values to estimate biovolumes for the taxa in the NLA.  They shared this data and we have summed that information on a per-
lake basis.

##References

### 2014-09-02 R/LakeTrophicModelling.R

@@ -1,0 +1,5 @@
+#' LakeTrophicModelling
+#'
+#' @name LakeTrophicModelling
+#' @docType package
+NULL

### 2014-09-02 R/runRandomForest.R

@@ -1,0 +1,37 @@
+#' runRandomForest
+#' 
+#' This function takes the output of iterVarSelRF and runs a random forest. 
+#' 
+#' @param indVar independent variables
+#' @param depVar dependent variable
+#' @param ntree to be passed to randomForest.  Had trouble getting ... to work with the snowfall wrapper 
+#'              function.  This was a quick hack.
+#' @param importance  to be passed to randomForest.  Had trouble getting ... to work with the snowfall wrapper 
+#'                    function.  This was a quick hack.
+#' @param proximity to be passed to randomForest.  Had trouble getting ... to work with the snowfall wrapper 
+#'                  function.  This was a quick hack.
+#' 
+#' @examples
+#' data(hkm2014Data)
+#' x2<-iterVarSelRF(hkm2014Data[predictors_gis],TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
+#' vars.drop.frac=NULL,vars.drop.num=1,time=FALSE)
+#' x2_vars<-unique(unlist(x2))
+#' gis_ts4_rf<-runRandomForest(hkm2014Data[x2_vars],TS_CHLA_4, ntree=10,importance=TRUE,proximity=TRUE)
+#' @export
+runRandomForest <- function(indVar, depVar, ntree = 100, 
+    importance = TRUE, proximity = TRUE) {
+    
+    # Only Pass Complete Cases
+    comp_case <- complete.cases(indVar, depVar)
+    if (length(indVar) > 1) {
+        indVar <- indVar[comp_case, ]
+    } else {
+        indVar <- indVar[comp_case]
+    }
+    depVar <- depVar[comp_case]
+    
+    
+    return(randomForest(indVar, depVar, ntree = ntree, 
+        importance = importance, proximity = proximity))
+    
+} 

### 2014-09-02 vignettes/install.R

@@ -1,0 +1,11 @@
+#!/usr/bin/Rscript
+library("methods")
+# Install packrat if not available
+options(repos=c(CRAN="http://cran.rstudio.com"))
+if(!require("devtools")) install.packages("devtools")
+if(!require("packrat")) devtools::install_github("rstudio/packrat")
+packrat::packify()
+packrat::restore()
+
+
+

### 2014-09-02 vignettes/restore-cache.R

@@ -1,0 +1,20 @@
+#!/usr/bin/R
+library("methods")
+library("utils")
+library("httr")
+
+setwd("components")
+archive <- "template-manuscript-cache.tar.gz"
+
+# restore cache file
+target <- paste0("http://two.ucdavis.edu/~cboettig/data/", archive)
+cache <- GET(target)
+if(cache$status_code != 404){
+  writeBin(content(cache, "raw"), archive)
+  untar(archive)
+  unlink(archive)
+}
+
+setwd("..")
+
+

### 2014-09-04 inst/doc/manuscript.Rmd
cleaned up uneeded files, update on RMD

@@
 ---
-title: "Modeling Lake Trophic State: A Data Mining Approach"
-bibliography: LakeTrophicModelling.bib
-layout: preprint
-csl: ecology.csl
-documentclass: elsarticle
+title: 'Modeling Lake Trophic State: A Data Mining Approach'
+author:
+- affiliation: aed
+  email: hollister.jeff(at)epa.gov
+  footnote: Corresponding author
+  name: Jeffrey W. Hollister
+- affiliation: aed
+  name: W. Bryan Milstead
+- affiliation: aed
+  name: Betty J. Kreakie
+address:
+- address: "US Environmental Protection Agency Office of Research and Development National
+    Health and Environmental Effects Research Laboratory Atlantic Ecology Division 27
+    Tarzwell Drive  Narragansett, RI, 02879, USA  "
+  code: aed
 output:
   pdf_document:
+    fig_caption: yes
+    keep_tex: yes
     template: elsarticle.latex
-    keep_tex: true
-    fig_caption: true
-author: 
-  - name: Carl Boettiger
-    affiliation: cstar
-    email: cboettig(at)gmail.com
-    footnote: Corresponding author
-  - name: First Coauthor 
-    affiliation: shortcode
-address: 
-  - code: cstar
-    address: | 
-      Center for Stock Assessment Research,
-      110 Shaffer Rd,
-      Santa Cruz, CA 95050, USA 
-  - code: shortcode
-    address: |
-      Full address of First Coauthor
-abstract: | 
-          Write the abstract of the paper here
+  html_document: null
+  word_document: null
+csl: plos.csl
+documentclass: elsarticle
+layout: review
+linenumbers: no
+bibliography: LakeTrophicModelling.bib
+abstract:  Productivity of lentic ecosystems has been well studied and it is 
+  widely accepted that as nutrient inputs increase,  productivity increases and
+  lakes transition from low trophic state (e.g. oligotrophic) to higher trophic 
+  states  (e.g. eutrophic). These broad trophic state classifications are good 
+  predictors of ecosystem health and ecosystem services/disservices (e.g. recreation,
+  aesthetics, fisheries, and harmful algal blooms). While the relationship between 
+  nutrients and trophic state provides reliable predictions, it requires *in situ* 
+  water quality data in order to paramterize the model.  This limits the application 
+  of these models to lakes with existing and, more importanly, available water quality
+  data.  To expand our ability to predict in lakes without water quality data, we take
+  advantage of the availability of a large national lakes water quality database, 
+  land use/land cover data, lake morphometry data, other universally available data, 
+  and modern data mining approaches to build and assess models of lake tropic state 
+  that may be more universally applied.  We use random forests and random forest 
+  variable selection to identify variables to be used for predicting trophic state and 
+  we compare the performance of two models of trophic state (as determined by 
+  chlorophyll a concentration).   The first model estimates trophic state with *in situ* 
+  as well as universally available data and the second model uses universally available 
+  data only. For each of these models we used three separate trophic state categories, 
+  for a total of six models. Overall accuracy for the *in situ* and universal data 
+  models ranged from xx% to xx% and xx, xx, and xx described the most variation in 
+  trophic state.  For the universal data only models, Overall accuraccy ranged from xx% 
+  to xx% and xx, xx, and xx described the most variation in trophic state. Lastly, it 
+  is believed that the presence and abundance of cyanobacteria is strongly associated 
+  with trophic state.  To test this we examine the association between estimates of 
+  cyanobacteria biovolume and the measured and predicted trophic state.  Expanding 
+  these preliminary results to include cyanobacteria taxa indicates that cyanobacteria 
+  are significantly more likely to be found in highly eutrophic lakes. These results 
+  suggest that predictive models of lake trophic state may be improved with additional 
+  information on the landscape surrounding lakes and that those models provide 
+  additional information on the presence of potentially harmful cyanobacteria taxa. 
 ---
@@
 %\VignetteEngine{knitr::rmarkdown}
-%\VignetteIndexEntry{Lake Trophic Modelling}
-%\usepackage[vmargin=0.5in,hmargin=0.5in]{geometry}
-%\usepackage{times}
+%\VignetteIndexEntry{Modeling Lake Trophic State: A Data Mining Approach}
 -->
 
-##Introduction
+#Introduction
 Cyanobacteria are an important taxonomic group associated with harmful algal blooms in lakes. Understanding the drivers of cyanobacteria presence has important implications for lake management and for
@@
 
-##Data and Modeling Methods
+#Data and Modeling Methods
 
-**Data**
+##Data
+
 We utilize four primary sources of data for this study. These are outlined below and in Table 1.
@@
  large numbers of lakes over broad regions.  To add this information we modeled lake morphometry [@hollister2010volume;@hollister_predicting_2011;@Lakmorpho2014;@NLMDInPrep].
-4. Estimated Cyanobacteria Biovolumes: Cyanobacteria biovolumes is a truer measure of Cyanobacteria dominance than abundance as there is great variability in the size within and between species.  To a
+4.
+Estimated Cyanobacteria Biovolumes: Cyanobacteria biovolumes are a truer measure of Cyanobacteria dominance than abundance as there is great variability in the size within and between species.
+To a
 ccount for this, Beaulieu *et al.* [-@beaulieu2013nutrients] used literature values to estimate biovolumes for the taxa in the NLA.  They shared this data and we have summed that information on a per-
@@
 
-##References
+#References

### 2014-09-09 inst/doc/manuscript.Rmd
Got vignette running

@@
 
+```{r setup, include=FALSE, echo=FALSE}
+devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+devtools::install_github("wesanderson","jhollist")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("xtable")
+library("e1071")
+#opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+data(hkm2014Data_all)
+#Checks for existing cache (from another project)
+if(file.exists("cache")){
+  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+```
+
 #Introduction
@@
 
+```{r Table1, results='asis', echo=FALSE}
+translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
+               "Percent Low Intensity Development","Percent Medium Intensity Development",
+               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
+               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
+               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
+               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
+               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
+               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
+               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
+               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
+               "Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions",
+               "Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum",
+               "Ion Balance","Estimate Organic Anions","Calculated Conductivity",
+               "D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days",
+               "Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", 
+               "Fetch form Northeast","Fetch from East", "Fetch from Southeast",
+               "Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth",
+               "Nitrogen:Phophorus Ratio")
+wq<-!predictors_all%in%predictors_gis
+type<-vector("character",length(predictors_all))
+type[wq]<-"Water Quality"
+type[!wq]<-"GIS"
+data_def<-data.frame(Variables=predictors_all,Description=translation,Type=type)
+# This is an unbelievably silly table... It is life sized!
+kable(data_def)
+#dfToImage(data_def,style="css/texttable.css",file="figure/data_def_table.jpg")
+```
+
+```{r results='asis', echo=FALSE}
+state<-map_data('state')
+lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
+p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
+lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
+lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
+lakes_dd<-data.frame(coordinates(lakes_dd))
+mycolor<-wes.palette(5,"Zissou2")
+mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
+names(lakes_dd)<-c("long","lat",c())
+nlaMap(state,lakes_dd,mycolor)
+#ggsave("figure/nla_map.jpg",nlaMap(state,lakes_dd,mycolor),
+#       width=18,units="in",dpi=600)
+```
+**Predicting Trophic State with Random Forests**
+
+Random forest is a machine learning algorithm that aggregates numerous decision trees in order to obtain a consensus prediction of the response categories [@breiman2001random].
+Bootstrapped sample data is recursively partitioned according to a given random subset of predictor variables and completely grown without pruning.
+With each new tree, both the sample data and predictor variable subset is randomly selected.
+
+While random forests are able to handle numerous correlated variables without a decrease in prediction accuracy, unusually large numbers of related variables can reduce accuracy and increase the chances of over-fitting the model.
+This is a problem often faced in gene selection and in that field, a variable selection method based on random forest has been succesfully applied [@diaz2006gene].
+We use varselRF in R to initially examine the importance of the water quality and GIS derived variables and select a subset, the reduced model, to then pass to random forest[@diaz-uriarte2010varSelRF].
+
+Using R’s randomForest package, we pass the reduced models selected with varSelRF and calculate confusion matrices, overall accuracy and kappa coeffecient [@liaw2002randomForest].
+From the reduced model random forests we collect a consensus prediction and calculate a confusion matrix and summary stats.
+
+**Model Details**
+
+Using a combination of the `varSelRF` and `randomForest` we ran models for six combinations of variables and trophic state classifications.
+These combinations included different combinations of the Chlorphyll *a* trophic states (Table 2) along with all variables and the GIS only variables (i.e.
+no *in situ* infromation).
+The six model combinations were:
+
+1. Chlorophyll *a* trophic state - 4 class = All variables (*in situ* water quality, lake morphometry, and landscape)
+2. Chlorophyll *a* trophic state - 3 class = All variables (*in situ* water quality, lake morphometry, and landscape)
+3. Chlorophyll *a* trophic state - 2 class = All variables (*in situ* water quality, lake morphometry, and landscape)
+4. Chlorophyll *a* trophic state - 4 class = All variables (lake morphometry, and landscape)
+5. Chlorophyll *a* trophic state - 3 class = All variables (lake morphometry, and landscape)
+6. Chlorophyll *a* trophic state - 2 class = All variables (lake morphometry, and landscape)
+
+```{r trophicStateTable, results='asis', echo=FALSE, fig.cap="Table 2. Chlorphyll a based trophic state cut-offs"}
+ts_4<-c("oligo","meso","eu","hyper")
+ts_3<-c("oligo","meso/eu","meso/eu","hyper")
+ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
+co<-c("<= 0.2",">2-7",">7-30",">30")
+xdf<-data.frame(ts_4,ts_3,ts_2,co)
+names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
+kable(xdf)
+#dfToImage(xdf,style="css/texttable.css",file="figure/ts_classes_table.jpg")
+```
+
+##Results
+
+```{r figSetup,echo=FALSE,eval=T,cache=TRUE}
+#create list of all selected variables and assign a color
+vars<-ls(pattern="_vars")
+all_vars<-vector()
+for(i in vars){
+  all_vars<-c(all_vars,get(i))
+}
+all_vars<-unique(all_vars)
+
+all_movies_palette<-vector()
+for(i in 1:dim(namelist)[1]){
+  all_movies_palette<-c(all_movies_palette,wes.palette(namelist[i,2],
+                                                       namelist[i,1]))
+}
+all_movies_palette<-unique(all_movies_palette)
+zissou<-c(wes.palette(5,"Zissou"),wes.palette(5,"Zissou2"))
+col_lu<-data.frame(variables=all_vars,hexcode=sample(all_movies_palette,length(all_vars)))
+```
+
+###Model 1: 4 Trophic States ~ All Variables
+
+```{r ALL_TS4, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts4<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_4,numCore=4,100,
+    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="all_ts4",time=TRUE)
+```
+
+```{r ALL_TS4_rf,echo=FALSE,eval=FALSE,cache=T}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts4_vars<-unique(unlist(all_ts4[1:100]))
+all_ts4_rf<-runRandomForest(hkm2014Data[all_ts4_vars],hkm2014Data$TS_CHLA_4, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model1,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+#dfToImage(sumTable(all_ts4[1:100]),file="figure/all_ts4_table.jpg",
+#         style="css/sumtable.css",width=8,include.rownames=FALSE)
+#dfToImage(formatC(round(all_ts4_rf$confusion,2)),file="figure/all_ts4_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(all_ts4[1:100]),row.names=F)
+kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
+
+all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
+importancePlot(all_ts4_rf,'acc',size=10,aes(colour=all_ts4_color))
+importancePlot(all_ts4_rf,'gini',size=10,aes(colour=all_ts4_color))
+```
+
+Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)`.
+
+###Model 2: 3 Trophic States ~ All Variables
+
+```{r ALL_TS3, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts3<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_3,numCore=4,100,
+  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="all_ts3",time=TRUE)
+```
+
+```{r all_TS3_rf,echo=FALSE,eval=FALSE,cache=T}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts3_vars<-unique(unlist(all_ts3[1:100]))
+all_ts3_rf<-runRandomForest(hkm2014Data[all_ts3_vars],hkm2014Data$TS_CHLA_3, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model2,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+#dfToImage(sumTable(all_ts3[1:100]),file="figure/all_ts3_table.jpg",
+#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+#dfToImage(formatC(round(all_ts3_rf$confusion,2)),file="figure/all_ts3_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(all_ts3[1:100]),row.names=F)
+kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
+
+all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
+importancePlot(all_ts3_rf,'acc',size=10,aes(colour=all_ts3_color))
+importancePlot(all_ts3_rf,'gini',size=10,aes(colour=all_ts3_color))
+```
+
+Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+###Model 3: 2 Trophic States ~ All Variables
+
+```{r ALL_TS2, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts2<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_2,numCore=4,100,
+  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="all_ts2",time=TRUE)
+```
+
+```{r all_TS2_rf,echo=FALSE,eval=FALSE,cache=T}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+all_ts2_vars<-unique(unlist(all_ts2[1:100]))
+all_ts2_rf<-runRandomForest(hkm2014Data[all_ts2_vars],hkm2014Data$TS_CHLA_2, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model3,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+#dfToImage(sumTable(all_ts2[1:100]),file="figure/all_ts2_table.jpg",
+#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+#dfToImage(formatC(round(all_ts2_rf$confusion,2)),file="figure/all_ts2_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(all_ts2[1:100]),row.names=F)
+kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
+
+all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
+importancePlot(all_ts2_rf,'acc',size=10,aes(colour=all_ts2_color))
+importancePlot(all_ts2_rf,'gini',size=10,aes(colour=all_ts2_color))
+```
+
+Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+###Model 4: 4 Trophic States ~ GIS Only Variables
+
+```{r GIS_TS4, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts4<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_4,numCore=4,100,
+      								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="gis_ts4",time=TRUE)
+```
+
+```{r GIS_TS4_rf,echo=FALSE,eval=FALSE,cache=T}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts4_vars<-unique(unlist(gis_ts4[1:100]))
+gis_ts4_rf<-runRandomForest(hkm2014Data[gis_ts4_vars],hkm2014Data$TS_CHLA_4, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model4,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8,fig.cap="Figure 1"}
+#dfToImage(sumTable(gis_ts4[1:100]),file="figure/gis_ts4_table.jpg",
+#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+
+#dfToImage(formatC(round(gis_ts4_rf$confusion,2)),file="figure/gis_ts4_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(gis_ts4[1:100]),row.names=F)
+kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
+
+gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
+importancePlot(gis_ts4_rf,'acc',size=10,aes(colour=gis_ts4_color))
+importancePlot(gis_ts4_rf,'gini',size=10,aes(colour=gis_ts4_color))
+```
+
+Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
+
+###Model 5: 3 Trophic States ~ GIS Only Variables
+
+```{r GIS_TS3, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts3<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_3,numCore=4,100,
+    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="gis_ts3",time=TRUE)
+```
+
+```{r GIS_TS3_rf,echo=FALSE,eval=FALSE,cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts3_vars<-unique(unlist(gis_ts3[1:100]))
+gis_ts3_rf<-runRandomForest(hkm2014Data[gis_ts3_vars],hkm2014Data$TS_CHLA_3, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model5,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+#dfToImage(sumTable(gis_ts3[1:100]),file="figure/gis_ts3_table.jpg",
+#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+#dfToImage(formatC(round(gis_ts3_rf$confusion,2)),file="figure/gis_ts3_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(gis_ts3[1:100]),row.names=F)
+kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
+
+gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
+importancePlot(gis_ts3_rf,'acc',size=10,aes(colour=gis_ts3_color))
+importancePlot(gis_ts3_rf,'gini',size=10,aes(colour=gis_ts3_color))
+```
+
+Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+###Model 6: 2 Trophic States ~ GIS Only Variables
+
+```{r GIS_TS2, echo=FALSE, eval=FALSE, cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts2<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_2,numCore=4,100,
+  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+											vars.drop.num=1,outStr="gis_ts2",time=TRUE)
+```
+
+```{r GIS_TS2_rf,echo=FALSE,eval=FALSE,cache=TRUE}
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+gis_ts2_vars<-unique(unlist(gis_ts2[1:100]))
+gis_ts2_rf<-runRandomForest(hkm2014Data[gis_ts2_vars],hkm2014Data$TS_CHLA_2, ntree=10000,importance=TRUE,proximity=TRUE)
+```
+
+```{r Figure_Model6,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+#dfToImage(sumTable(gis_ts2[1:100]),file="figure/gis_ts2_table.jpg",
+#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+#dfToImage(formatC(round(gis_ts2_rf$confusion,2)),file="figure/gis_ts2_confusion.jpg",
+#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+kable(sumTable(gis_ts2[1:100]),row.names=F)
+kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
+
+gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
+importancePlot(gis_ts2_rf,'acc',size=10,aes(colour=gis_ts2_color))
+importancePlot(gis_ts2_rf,'gini',size=10,aes(colour=gis_ts2_color))
+```
+
+Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+###Associating Trophic State and Cyanobacteria
+
+```{r ts_4_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,
+                                     cdf_colors=zissou[6:9],y='Percent',
+                                     x='Log10(Cyanobacteria Biovolume)',
+                                     title=expression(paste('CDF for Chlorophyll ', 
+                                     italic("a"),' Trophic States (4 Categories)')),
+                                     color="Trophic State\nCategories")
+```
+
+```{r ts_3_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
+       y='Percent',x='Log10(Cyanobacteria Biovolume)',
+       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
+       color="Trophic State\nCategories")
+```
+
+```{r ts_2_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
+       y='Percent',x='Log10(Cyanobacteria Biovolume)',
+       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (2 Categories)')),
+       color="Trophic State\nCategories")
+```
+
+```{r scatterplot,echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
+scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
+             title=expression(paste("Chlorophyll ", italic("a")," and Cyanobacteria Relationship")),
+             x=expression(paste('Log10(Chl ', italic("a"),')')),
+             y="Log10(Cyanobaterial Biovolumes + 1)")
+```
+
 #References

### 2014-09-09 inst/doc/manuscript.R

@@ -1,0 +1,175 @@
+## ----setup, include=FALSE, echo=FALSE------------------------------------
+devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+devtools::install_github("wesanderson","jhollist")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("xtable")
+library("e1071")
+#opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+data(hkm2014Data_all)
+#Checks for existing cache (from another project)
+if(file.exists("cache")){
+  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+
+## ----Table1, results='asis', echo=FALSE----------------------------------
+translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
+               "Percent Low Intensity Development","Percent Medium Intensity Development",
+               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
+               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
+               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
+               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
+               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
+               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
+               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
+               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
[…]

### 2014-09-09 vignettes/manuscript.R

@@ -1,49 +1,0 @@
-## ----setup, include=FALSE, echo=FALSE------------------------------------
-
-
-devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
-                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-devtools::install_github("wesanderson","jhollist")
-library("wesanderson")
-library("LakeTrophicModelling")
-library("knitr")
-library("ggplot2")
-library("sp")
-library("rgdal")
-opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
-data(hkm2014Data_all)
-#Checks for existing cache (from another project)
-if(file.exists("cache")){
-  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
-    lazyLoad(i)
-  }
-}
-
-## ----Table1, results='asis', echo=FALSE----------------------------------
-translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
-               "Percent Low Intensity Development","Percent Medium Intensity Development",
-               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
-               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
-               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
-               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
-               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
-               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
-               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
-               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
-               "Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions",
-               "Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum",
-               "Ion Balance","Estimate Organic Anions","Calculated Conductivity",
-               "D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days",
[…]

### 2014-09-10 inst/doc/manuscript.Rmd
updated manuscript.Rmd and rendered vignette

@@
 library("e1071")
-#opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
 data(hkm2014Data_all)
@@
 #Introduction
-Cyanobacteria are an important taxonomic group associated with harmful algal blooms in lakes. Understanding the drivers of cyanobacteria presence has important implications for lake management and for
- the protection of human and ecosystem health.  Chlorophyll a concentration, a measure of the biological productivity of a lake, is one such driver and is largely, although not exclusively, determined
- by nutrient inputs.   As nutrient inputs increase, productivity increases and lakes transition from low trophic state (e.g. oligotrophic) to higher trophic states (e.g. hypereutrophic).  These broad
-trophic state classifications are associated with ecosystem health and ecosystem services/disservices (e.g. recreation, aesthetics, fisheries, and harmful algal blooms).  Thus, models of trophic state
- might be used to predict things like cyanobacteria.
 
-We have three goals for this preliminary research:
+Productivity in lentic systems is often categorized across a range of tropic states (e.g.
+the tropic continuum) from early succesional (i.e.
+oligotrophic)to late successional lakes (i.e.
+hypereutrophic) [@carlson1977trophic].
+Lakes naturally occur across the range of trophic state and higher primary productivity is not necessarily a predictor of poor ecological condition.
+Lakes that are naturally oligotrophic occur in nutrient poor areas or have a more recent geologic history.
+These lakes are often found in higher elevations, have clear water, and are often favored for drinking water or direct contact recreation (e.g.
+swimming).
+Lakes with higher productivity (e.g.
+eutrophic lakes) have greater nutrient loads, tend to be less clear, have greater density of aquatic plants, and often support more diverse and abundant fish communities.
+Lakes will naturally shift to higher trophic states but this is a slow process.
+Given this fact, monitoring trophic state allows the identification of rapid shifts in trophic state or locating lakes with unusually high productivity (e.g.
+hypereutrophic).
+These cases are indicative of lakes under greater anthropogenic nutrient loads, also known as cultural eutrophication, and are more likely to be at risk of fish kills, fouling, and harmful algal blooms[@smith1998cultural;@smith1999eutrophication;@smith2006eutrophication].
+Given the association between trophic state and many ecosystem services and disservices, being able to model trophic state could allow for estimating trophic state in unmonitored lakes and provide a first cut at identifying lakes with the potential for harmful algal blooms and other problems associated with cultural eutrophication.
 
-1. Build and assess models of lake trophic state
-2. Assess ability to predict trophic state in lakes without available *in situ* water quality data
-3. Explore association between cyanobacteria and trophic in order to expand models.
+Cyanobacteria are an important taxonomic group associated with harmful algal blooms in lakes.
+Understanding the drivers of cyanobacteria presence has important implications for lake management and for the protection of human and ecosystem health.
+Chlorophyll a concentration, a measure of the biological productivity of a lake, is one such driver and is largely, although not exclusively, determined by nutrient inputs.
+As nutrient inputs increase, productivity increases and lakes transition from low trophic state (e.g.
+oligotrophic) to higher trophic states (e.g.
+hypereutrophic).
+These broad trophic state classifications are associated with ecosystem health and ecosystem services/disservices (e.g.
+recreation, aesthetics, fisheries, and harmful algal blooms).
+Thus, models of trophic state might be used to predict things like cyanobacteria.
 
-#Data and Modeling Methods
+We have three goals for this preliminary research.
+First, we build and assess multiple models of lake trophic state using a full suite of data including *in situ* water quality and universally available data (e.g.
+landscape data).
+Second, we assess the accuracy of predicted trophic state in lakes with only the universally available data.
+Lastly, we explore associations between trophic state and cyanobacteria to explore.
 
-##Data
+#Methods
 
-We utilize four primary sources of data for this study. These are outlined below and in Table 1.
+##Data and Study Area
+We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].
+All datasets are national in scale and provide a unique snapshot view of the condition of lakes in the United States'.
 
-1. National Lakes Assessment (NLA) 2007: The NLA data were collected during the summer of 2007 and the final data were released in 2009.  With consistent methods and metrics collected at 1056 location
-s across the conterminous United States (Map 1), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.  The NLA collected data on biophysical measures of lake wat
-er quality and habitat.  For this analysis we primarily examined the water quality measurements from the NLA [@usepa2009national].
-2. National Land Cover Dataset (NLCD) 2006: The NLCD is a nationally collected land use land cover dataset.  We collected total land use land cover and total percent impervious surface within a 3 kilo
-meter buffer surrounding the lake to examine larger landscape-level effects [@homer2004development;@xian2009updating].
-3. Modeled lake morphometry: Various measures of lake morphometry (i.e. depth, volume, fetch, etc.) are important in understanding lake productivity, yet many of these data are difficult to obtain for
- large numbers of lakes over broad regions.  To add this information we modeled lake morphometry [@hollister2010volume;@hollister_predicting_2011;@Lakmorpho2014;@NLMDInPrep].
-4.
-Estimated Cyanobacteria Biovolumes: Cyanobacteria biovolumes are a truer measure of Cyanobacteria dominance than abundance as there is great variability in the size within and between species.
-To a
-ccount for this, Beaulieu *et al.* [-@beaulieu2013nutrients] used literature values to estimate biovolumes for the taxa in the NLA.  They shared this data and we have summed that information on a per-
-lake basis.
+The NLA data were collected during the summer of 2007 and the final data were released in 2009.
+With consistent methods and metrics collected at 1056 locations across the conterminous United States (Map 1), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.
+The NLA collected data on biophysical measures of lake water quality and habitat.
+For this analysis we primarily examined the water quality measurements from the NLA [@usepa2009national].
+Adding to the monitoring data collected via the NLA, we use the 2006 NLCD data to examine the possible landscape-level drivers of trophic status in lakes.
+The NLCD is a nationally collected land use land cover dataset that also provides estimates of impervious surface.
+We collected total land use land cover and total percent impervious surface within a 3 kilometer buffer surrounding the lake to examine larger landscape-level effect [@homer2004development;@xian2009updating].
+We also used various measures of lake morphometry (i.e.
+depth, volume, fetch, etc.) as they are important in understanding lake productivity, yet many of these data are difficult to obtain for large numbers of lakes over broad regions.
+To add this information we modeled lake morphometry [@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@NLMDInPrep].
+Lastly, to explore associations between trophic state and cyanobacteria, we used estimates of cyanobacterial biovoulme caluclated by Beaulieu *et al.* [-@beaulieu2013nutrients].
+Cyanobacteria biovolumes are a truer measure of cyanobacteria dominance than abundance as there is great variability in the size within and between species.
+We have consolidated the taxa level estimates from Beaulieu *et al.* [-@beaulieu2013nutrients] and  summed that information on a per-lake basis.
 

### 2014-09-10 inst/doc/manuscript.R

@@ -1,0 +1,271 @@
+## ----setup, include=FALSE, echo=FALSE------------------------------------
+devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+devtools::install_github("wesanderson","jhollist")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("xtable")
+library("e1071")
+opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+data(hkm2014Data_all)
+#Checks for existing cache (from another project)
+if(file.exists("cache")){
+  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+
+## ----Table1, results='asis', echo=FALSE----------------------------------
+translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
+               "Percent Low Intensity Development","Percent Medium Intensity Development",
+               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
+               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
+               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
+               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
+               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
+               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
+               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
+               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
[…]

### 2014-09-11 inst/doc/manuscript.Rmd
updated vignette build from new Rmd

@@
   word_document: null
+fontsize: 11pt
 csl: plos.csl
@@
 layout: review
-linenumbers: no
+linenumbers: yes
 bibliography: LakeTrophicModelling.bib
@@
 %\VignetteIndexEntry{Modeling Lake Trophic State: A Data Mining Approach}
+
 -->
@@
 ```{r setup, include=FALSE, echo=FALSE}
-devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
-                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-devtools::install_github("wesanderson","jhollist")
+#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+#                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+#devtools::install_github("wesanderson","jhollist")
 library("wesanderson")
@@
 data(hkm2014Data_all)
+
 #Checks for existing cache (from another project)
@@
 }
+
+# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+
+knit_hooks$set(tab.cap = function(before, options, envir) {
+  if(!before)
+    paste('\n\n:', options$tab.cap, sep='') 
+})
+default_output_hook = knit_hooks$get("output")
+knit_hooks$set(output = function(x, options) {
+  if (is.null(options$tab.cap) == F)  
+    x
+  else
+    default_output_hook(x,options)
+})
 ```
@@
 ##Data and Study Area
+
 We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].
@@
 
-```{r Table1, results='asis', echo=FALSE}
-translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
-               "Percent Low Intensity Development","Percent Medium Intensity Development",
-               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
-               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
-               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
-               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
-               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
-               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
-               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
-               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
-               "Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions",
-               "Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum",
-               "Ion Balance","Estimate Organic Anions","Calculated Conductivity",
-               "D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days",
-               "Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", 
-               "Fetch form Northeast","Fetch from East", "Fetch from Southeast",
-               "Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth",
-               "Nitrogen:Phophorus Ratio")
+```{r Variables, results='asis', eval=FALSE, echo=FALSE}
+#tab.cap="Definition of all variables considered"
+translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space","Percent Low Intensity Development","Percent Medium Intensity Development", "Percent High Intensity Development","Percent Barren","Percent Decidous Forest","Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland", "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter","Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area","Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity","Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon","Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate","Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions","Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum","Ion Balance","Estimate Organic Anions","Calculated Conductivity","D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days","Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", "Fetch form Northeast","Fetch from East", "Fetch from Southeast","Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth","Nitrogen:Phophorus Ratio")
 wq<-!predictors_all%in%predictors_gis
@@
 
-```{r results='asis', echo=FALSE}
+```{r results='asis', echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations"}
 state<-map_data('state')
@@
 nlaMap(state,lakes_dd,mycolor)
-#ggsave("figure/nla_map.jpg",nlaMap(state,lakes_dd,mycolor),
-#       width=18,units="in",dpi=600)
 ```
+
 **Predicting Trophic State with Random Forests**
@@
 
-```{r trophicStateTable, results='asis', echo=FALSE, fig.cap="Table 2. Chlorphyll a based trophic state cut-offs"}
+```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorphyll a based trophic state cut-offs"}
 ts_4<-c("oligo","meso","eu","hyper")
@@
 kable(xdf)
-#dfToImage(xdf,style="css/texttable.css",file="figure/ts_classes_table.jpg")
 ```
@@
 
-```{r Figure_Model1,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-#dfToImage(sumTable(all_ts4[1:100]),file="figure/all_ts4_table.jpg",
-#         style="css/sumtable.css",width=8,include.rownames=FALSE)
-#dfToImage(formatC(round(all_ts4_rf$confusion,2)),file="figure/all_ts4_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)`.
+
+```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1"}
 kable(sumTable(all_ts4[1:100]),row.names=F)
+```
+
+```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1"}
 kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1"}
 all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
-importancePlot(all_ts4_rf,'acc',size=10,aes(colour=all_ts4_color))
 importancePlot(all_ts4_rf,'gini',size=10,aes(colour=all_ts4_color))
 ```
-
-Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)`.
 
@@
 
-```{r Figure_Model2,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-#dfToImage(sumTable(all_ts3[1:100]),file="figure/all_ts3_table.jpg",
-#          style="css/sumtable.css",width=8,include.rownames=FALSE)
-#dfToImage(formatC(round(all_ts3_rf$confusion,2)),file="figure/all_ts3_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2"}
 kable(sumTable(all_ts3[1:100]),row.names=F)
+```
+
+```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2"}
 kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2"}
 all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
-importancePlot(all_ts3_rf,'acc',size=10,aes(colour=all_ts3_color))
 importancePlot(all_ts3_rf,'gini',size=10,aes(colour=all_ts3_color))
 ```
-
-Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
 
@@
 
-```{r Figure_Model3,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-#dfToImage(sumTable(all_ts2[1:100]),file="figure/all_ts2_table.jpg",
-#          style="css/sumtable.css",width=8,include.rownames=FALSE)
-#dfToImage(formatC(round(all_ts2_rf$confusion,2)),file="figure/all_ts2_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3"}
 kable(sumTable(all_ts2[1:100]),row.names=F)
+```
+
+```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3"}
 kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3"}
 all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
-importancePlot(all_ts2_rf,'acc',size=10,aes(colour=all_ts2_color))
 importancePlot(all_ts2_rf,'gini',size=10,aes(colour=all_ts2_color))
 ```
-
-Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
 
@@
 
-```{r Figure_Model4,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8,fig.cap="Figure 1"}
-#dfToImage(sumTable(gis_ts4[1:100]),file="figure/gis_ts4_table.jpg",
-#          style="css/sumtable.css",width=8,include.rownames=FALSE)
+Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
 
-#dfToImage(formatC(round(gis_ts4_rf$confusion,2)),file="figure/gis_ts4_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4" }
 kable(sumTable(gis_ts4[1:100]),row.names=F)
+```
+
+```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4" }
 kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 3"}
 gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
-importancePlot(gis_ts4_rf,'acc',size=10,aes(colour=gis_ts4_color))
 importancePlot(gis_ts4_rf,'gini',size=10,aes(colour=gis_ts4_color))
 ```
-
-Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
 
@@
 
-```{r Figure_Model5,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-#dfToImage(sumTable(gis_ts3[1:100]),file="figure/gis_ts3_table.jpg",
-#          style="css/sumtable.css",width=8,include.rownames=FALSE)
-#dfToImage(formatC(round(gis_ts3_rf$confusion,2)),file="figure/gis_ts3_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5"}
 kable(sumTable(gis_ts3[1:100]),row.names=F)
+```
+
+```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5"}
 kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5"}
 gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
-importancePlot(gis_ts3_rf,'acc',size=10,aes(colour=gis_ts3_color))
 importancePlot(gis_ts3_rf,'gini',size=10,aes(colour=gis_ts3_color))
 ```
-
-Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
 
@@
 
-```{r Figure_Model6,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-#dfToImage(sumTable(gis_ts2[1:100]),file="figure/gis_ts2_table.jpg",
-#          style="css/sumtable.css",width=8,include.rownames=FALSE)
-#dfToImage(formatC(round(gis_ts2_rf$confusion,2)),file="figure/gis_ts2_confusion.jpg",
-#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
+Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6"}
 kable(sumTable(gis_ts2[1:100]),row.names=F)
+```
+
+```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6"}
 kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
+```
 
+```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6"}
 gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
-importancePlot(gis_ts2_rf,'acc',size=10,aes(colour=gis_ts2_color))
 importancePlot(gis_ts2_rf,'gini',size=10,aes(colour=gis_ts2_color))
@@
 
-Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
-
 ###Associating Trophic State and Cyanobacteria
 
-```{r ts_4_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
-plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,
-                                     cdf_colors=zissou[6:9],y='Percent',
-                                     x='Log10(Cyanobacteria Biovolume)',
+```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes"}
+plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:9],
+        y='Percent',x='Log10(Cyanobacteria Biovolume)',
         title=expression(paste('CDF for Chlorophyll ', 
@@
 
-```{r ts_3_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes" }
 plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
        y='Percent',x='Log10(Cyanobacteria Biovolume)', 
-       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
+       title=expression(paste('CDF for Chlorophyll ', 
+                              italic("a"),' Trophic States (3 Categories)')),
        color="Trophic State\nCategories")
@@
 
-```{r ts_2_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes"}
 plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
        y='Percent',x='Log10(Cyanobacteria Biovolume)',
-       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (2 Categories)')),
+       title=expression(paste('CDF for Chlorophyll ', 
+                              italic("a"),' Trophic States (2 Categories)')),
        color="Trophic State\nCategories")
@@
 
-```{r scatterplot,echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
+```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot"}
 scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
 scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
-             title=expression(paste("Chlorophyll ", italic("a")," and Cyanobacteria Relationship")),
+             title=expression(paste("Chlorophyll ", 
+                                    italic("a")," and Cyanobacteria Relationship")),
              x=expression(paste('Log10(Chl ', italic("a"),')')),
@@
 
+\newpage
+
 #References

### 2014-09-11 vignettes/temp.R

@@ -1,0 +1,298 @@
+
+```{r nla_map,echo=FALSE,fig.cap=cap_num('Map of 2007 National Lake Assessment Sample Locations','figure')}
+state<-map_data('state')
+lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
+p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
+lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
+lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
+lakes_dd<-data.frame(coordinates(lakes_dd))
+mycolor<-wes.palette(5,"Zissou2")
+mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
+names(lakes_dd)<-c("long","lat",c())
+nlaMap(state,lakes_dd,mycolor)
+#ggsave("figure/nla_map.jpg",nlaMap(state,lakes_dd,mycolor),
+#       width=18,units="in",dpi=600)
+```
+**Predicting Trophic State with Random Forests**
+  
+  Random forest is a machine learning algorithm that aggregates numerous decision trees in order to obtain a consensus prediction of the response categories [@breiman2001random].  Bootstrapped sample data is recursively partitioned according to a given random subset of predictor variables and completely grown without pruning.   With each new tree, both the sample data and predictor variable subset is randomly selected.  
+
[…]

### 2014-09-12 inst/doc/manuscript.Rmd
cleaned up vignettes, got table and figure labels and captions working, rebuilt example in inst/doc

@@
     keep_tex: yes
-    template: elsarticle.latex
+    number_sections: yes
+    template: components/elsarticle.latex
   html_document: null
@@
 fontsize: 11pt
-csl: plos.csl
+csl: components/ecology.csl
 documentclass: elsarticle
 layout: review
-linenumbers: yes
-bibliography: LakeTrophicModelling.bib
+linenumbers: no
+bibliography: components/LakeTrophicModelling.bib
 abstract:  Productivity of lentic ecosystems has been well studied and it is 
@@
 %\VignetteIndexEntry{Modeling Lake Trophic State: A Data Mining Approach}
-
 -->
@@
 knit_hooks$set(tab.cap = function(before, options, envir) {
-  if(!before)
+                  if(!before) { 
                     paste('\n\n:', options$tab.cap, sep='') 
+                  }
                 })
@@
 knit_hooks$set(output = function(x, options) {
-  if (is.null(options$tab.cap) == F)  
+  if (is.null(options$tab.cap) == FALSE) {
     x
-  else
+  } else
     default_output_hook(x,options)
 })
+
 ```
@@
 The NLA data were collected during the summer of 2007 and the final data were released in 2009.
-With consistent methods and metrics collected at 1056 locations across the conterminous United States (Map 1), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.
+With consistent methods and metrics collected at 1056 locations across the conterminous United States (Figure \ref{fig:nlaMap}), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.
 The NLA collected data on biophysical measures of lake water quality and habitat.
@@
 
-```{r results='asis', echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations"}
+```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}"}
 state<-map_data('state')
@@
 
-**Predicting Trophic State with Random Forests**
+##Predicting Trophic State with Random Forests
 
@@
 
-**Model Details**
+##Model Details
 
 Using a combination of the `varSelRF` and `randomForest` we ran models for six combinations of variables and trophic state classifications.
-These combinations included different combinations of the Chlorphyll *a* trophic states (Table 2) along with all variables and the GIS only variables (i.e.
+These combinations included different combinations of the Chlorphyll *a* trophic states (Table \ref{tab:trophicStateTable}) along with all variables and the GIS only variables (i.e.
 no *in situ* infromation).
@@
 
-```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorphyll a based trophic state cut-offs"}
+```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}"}
 ts_4<-c("oligo","meso","eu","hyper")
@@
 
-##Results
+#Results
 
-```{r figSetup,echo=FALSE,eval=T,cache=TRUE}
+```{r figSetup,echo=FALSE,eval=TRUE,cache=TRUE}
 #create list of all selected variables and assign a color
@@
 
-###Model 1: 4 Trophic States ~ All Variables
+##Model 1: 4 Trophic States ~ All Variables
 
@@
 
-Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)`.
+The selected variables that made up Model 1 were Potassium, Nitrogen:Phosphorus, Total Nitrogen, Total Phosphorus, Total Organic Carbon, Turbidity, Ecoregion, Organic Ions, Dissolved Organic Carbon, and Maximum Lake Depth (Table \ref{tab:VarSel_Model1}).
+Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)` (Table \ref{tab:Confusion_Model1}).
 
-```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1"}
+```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}"}
 kable(sumTable(all_ts4[1:100]),row.names=F)
@@
 
-```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1"}
+```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}"}
 kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1"}
+Lastly, tubidity, total phosphorus, total nitrogen, and total organic carbon were the most important predictors of the 4 classes of trophic state (Figure \ref{fig:Importance_Model1}).
+
+```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1\\label{fig:Importance_Model1}"}
 all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
@@
 
-###Model 2: 3 Trophic States ~ All Variables
+##Model 2: 3 Trophic States ~ All Variables
 
@@
 
-###Model 3: 2 Trophic States ~ All Variables
+##Model 3: 2 Trophic States ~ All Variables
 
@@
 
-###Model 4: 4 Trophic States ~ GIS Only Variables
+##Model 4: 4 Trophic States ~ GIS Only Variables
 
@@
 
-```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 3"}
+```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 4"}
 gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
@@
 
-###Model 5: 3 Trophic States ~ GIS Only Variables
+##Model 5: 3 Trophic States ~ GIS Only Variables
 
@@
 
-###Model 6: 2 Trophic States ~ GIS Only Variables
+##Model 6: 2 Trophic States ~ GIS Only Variables
 
@@
 
-###Associating Trophic State and Cyanobacteria
+##Associating Trophic State and Cyanobacteria
 
@@
 ```
-
+\newpage
 ```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes" }
@@
 ```
-
+\newpage
 ```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes"}
@@
 ```
-
+\newpage
 ```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot"}

### 2014-09-12 inst/doc/manuscript.R

@@ -23,16 +23,19 @@ if(file.exists("cache")){
 # Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
 
 knit_hooks$set(tab.cap = function(before, options, envir) {
-  if(!before)
-    paste('\n\n:', options$tab.cap, sep='') 
-})
+                  if(!before) { 
+                    paste('\n\n:', options$tab.cap, sep='') 
+                  }
+                })
 default_output_hook = knit_hooks$get("output")
 knit_hooks$set(output = function(x, options) {
-  if (is.null(options$tab.cap) == F)  
+  if (is.null(options$tab.cap) == FALSE) {
     x
-  else
+  } else
     default_output_hook(x,options)
 })
+
+
 
 ## ----Variables, results='asis', eval=FALSE, echo=FALSE-------------------
 #  #tab.cap="Definition of all variables considered"
@@ -46,7 +49,7 @@ knit_hooks$set(output = function(x, options) {
 #  kable(data_def)
 #  #dfToImage(data_def,style="css/texttable.css",file="figure/data_def_table.jpg")
 
-## ----results='asis', echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations"----
+## ----nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}"----
 state<-map_data('state')
 lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
 p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
@@ -59,7 +62,7 @@ mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
 names(lakes_dd)<-c("long","lat",c())
 nlaMap(state,lakes_dd,mycolor)
 
-## ----trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorphyll a based trophic state cut-offs"----
[…]

### 2014-09-12 vignettes/temp.R

@@ -1,298 +1,0 @@
-
-```{r nla_map,echo=FALSE,fig.cap=cap_num('Map of 2007 National Lake Assessment Sample Locations','figure')}
-state<-map_data('state')
-lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
-p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
-ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
-lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
-lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
-lakes_dd<-data.frame(coordinates(lakes_dd))
-mycolor<-wes.palette(5,"Zissou2")
-mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
-names(lakes_dd)<-c("long","lat",c())
-nlaMap(state,lakes_dd,mycolor)
-#ggsave("figure/nla_map.jpg",nlaMap(state,lakes_dd,mycolor),
-#       width=18,units="in",dpi=600)
-```
-**Predicting Trophic State with Random Forests**
-  
-  Random forest is a machine learning algorithm that aggregates numerous decision trees in order to obtain a consensus prediction of the response categories [@breiman2001random].  Bootstrapped sample data is recursively partitioned according to a given random subset of predictor variables and completely grown without pruning.   With each new tree, both the sample data and predictor variable subset is randomly selected.  
-
[…]

### 2014-09-12 inst/doc/manuscript.Rmd
monkeying with latex

@@
 title: 'Modeling Lake Trophic State: A Data Mining Approach'
-author:
-- affiliation: aed
-  email: hollister.jeff(at)epa.gov
-  footnote: Corresponding author
-  name: Jeffrey W. Hollister
-- affiliation: aed
-  name: W. Bryan Milstead
-- affiliation: aed
-  name: Betty J. Kreakie
-address:
-- address: "US Environmental Protection Agency Office of Research and Development National
-    Health and Environmental Effects Research Laboratory Atlantic Ecology Division 27
-    Tarzwell Drive  Narragansett, RI, 02879, USA  "
-  code: aed
 output:
@@
 fontsize: 11pt
+capsize: normalsize
 csl: components/ecology.csl
 documentclass: elsarticle
-layout: review
+spacing: doublespacing
 linenumbers: no
 bibliography: components/LakeTrophicModelling.bib
-abstract:  Productivity of lentic ecosystems has been well studied and it is 
-  widely accepted that as nutrient inputs increase,  productivity increases and
-  lakes transition from low trophic state (e.g. oligotrophic) to higher trophic 
-  states  (e.g. eutrophic). These broad trophic state classifications are good 
-  predictors of ecosystem health and ecosystem services/disservices (e.g. recreation,
-  aesthetics, fisheries, and harmful algal blooms). While the relationship between 
-  nutrients and trophic state provides reliable predictions, it requires *in situ* 
-  water quality data in order to paramterize the model.  This limits the application 
-  of these models to lakes with existing and, more importanly, available water quality
-  data.  To expand our ability to predict in lakes without water quality data, we take
-  advantage of the availability of a large national lakes water quality database, 
-  land use/land cover data, lake morphometry data, other universally available data, 
-  and modern data mining approaches to build and assess models of lake tropic state 
-  that may be more universally applied.  We use random forests and random forest 
-  variable selection to identify variables to be used for predicting trophic state and 
-  we compare the performance of two models of trophic state (as determined by 
-  chlorophyll a concentration).   The first model estimates trophic state with *in situ* 
-  as well as universally available data and the second model uses universally available 
-  data only. For each of these models we used three separate trophic state categories, 
-  for a total of six models. Overall accuracy for the *in situ* and universal data 
-  models ranged from xx% to xx% and xx, xx, and xx described the most variation in 
-  trophic state.  For the universal data only models, Overall accuraccy ranged from xx% 
-  to xx% and xx, xx, and xx described the most variation in trophic state. Lastly, it 
-  is believed that the presence and abundance of cyanobacteria is strongly associated 
-  with trophic state.  To test this we examine the association between estimates of 
-  cyanobacteria biovolume and the measured and predicted trophic state.  Expanding 
-  these preliminary results to include cyanobacteria taxa indicates that cyanobacteria 
-  are significantly more likely to be found in highly eutrophic lakes. These results 
-  suggest that predictive models of lake trophic state may be improved with additional 
-  information on the landscape surrounding lakes and that those models provide 
-  additional information on the presence of potentially harmful cyanobacteria taxa. 
 ---
@@
 })
+```
+\begin{center} 
+Jeffrey W. Hollister*, W. Bryan Milstead, Betty J. Kreakie
 
-```
+*US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02879, USA*
+
+------
+*corresponding author:hollister.jeff@epa.gov
+\end{center}
+
+*#Abstract
+Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from low trophic state (e.g.
+oligotrophic) to higher trophic states  (e.g.
+eutrophic).
+These broad trophic state classifications are good predictors of ecosystem health and ecosystem services/disservices (e.g.
+recreation, aesthetics, fisheries, and harmful algal blooms).
+While the relationship between  nutrients and trophic state provides reliable predictions, it requires *in situ* water quality data in order to paramterize the model.
+This limits the application of these models to lakes with existing and, more importanly, available water quality data.
+To expand our ability to predict in lakes without water quality data, we take advantage of the availability of a large national lakes water quality database, land use/land cover data, lake morphometry data, other universally available data, and modern data mining approaches to build and assess models of lake tropic state that may be more universally applied.
+We use random forests and random forest variable selection to identify variables to be used for predicting trophic state and we compare the performance of two models of trophic state (as determined by chlorophyll *a* concentration).
+The first model estimates trophic state with *in situ* as well as universally available data and the second model uses universally available data only.
+For each of these models we used three separate trophic state categories, for a total of six models.
+Overall accuracy for the *in situ* and universal data models ranged from xx% to xx% and xx, xx, and xx described the most variation in trophic state.
+For the universal data only models, Overall accuraccy ranged from xx% to xx% and xx, xx, and xx described the most variation in trophic state.
+Lastly, it is believed that the presence and abundance of cyanobacteria is strongly associated with trophic state.
+To test this we examine the association between estimates of cyanobacteria biovolume and the measured and predicted trophic state.
+Expanding these preliminary results to include cyanobacteria taxa indicates that cyanobacteria are significantly more likely to be found in highly eutrophic lakes.
+These results suggest that predictive models of lake trophic state may be improved with additional information on the landscape surrounding lakes and that those models provide additional information on the presence of potentially harmful cyanobacteria taxa.
+\end{Abstract}
+\end{singlespace}
 
 #Introduction
-
 Productivity in lentic systems is often categorized across a range of tropic states (e.g.
@@
 ##Data and Study Area
-
 We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].
@@
 ##Predicting Trophic State with Random Forests
-
 Random forest is a machine learning algorithm that aggregates numerous decision trees in order to obtain a consensus prediction of the response categories [@breiman2001random].
@@
 ##Model Details
-
 Using a combination of the `varSelRF` and `randomForest` we ran models for six combinations of variables and trophic state classifications.
@@
 #Results
-
 ```{r figSetup,echo=FALSE,eval=TRUE,cache=TRUE}
@@
 ##Model 1: 4 Trophic States ~ All Variables
-
 ```{r ALL_TS4, echo=FALSE, eval=FALSE, cache=TRUE}

### 2014-09-12 vignettes/manuscript.R

@@ -150,11 +150,3 @@ plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
                               italic("a"),' Trophic States (2 Categories)')),
        color="Trophic State\nCategories")
 
-## ----scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot"----
-scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
-scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
-             title=expression(paste("Chlorophyll ", 
-                                    italic("a")," and Cyanobacteria Relationship")),
-             x=expression(paste('Log10(Chl ', italic("a"),')')),
-             y="Log10(Cyanobaterial Biovolumes + 1)")
-

### 2014-09-12 inst/doc/manuscript.Rmd
monkeying with latex

@@
 title: 'Modeling Lake Trophic State: A Data Mining Approach'
+authors: Jeffrey W. Hollister; W. Bryan Milstead; Betty J. Kreakie
 output:
@@
     number_sections: yes
-    template: components/elsarticle.latex
   html_document: null
@@
 csl: components/ecology.csl
-documentclass: elsarticle
+documentclass: article
 spacing: doublespacing
@@
 ```
-\begin{center} 
-Jeffrey W. Hollister*, W. Bryan Milstead, Betty J. Kreakie
 
@@
 
-------
-*corresponding author:hollister.jeff@epa.gov
-\end{center}
+*corresponding author:hollister.jeff@epa.gov*
 
-*#Abstract
+\begin{abstract}
 Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from low trophic state (e.g.
@@
 These results suggest that predictive models of lake trophic state may be improved with additional information on the landscape surrounding lakes and that those models provide additional information on the presence of potentially harmful cyanobacteria taxa.
-\end{Abstract}
-\end{singlespace}
+\end{abstract}
 

### 2014-09-12 vignettes/manuscript.R

@@ -1,152 +1,0 @@
-## ----setup, include=FALSE, echo=FALSE------------------------------------
-#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
-#                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-#devtools::install_github("wesanderson","jhollist")
-library("wesanderson")
-library("LakeTrophicModelling")
-library("knitr")
-library("ggplot2")
-library("sp")
-library("rgdal")
-library("xtable")
-library("e1071")
-opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
-data(hkm2014Data_all)
-
-#Checks for existing cache (from another project)
-if(file.exists("cache")){
-  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
-    lazyLoad(i)
-  }
-}
-
-# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
-
-knit_hooks$set(tab.cap = function(before, options, envir) {
-                  if(!before) { 
-                    paste('\n\n:', options$tab.cap, sep='') 
-                  }
-                })
-default_output_hook = knit_hooks$get("output")
-knit_hooks$set(output = function(x, options) {
-  if (is.null(options$tab.cap) == FALSE) {
-    x
-  } else
-    default_output_hook(x,options)
-})
-
-## ----Variables, results='asis', eval=FALSE, echo=FALSE-------------------
-#  #tab.cap="Definition of all variables considered"
[…]

### 2014-09-14 inst/doc/manuscript.R
Tried messing with placement of \begin{abstract}, moved back to front, fized authors yaml

@@ -1,6 +1,6 @@
 ## ----setup, include=FALSE, echo=FALSE------------------------------------
 #devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
-#                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+ #                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
 #devtools::install_github("wesanderson","jhollist")
 library("wesanderson")
 library("LakeTrophicModelling")

### 2014-09-15 inst/doc/manuscript.Rmd
devtools rebuild with new manuscript.latex template from manuscriptPackage

@@
 title: 'Modeling Lake Trophic State: A Data Mining Approach'
-authors: Jeffrey W. Hollister; W. Bryan Milstead; Betty J. Kreakie
+author: 
+- name: Jeffrey W. Hollister
+  affilnum: 1
+  email: hollister.jeff@epa.gov
+- name: W. Bryan Milstead
+  affilnum: 1
+- name: Betty J. Kreakie
+  affilnum: 1
+affiliation:
+- affilnum: 1
+  affil: US Environmental Protection Agency, Office of Research and Development, National     Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
 output:
@@
     number_sections: yes
+    template: components/manuscript.latex
   html_document: null
@@
 bibliography: components/LakeTrophicModelling.bib
+
 ---
+
 <!--
@@
 
-*US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02879, USA*
+```{r analysis , include=FALSE, echo=FALSE}
+#All analysis in here, that way all bits of the paper have access to the final objects
+#Place tables and figures and numerical results where they need to go.
+```
 
-*corresponding author:hollister.jeff@epa.gov*
-
+\begin{singlespace}
 \begin{abstract}
@@
 \end{abstract}
+\end{singlespace}
 
@@
 
-```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}"}
+```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}",cache=TRUE}
 state<-map_data('state')
@@
 
-```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}"}
+```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}",cache=TRUE}
 ts_4<-c("oligo","meso","eu","hyper")
@@
 
-```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}"}
+```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}",cache=TRUE}
 kable(sumTable(all_ts4[1:100]),row.names=F)
@@
 
-```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}"}
+```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}",cache=TRUE}
 kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1\\label{fig:Importance_Model1}"}
+```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1\\label{fig:Importance_Model1}",cache=TRUE}
 all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
@@
 
-```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2"}
+```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2",cache=TRUE}
 kable(sumTable(all_ts3[1:100]),row.names=F)
@@
 
-```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2"}
+```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2",cache=TRUE}
 kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2"}
+```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2",cache=TRUE}
 all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
@@
 
-```{r all_TS2_rf,echo=FALSE,eval=FALSE,cache=T}
+```{r all_TS2_rf,echo=FALSE,eval=FALSE,cache=TRUE}
 #Included for completeness.  Results already run in another project
@@
 
-```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3"}
+```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3",cache=TRUE}
 kable(sumTable(all_ts2[1:100]),row.names=F)
@@
 
-```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3"}
+```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3",cache=TRUE}
 kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3"}
+```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3",cache=TRUE}
 all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
@@
 
-```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4" }
+```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4",cache=TRUE }
 kable(sumTable(gis_ts4[1:100]),row.names=F)
@@
 
-```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4" }
+```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4",cache=TRUE }
 kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 4"}
+```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 4",cache=TRUE}
 gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
@@
 
-```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5"}
+```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5",cache=TRUE}
 kable(sumTable(gis_ts3[1:100]),row.names=F)
@@
 
-```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5"}
+```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5",cache=TRUE}
 kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5"}
+```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5",cache=TRUE}
 gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
@@
 
-```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6"}
+```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6",cache=TRUE}
 kable(sumTable(gis_ts2[1:100]),row.names=F)
@@
 
-```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6"}
+```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6",cache=TRUE}
 kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
@@
 
-```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6"}
+```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6",cache=TRUE}
 gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
@@
 
-```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes"}
+```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes",cache=TRUE}
 plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:9],
@@
 \newpage
-```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes" }
+```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes",cache=TRUE }
 plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
@@
 \newpage
-```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes"}
+```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes",cache=TRUE}
 plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
@@
 \newpage
-```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot"}
+```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot",cache=TRUE}
 scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)

### 2014-09-19 inst/doc/manuscript.Rmd
Updating analysis, new function added, and updated data

@@
 spacing: doublespacing
-linenumbers: no
+linenumbers: yes
 bibliography: components/LakeTrophicModelling.bib
-
 ---
-
 <!--
@@
 
-```{r analysis , include=FALSE, echo=FALSE}
+```{r analysis , eval=FALSE, include=FALSE, echo=FALSE}
+##########################################################################################
 #All analysis in here, that way all bits of the paper have access to the final objects
-#Place tables and figures and numerical results where they need to go.
+#Tables, figures, and numerical results placed at end of paper.
+#Included for completeness.  Results already run in another project
+#Those results loaded via lazyLoad in setup chunk
+##########################################################################################
+#Model 1: All Variables, 4 Trophic State Classes
+  #Variable Selction
+  all_ts4<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_4,numCore=4,100,
+        								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="all_ts4",time=TRUE)
+  #Random Forest
+  all_ts4_vars<-unique(unlist(all_ts4[1:100]))
+  all_ts4_rf<-runRandomForest(hkm2014Data[all_ts4_vars],hkm2014Data$TS_CHLA_4, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+##########################################################################################
+#Model 2: All Variables, 3 Trophic State Classes
+  #Variable Selction
+  all_ts3<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_3,numCore=4,100,
+      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="all_ts3",time=TRUE)
+  #Random Forest
+  all_ts3_vars<-unique(unlist(all_ts3[1:100]))
+  all_ts3_rf<-runRandomForest(hkm2014Data[all_ts3_vars],hkm2014Data$TS_CHLA_3, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+##########################################################################################
+#Model 3: All Variables, 2 Trophic State Classes
+  #Variable Selction
+  all_ts2<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_2,numCore=4,100,
+      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="all_ts2",time=TRUE)
+
+  #Random Forest
+  all_ts2_vars<-unique(unlist(all_ts2[1:100]))
+  all_ts2_rf<-runRandomForest(hkm2014Data[all_ts2_vars],hkm2014Data$TS_CHLA_2, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+##########################################################################################
+#Model 4: GIS Only Variables, 4 Trophic State Classes
+  #Variable Selction
+  gis_ts4<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_4,numCore=4,100,
+          							ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="gis_ts4",time=TRUE)
+
+  #Random Forest
+  gis_ts4_vars<-unique(unlist(gis_ts4[1:100]))
+  gis_ts4_rf<-runRandomForest(hkm2014Data[gis_ts4_vars],hkm2014Data$TS_CHLA_4, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+##########################################################################################
+#Model 5: GIS Only Variables, 3 Trophic State Classes
+  #Variable Selction
+  gis_ts3<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_3,numCore=4,100,
+        								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="gis_ts3",time=TRUE)
+
+  #Random Forest
+  gis_ts3_vars<-unique(unlist(gis_ts3[1:100]))
+  gis_ts3_rf<-runRandomForest(hkm2014Data[gis_ts3_vars],hkm2014Data$TS_CHLA_3, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+##########################################################################################
+#Model 6: GIS Only Variables, 2 Trophic State Classes
+  #Variable Selction
+  gis_ts2<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_2,numCore=4,100,
+      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  											vars.drop.num=1,outStr="gis_ts2",time=TRUE)
+  #Random Forest
+  gis_ts2_vars<-unique(unlist(gis_ts2[1:100]))
+  gis_ts2_rf<-runRandomForest(hkm2014Data[gis_ts2_vars],hkm2014Data$TS_CHLA_2, 
+                              ntree=10000,importance=TRUE,proximity=TRUE)
+
 ```
 
-\begin{singlespace}
-\begin{abstract}
+```{r summaryStats, echo=FALSE, cache=TRUE}
+acc_results<-data.frame(model_number=c("Model 1","Model 2","Model 3","Model 4","Model 5","Model 6"),
+                        total_accuracy=c(classAgreement(all_ts4_rf$confusion[,1:4])$diag,
+                                         classAgreement(all_ts3_rf$confusion[,1:3])$diag,
+                                         classAgreement(all_ts2_rf$confusion[,1:2])$diag,
+                                         classAgreement(gis_ts4_rf$confusion[,1:4])$diag,
+                                         classAgreement(gis_ts3_rf$confusion[,1:3])$diag,
+                                         classAgreement(gis_ts2_rf$confusion[,1:2])$diag),
+                        kappa_coeff=c(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,
+                                         classAgreement(all_ts3_rf$confusion[,1:3])$kappa,
+                                         classAgreement(all_ts2_rf$confusion[,1:2])$kappa,
+                                         classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,
+                                         classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,
+                                         classAgreement(gis_ts2_rf$confusion[,1:2])$kappa))
+
+```
+
+\singlespace
+\hrule
+
+**Abstract**
+
 Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from low trophic state (e.g.
@@
 For each of these models we used three separate trophic state categories, for a total of six models.
-Overall accuracy for the *in situ* and universal data models ranged from xx% to xx% and xx, xx, and xx described the most variation in trophic state.
-For the universal data only models, Overall accuraccy ranged from xx% to xx% and xx, xx, and xx described the most variation in trophic state.
+Overall accuracy for the *in situ* and universal data models ranged from `r round(min(acc_results[1:3,2]),3)`% to `r round(max(acc_results[1:3,2]),3)`% and xx, xx, and xx described the most variation in trophic state.
+For the universal data only models, Overall accuraccy ranged from `r round(min(acc_results[4:6,2]),3)`% to `r round(max(acc_results[4:6,2]),3)`% and xx, xx, and xx described the most variation in trophic state.
 Lastly, it is believed that the presence and abundance of cyanobacteria is strongly associated with trophic state.
@@
 These results suggest that predictive models of lake trophic state may be improved with additional information on the landscape surrounding lakes and that those models provide additional information on the presence of potentially harmful cyanobacteria taxa.
-\end{abstract}
-\end{singlespace}
+
+\hrule
+\doublespace
 
@@
 ##Model 1: 4 Trophic States ~ All Variables
-```{r ALL_TS4, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts4<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_4,numCore=4,100,
-    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="all_ts4",time=TRUE)
-```
-
-```{r ALL_TS4_rf,echo=FALSE,eval=FALSE,cache=T}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts4_vars<-unique(unlist(all_ts4[1:100]))
-all_ts4_rf<-runRandomForest(hkm2014Data[all_ts4_vars],hkm2014Data$TS_CHLA_4, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 The selected variables that made up Model 1 were Potassium, Nitrogen:Phosphorus, Total Nitrogen, Total Phosphorus, Total Organic Carbon, Turbidity, Ecoregion, Organic Ions, Dissolved Organic Carbon, and Maximum Lake Depth (Table \ref{tab:VarSel_Model1}).
@@
 
-```{r ALL_TS3, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts3<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_3,numCore=4,100,
-  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="all_ts3",time=TRUE)
-```
-
-```{r all_TS3_rf,echo=FALSE,eval=FALSE,cache=T}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts3_vars<-unique(unlist(all_ts3[1:100]))
-all_ts3_rf<-runRandomForest(hkm2014Data[all_ts3_vars],hkm2014Data$TS_CHLA_3, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
@@
 ##Model 3: 2 Trophic States ~ All Variables
-
-```{r ALL_TS2, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts2<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_2,numCore=4,100,
-  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="all_ts2",time=TRUE)
-```
-
-```{r all_TS2_rf,echo=FALSE,eval=FALSE,cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-all_ts2_vars<-unique(unlist(all_ts2[1:100]))
-all_ts2_rf<-runRandomForest(hkm2014Data[all_ts2_vars],hkm2014Data$TS_CHLA_2, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
@@
 ##Model 4: 4 Trophic States ~ GIS Only Variables
-
-```{r GIS_TS4, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts4<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_4,numCore=4,100,
-      								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="gis_ts4",time=TRUE)
-```
-
-```{r GIS_TS4_rf,echo=FALSE,eval=FALSE,cache=T}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts4_vars<-unique(unlist(gis_ts4[1:100]))
-gis_ts4_rf<-runRandomForest(hkm2014Data[gis_ts4_vars],hkm2014Data$TS_CHLA_4, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
@@
 ##Model 5: 3 Trophic States ~ GIS Only Variables
-
-```{r GIS_TS3, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts3<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_3,numCore=4,100,
-    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="gis_ts3",time=TRUE)
-```
-
-```{r GIS_TS3_rf,echo=FALSE,eval=FALSE,cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts3_vars<-unique(unlist(gis_ts3[1:100]))
-gis_ts3_rf<-runRandomForest(hkm2014Data[gis_ts3_vars],hkm2014Data$TS_CHLA_3, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
@@
 ##Model 6: 2 Trophic States ~ GIS Only Variables
-
-```{r GIS_TS2, echo=FALSE, eval=FALSE, cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts2<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_2,numCore=4,100,
-  										ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-											vars.drop.num=1,outStr="gis_ts2",time=TRUE)
-```
-
-```{r GIS_TS2_rf,echo=FALSE,eval=FALSE,cache=TRUE}
-#Included for completeness.  Results already run in another project
-#Those results loaded via lazyLoad in setup chunk
-gis_ts2_vars<-unique(unlist(gis_ts2[1:100]))
-gis_ts2_rf<-runRandomForest(hkm2014Data[gis_ts2_vars],hkm2014Data$TS_CHLA_2, ntree=10000,importance=TRUE,proximity=TRUE)
-```
-
 Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.

### 2014-09-19 R/varImportance.R

@@ -1,0 +1,19 @@
+#' Variable Importance Summary
+#' 
+#' @param rf Random Forest Object
+#' @param label A label for the model
+#' @export
+varImportance<-function(rf,label=NULL){
+  if(!is.null(label)){
+    xdf<-data.frame(model_id=rep(label,dim(rf$importance)[1]),
+             variable_names=row.names(rf$importance),
+             mean_decrease_acc=data.frame(rf$importance)$MeanDecreaseAccuracy,
+             mean_decrease_gini=data.frame(rf$importance)$MeanDecreaseGini)
+  } else {
+    xdf<-data.frame(model_id=rep(NA,dim(rf$importance)[1]),
+                    variable_names=row.names(rf$importance),
+                    mean_decrease_acc=data.frame(rf$importance)$MeanDecreaseAccuracy,
+                    mean_decrease_gini=data.frame(rf$importance)$MeanDecreaseGini)
+  }   
+  return(xdf)
+}

### 2014-09-19 inst/doc/manuscript.R

@@ -1,0 +1,120 @@
+## ----setup, include=FALSE, echo=FALSE------------------------------------
+#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+ #                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+#devtools::install_github("wesanderson","jhollist")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("xtable")
+library("e1071")
+opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+data(hkm2014Data_all)
+
+#Checks for existing cache (from another project)
+if(file.exists("cache")){
+  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+
+# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+
+knit_hooks$set(tab.cap = function(before, options, envir) {
+                  if(!before) { 
+                    paste('\n\n:', options$tab.cap, sep='') 
+                  }
+                })
+default_output_hook = knit_hooks$get("output")
+knit_hooks$set(output = function(x, options) {
+  if (is.null(options$tab.cap) == FALSE) {
+    x
+  } else
+    default_output_hook(x,options)
+})
+
+## ----analysis , eval=FALSE, include=FALSE, echo=FALSE--------------------
+#  ##########################################################################################
+#  #All analysis in here, that way all bits of the paper have access to the final objects
+#  #Tables, figures, and numerical results placed at end of paper.
+#  #Included for completeness.  Results already run in another project
+#  #Those results loaded via lazyLoad in setup chunk
+#  ##########################################################################################
+#  #Model 1: All Variables, 4 Trophic State Classes
+#    #Variable Selction
+#    all_ts4<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_4,numCore=4,100,
+#          								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
[…]

### 2014-09-22 inst/doc/manuscript.R
cleaning up Rmd - moved fig and tab to end

@@ -1,7 +1,7 @@
 ## ----setup, include=FALSE, echo=FALSE------------------------------------
-#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
- #                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-#devtools::install_github("wesanderson","jhollist")
+#devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE,
+#                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+#devtools::install_github("jhollist/wesanderson")
 library("wesanderson")
 library("LakeTrophicModelling")
 library("knitr")
@@ -10,17 +10,20 @@ library("sp")
 library("rgdal")
 library("xtable")
 library("e1071")
+library("dplyr")
+library("tidyr")
 opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
-data(hkm2014Data_all)
+data(LakeTrophicModelling)
 
 #Checks for existing cache (from another project)
-if(file.exists("cache")){
-  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+if(file.exists("prior_cache")){
+  for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
     lazyLoad(i)
   }
 }
 
-# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+# Table Captions from @DeanK on 
+#http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
 
 knit_hooks$set(tab.cap = function(before, options, envir) {
                   if(!before) { 
@@ -34,87 +37,4 @@ knit_hooks$set(output = function(x, options) {
   } else
     default_output_hook(x,options)
 })
-
-## ----analysis , eval=FALSE, include=FALSE, echo=FALSE--------------------
-#  ##########################################################################################
-#  #All analysis in here, that way all bits of the paper have access to the final objects
-#  #Tables, figures, and numerical results placed at end of paper.
-#  #Included for completeness.  Results already run in another project
[…]

### 2014-09-23 inst/doc/manuscript.Rmd
saving line endings

@@
 bibliography: components/LakeTrophicModelling.bib
+abstract: no
 ---
@@
 ```{r setup, include=FALSE, echo=FALSE}
-#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
+#devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE,
 #                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-#devtools::install_github("wesanderson","jhollist")
+#devtools::install_github("jhollist/wesanderson")
 library("wesanderson")
@@
 library("e1071")
+library("dplyr")
+library("tidyr")
 opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
-data(hkm2014Data_all)
+data(LakeTrophicModelling)
 
 #Checks for existing cache (from another project)
-if(file.exists("cache")){
-  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
+if(file.exists("prior_cache")){
+  for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
     lazyLoad(i)
@@
 
-# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+# Table Captions from @DeanK on 
+#http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
 
@@
 
-```{r analysis , eval=FALSE, include=FALSE, echo=FALSE}
-##########################################################################################
-#All analysis in here, that way all bits of the paper have access to the final objects
+```{r analysis , eval=FALSE, include=FALSE, echo=FALSE, cache=TRUE}
+################################################################################
+#All analysis in here, that way all bits of paper have access to final objects
 #Tables, figures, and numerical results placed at end of paper.
@@
 #Those results loaded via lazyLoad in setup chunk
-##########################################################################################
+#This chunk takes several days to complete - DO NOT EDIT!!!
+################################################################################
 #Model 1: All Variables, 4 Trophic State Classes
   #Variable Selction
-  all_ts4<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_4,numCore=4,100,
+  all_ts4<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_4,numCore=4,100,
         								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   all_ts4_vars<-unique(unlist(all_ts4[1:100]))
-  all_ts4_rf<-runRandomForest(hkm2014Data[all_ts4_vars],hkm2014Data$TS_CHLA_4, 
+  all_ts4_rf<-runRandomForest(ltmData[all_ts4_vars],ltmData$TS_CHLA_4, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
-##########################################################################################
+################################################################################
 #Model 2: All Variables, 3 Trophic State Classes
   #Variable Selction
-  all_ts3<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_3,numCore=4,100,
+  all_ts3<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_3,numCore=4,100,
       									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   all_ts3_vars<-unique(unlist(all_ts3[1:100]))
-  all_ts3_rf<-runRandomForest(hkm2014Data[all_ts3_vars],hkm2014Data$TS_CHLA_3, 
+  all_ts3_rf<-runRandomForest(ltmData[all_ts3_vars],ltmData$TS_CHLA_3, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
-##########################################################################################
+################################################################################
 #Model 3: All Variables, 2 Trophic State Classes
   #Variable Selction
-  all_ts2<-iterVarSelRF(hkm2014Data[predictors_all],hkm2014Data$TS_CHLA_2,numCore=4,100,
+  all_ts2<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_2,numCore=4,100,
       									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   all_ts2_vars<-unique(unlist(all_ts2[1:100]))
-  all_ts2_rf<-runRandomForest(hkm2014Data[all_ts2_vars],hkm2014Data$TS_CHLA_2, 
+  all_ts2_rf<-runRandomForest(ltmData[all_ts2_vars],ltmData$TS_CHLA_2, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
-##########################################################################################
+################################################################################
 #Model 4: GIS Only Variables, 4 Trophic State Classes
   #Variable Selction
-  gis_ts4<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_4,numCore=4,100,
+  gis_ts4<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,numCore=4,100,
           							ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   gis_ts4_vars<-unique(unlist(gis_ts4[1:100]))
-  gis_ts4_rf<-runRandomForest(hkm2014Data[gis_ts4_vars],hkm2014Data$TS_CHLA_4, 
+  gis_ts4_rf<-runRandomForest(ltmData[gis_ts4_vars],ltmData$TS_CHLA_4, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
-##########################################################################################
+################################################################################
 #Model 5: GIS Only Variables, 3 Trophic State Classes
   #Variable Selction
-  gis_ts3<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_3,numCore=4,100,
+  gis_ts3<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_3,numCore=4,100,
         								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   gis_ts3_vars<-unique(unlist(gis_ts3[1:100]))
-  gis_ts3_rf<-runRandomForest(hkm2014Data[gis_ts3_vars],hkm2014Data$TS_CHLA_3, 
+  gis_ts3_rf<-runRandomForest(ltmData[gis_ts3_vars],ltmData$TS_CHLA_3, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
-##########################################################################################
+################################################################################
 #Model 6: GIS Only Variables, 2 Trophic State Classes
   #Variable Selction
-  gis_ts2<-iterVarSelRF(hkm2014Data[predictors_gis],hkm2014Data$TS_CHLA_2,numCore=4,100,
+  gis_ts2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_2,numCore=4,100,
       									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
@@
   gis_ts2_vars<-unique(unlist(gis_ts2[1:100]))
-  gis_ts2_rf<-runRandomForest(hkm2014Data[gis_ts2_vars],hkm2014Data$TS_CHLA_2, 
+  gis_ts2_rf<-runRandomForest(ltmData[gis_ts2_vars],ltmData$TS_CHLA_2, 
                               ntree=10000,importance=TRUE,proximity=TRUE)
@@
 
-```{r summaryStats, echo=FALSE, cache=TRUE}
+```{r summaryStats, echo=FALSE, cache=TRUE, warning=FALSE}
 acc_results<-data.frame(model_number=c("Model 1","Model 2","Model 3","Model 4","Model 5","Model 6"),
@@
                                          classAgreement(gis_ts2_rf$confusion[,1:2])$kappa))
+var_importance<-rbind_list(varImportance(all_ts4_rf,"Model 1"),
+                           varImportance(all_ts3_rf,"Model 2"),
+                           varImportance(all_ts2_rf,"Model 3"),
+                           varImportance(gis_ts4_rf,"Model 4"),
+                           varImportance(gis_ts3_rf,"Model 5"),
+                           varImportance(gis_ts2_rf,"Model 6"))
+var_importance$variable_names<-factor(var_importance$variable_names)
+#var_importance$model_id<-factor(var_importance$model_id)
+var_importance<-inner_join(var_importance,data_def[c("variable_names","description")],"variable_names")
+var_importance<-var_importance[order(var_importance$model_id,-var_importance$mean_decrease_gini, decreasing=FALSE),]
 
@@
 \singlespace
-\hrule
+
+\vspace{2mm}\hrule
 
 **Abstract**
-
 Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from low trophic state (e.g.
@@
 For each of these models we used three separate trophic state categories, for a total of six models.
-Overall accuracy for the *in situ* and universal data models ranged from `r round(min(acc_results[1:3,2]),3)`% to `r round(max(acc_results[1:3,2]),3)`% and xx, xx, and xx described the most variation in trophic state.
-For the universal data only models, Overall accuraccy ranged from `r round(min(acc_results[4:6,2]),3)`% to `r round(max(acc_results[4:6,2]),3)`% and xx, xx, and xx described the most variation in trophic state.
+Overall accuracy for models buils from *in situ* and universal data ranged from `r round(min(acc_results[1:3,2]),3)`% to `r round(max(acc_results[1:3,2]),3)`%.
+For the universal data only models, Overall accuraccy ranged from `r round(min(acc_results[4:6,2]),3)`% to `r round(max(acc_results[4:6,2]),3)`%.
 Lastly, it is believed that the presence and abundance of cyanobacteria is strongly associated with trophic state.
-To test this we examine the association between estimates of cyanobacteria biovolume and the measured and predicted trophic state.
+To test this we examine the association between estimates of cyanobacteria biovolume and the measured and predicted trophic state and find a positive relationship.
 Expanding these preliminary results to include cyanobacteria taxa indicates that cyanobacteria are significantly more likely to be found in highly eutrophic lakes.
@@
 
-\hrule
+\vspace{2mm}\hrule
 \doublespace
@@
 oligotrophic)to late successional lakes (i.e.
-hypereutrophic) [@carlson1977trophic].
-Lakes naturally occur across the range of trophic state and higher primary productivity is not necessarily a predictor of poor ecological condition.
-Lakes that are naturally oligotrophic occur in nutrient poor areas or have a more recent geologic history.
+hypereutrophic) [@carlson1977trophic] and lakes naturally occur across this range.
+Naturally oligotrophic lakes occur in nutrient poor areas or have a more recent geologic history.
 These lakes are often found in higher elevations, have clear water, and are often favored for drinking water or direct contact recreation (e.g.
@@
 eutrophic lakes) have greater nutrient loads, tend to be less clear, have greater density of aquatic plants, and often support more diverse and abundant fish communities.
-Lakes will naturally shift to higher trophic states but this is a slow process.
+Higher primary productivity is not necessarily a predictor of poor ecological condition.
+
+It is natural for lakes to shift from lower to higher trophic states but this is a slow process.
 Given this fact, monitoring trophic state allows the identification of rapid shifts in trophic state or locating lakes with unusually high productivity (e.g.
@@
 
-Cyanobacteria are an important taxonomic group associated with harmful algal blooms in lakes.
-Understanding the drivers of cyanobacteria presence has important implications for lake management and for the protection of human and ecosystem health.
-Chlorophyll a concentration, a measure of the biological productivity of a lake, is one such driver and is largely, although not exclusively, determined by nutrient inputs.
-As nutrient inputs increase, productivity increases and lakes transition from low trophic state (e.g.
-oligotrophic) to higher trophic states (e.g.
-hypereutrophic).
-These broad trophic state classifications are associated with ecosystem health and ecosystem services/disservices (e.g.
-recreation, aesthetics, fisheries, and harmful algal blooms).
-Thus, models of trophic state might be used to predict things like cyanobacteria.
+As trophic state can be defined by a number of *in situ* water quality measurements, most models have used this information as predictors.
+This leads to accurate models, but also requires data that is sparse and not always availble, limiting the population of lakes that can be modeled.
+Landscape data is ubiqutious ...
 
@@
 
-```{r Variables, results='asis', eval=FALSE, echo=FALSE}
-#tab.cap="Definition of all variables considered"
-translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space","Percent Low Intensity Development","Percent Medium Intensity Development", "Percent High Intensity Development","Percent Barren","Percent Decidous Forest","Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland", "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter","Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area","Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity","Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon","Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate","Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions","Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum","Ion Balance","Estimate Organic Anions","Calculated Conductivity","D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days","Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", "Fetch form Northeast","Fetch from East", "Fetch from Southeast","Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth","Nitrogen:Phophorus Ratio")
-wq<-!predictors_all%in%predictors_gis
-type<-vector("character",length(predictors_all))
-type[wq]<-"Water Quality"
-type[!wq]<-"GIS"
-data_def<-data.frame(Variables=predictors_all,Description=translation,Type=type)
-# This is an unbelievably silly table... It is life sized!
-kable(data_def)
-#dfToImage(data_def,style="css/texttable.css",file="figure/data_def_table.jpg")
-```
-
-```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}",cache=TRUE}
-state<-map_data('state')
-lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
-p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
-ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
-lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
-lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
-lakes_dd<-data.frame(coordinates(lakes_dd))
-mycolor<-wes.palette(5,"Zissou2")
-mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
-names(lakes_dd)<-c("long","lat",c())
-nlaMap(state,lakes_dd,mycolor)
-```
-
 ##Predicting Trophic State with Random Forests
@@
 
-Using R’s randomForest package, we pass the reduced models selected with varSelRF and calculate confusion matrices, overall accuracy and kappa coeffecient [@liaw2002randomForest].
+Using R's randomForest package, we pass the reduced models selected with varSelRF and calculate confusion matrices, overall accuracy and kappa coeffecient [@liaw2002randomForest].
 From the reduced model random forests we collect a consensus prediction and calculate a confusion matrix and summary stats.
@@
 
-```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}",cache=TRUE}
-ts_4<-c("oligo","meso","eu","hyper")
-ts_3<-c("oligo","meso/eu","meso/eu","hyper")
-ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
-co<-c("<= 0.2",">2-7",">7-30",">30")
-xdf<-data.frame(ts_4,ts_3,ts_2,co)
-names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
-kable(xdf)
-```
+#Results and Discussion
 
-#Results
+##Model 1: 4 Trophic States ~ All Variables
+The selected variables that made up Model 1 were Potassium, Nitrogen:Phosphorus, Total Nitrogen, Total Phosphorus, Total Organic Carbon, Turbidity, Ecoregion, Organic Ions, Dissolved Organic Carbon, and Maximum Lake Depth (Table \ref{tab:VarSel_Model1}).
+Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)` (Table \ref{tab:Confusion_Model1}).
+
+Lastly, tubidity, total phosphorus, total nitrogen, and total organic carbon were the most important predictors of the 4 classes of trophic state (Figure \ref{fig:Importance_Model1}).
+
+##Model 2: 3 Trophic States ~ All Variables
+
+Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+##Model 3: 2 Trophic States ~ All Variables
+Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+##Model 4: 4 Trophic States ~ GIS Only Variables
+Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
+
+##Model 5: 3 Trophic States ~ GIS Only Variables
+Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
+
+##Model 6: 2 Trophic States ~ GIS Only Variables
+Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
+
+##Associating Trophic State and Cyanobacteria
+Arm waving goes here.
+
+\newpage
+
+#Figures
+
 ```{r figSetup,echo=FALSE,eval=TRUE,cache=TRUE}
@@
 
-##Model 1: 4 Trophic States ~ All Variables
-The selected variables that made up Model 1 were Potassium, Nitrogen:Phosphorus, Total Nitrogen, Total Phosphorus, Total Organic Carbon, Turbidity, Ecoregion, Organic Ions, Dissolved Organic Carbon, and Maximum Lake Depth (Table \ref{tab:VarSel_Model1}).
-Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)` (Table \ref{tab:Confusion_Model1}).
-
-```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}",cache=TRUE}
-kable(sumTable(all_ts4[1:100]),row.names=F)
+```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}",cache=TRUE}
+state<-map_data('state')
+lakes_alb<-data.frame(ltmData[["AlbersX"]],ltmData[["AlbersY"]])
+p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
+lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
+lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
+lakes_dd<-data.frame(coordinates(lakes_dd))
+mycolor<-wes.palette(5,"Zissou2")
+mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
+names(lakes_dd)<-c("long","lat",c())
+nlaMap(state,lakes_dd,mycolor)
 ```
 
-```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}",cache=TRUE}
-kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
-```
-
-Lastly, tubidity, total phosphorus, total nitrogen, and total organic carbon were the most important predictors of the 4 classes of trophic state (Figure \ref{fig:Importance_Model1}).
+\newpage
 
@@
 
-##Model 2: 3 Trophic States ~ All Variables
-
-Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
-
-```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2",cache=TRUE}
-kable(sumTable(all_ts3[1:100]),row.names=F)
-```
-
-```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2",cache=TRUE}
-kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
-```
+\newpage
 
@@
 
-##Model 3: 2 Trophic States ~ All Variables
-Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
-
-```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3",cache=TRUE}
-kable(sumTable(all_ts2[1:100]),row.names=F)
-```
-
-```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3",cache=TRUE}
-kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
-```
+\newpage
 
@@
 
-##Model 4: 4 Trophic States ~ GIS Only Variables
-Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
-
-```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4",cache=TRUE }
-kable(sumTable(gis_ts4[1:100]),row.names=F)
-```
-
-```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4",cache=TRUE }
-kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
-```
+\newpage
 
@@
 
-##Model 5: 3 Trophic States ~ GIS Only Variables
-Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
-
-```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5",cache=TRUE}
-kable(sumTable(gis_ts3[1:100]),row.names=F)
-```
-
-```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5",cache=TRUE}
-kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
-```
+\newpage
 
@@
 
-##Model 6: 2 Trophic States ~ GIS Only Variables
-Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
-
-```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6",cache=TRUE}
-kable(sumTable(gis_ts2[1:100]),row.names=F)
-```
-
-```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6",cache=TRUE}
-kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
-```
+\newpage
 
@@
 
-##Associating Trophic State and Cyanobacteria
+\newpage
 
 ```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes",cache=TRUE}
-plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:9],
+plotCdf(ltmData$TS_CHLA_4,ltmData$sumBioV+1,cdf_colors=zissou[6:9],
         y='Percent',x='Log10(Cyanobacteria Biovolume)',
@@
 ```
+
 \newpage
+
 ```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes",cache=TRUE }
-plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
+plotCdf(ltmData$TS_CHLA_3,ltmData$sumBioV+1,cdf_colors=zissou[6:8],
        y='Percent',x='Log10(Cyanobacteria Biovolume)', 
@@
 ```
+
 \newpage
+
 ```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes",cache=TRUE}
-plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
+plotCdf(ltmData$TS_CHLA_2,ltmData$sumBioV+1,cdf_colors=zissou[6:7],
        y='Percent',x='Log10(Cyanobacteria Biovolume)',
@@
 ```
+
 \newpage
+
 ```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot",cache=TRUE}
-scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
+scp_df<-data.frame(chla=ltmData[["CHLA"]],biovp1=ltmData[["sumBioV"]]+1)
 scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
@@
 
+#Tables
+
+```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}",cache=TRUE}
+ts_4<-c("oligo","meso","eu","hyper")
+ts_3<-c("oligo","meso/eu","meso/eu","hyper")
+ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
+co<-c("<= 0.2",">2-7",">7-30",">30")
+xdf<-data.frame(ts_4,ts_3,ts_2,co)
+names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
+kable(xdf)
+```
+
+\newpage
+
+```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}",cache=TRUE}
+kable(sumTable(all_ts4[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}",cache=TRUE}
+kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
+```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2",cache=TRUE}
+kable(sumTable(all_ts3[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2",cache=TRUE}
+kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
+```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3",cache=TRUE}
+kable(sumTable(all_ts2[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3",cache=TRUE}
+kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
+```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4",cache=TRUE }
+kable(sumTable(gis_ts4[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4",cache=TRUE }
+kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
+```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5",cache=TRUE}
+kable(sumTable(gis_ts3[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5",cache=TRUE}
+kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
+```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6",cache=TRUE}
+kable(sumTable(gis_ts2[1:100]),row.names=F)
+```
+
+\newpage
+
+```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6",cache=TRUE}
+kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
+```
+
+\newpage
+
 #References

### 2014-09-23 R/dfToImage.R
fixing problems after moving

@@ -1,41 +1,0 @@
-#' dfToImage
-#' 
-#' This function will take a data.frame, convert it to a table, style it with 
-#' CSS (optional) and save as an image file.
-#' 
-#' @param df the data.frame to convert 
-#' @param file the output image file to save to include file type in the 
-#'        extension.  Have tested , 'png', and 'jpg'
-#' @param style the path/name for the style sheet.
-#' @param width width of output image.  defaults to 1024.
-#' @param include.rownames logical to print or not the row names.
-#' @param format.args list of formatC paramaters to control formating of text
-#' @param keepHTML Logical if html file should be saved (best for troubleshooting
-#'                  the css styling)
-#' @param ... parameters passed through to \code{\link{xtable}}.  For instance,
-#'            allows use of \code{caption} option. 
-#' 
-#' @examples
-#' dfToImage(head(mtcars),'test.jpg')
-#' 
-#' @export
-dfToImage <- function(df, file, style = NULL, width = 1024, 
-    include.rownames = FALSE, keepHTML = FALSE, ...) {
-    tmpfile <- paste(tempfile(, tmpdir = getwd()), 
-        ".html", sep = "")
-    print(xtable(df, ...), type = "html", file = tmpfile, 
-        include.rownames = include.rownames)
-    if (!is.null(style)) {
-        csslink <- paste("<link rel='stylesheet' href='", 
-            style, "' type='text/css' />\n", sep = "")
-        fileConn <- file(tmpfile, open = "r+")
-        Lines <- readLines(fileConn)
-        writeLines(c(csslink, Lines), con = fileConn)
-        close(fileConn)
-    }
-    system(paste("wkhtmltoimage --quality 100 --width", 
-        width, tmpfile, file, sep = " "))
-    if (!keepHTML) {
-        system(paste("rm", tmpfile))
-    }
-} 

### 2014-09-23 R/iterVarSelRF.R

@@ -16,10 +16,10 @@
 #' @param time logical true will output overall time for the run.  
 #' 
 #' @examples
-#' data(hkm2014Data)
-#' x1<-iterVarSelRF(hkm2014Data[predictors_gis],TS_CHLA_4,10,ntree=10,ntreeIterat=5,
+#' data(LakeTrophicModelling)
+#' x1<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,10,ntree=10,ntreeIterat=5,
 #' vars.drop.frac=NULL,vars.drop.num=1,time=TRUE)
-#' x2<-iterVarSelRF(hkm2014Data[predictors_gis],TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
+#' x2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
 #' vars.drop.frac=NULL,vars.drop.num=1,time=TRUE)
 #' @export
 #' @import varSelRF snowfall

### 2014-09-23 R/runRandomForest.R

@@ -12,11 +12,11 @@
 #'                  function.  This was a quick hack.
 #' 
 #' @examples
-#' data(hkm2014Data)
-#' x2<-iterVarSelRF(hkm2014Data[predictors_gis],TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
+#' data(LakeTrophicModelling)
+#' x2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
 #' vars.drop.frac=NULL,vars.drop.num=1,time=FALSE)
 #' x2_vars<-unique(unlist(x2))
-#' gis_ts4_rf<-runRandomForest(hkm2014Data[x2_vars],TS_CHLA_4, ntree=10,importance=TRUE,proximity=TRUE)
+#' gis_ts4_rf<-runRandomForest(ltmData[x2_vars],ltmData$TS_CHLA_4, ntree=10,importance=TRUE,proximity=TRUE)
 #' @export
 #' @import randomForest
 runRandomForest <- function(indVar, depVar, ntree = 100, 

### 2014-09-23 vignettes/manuscript.R

@@ -1,0 +1,39 @@
+## ----setup, include=FALSE, echo=FALSE------------------------------------
+#devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE,
+#                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+#devtools::install_github("jhollist/wesanderson")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("e1071")
+library("dplyr")
+library("tidyr")
+opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+data(LakeTrophicModelling)
+
+#Checks for existing cache (from another project)
+if(file.exists("prior_cache")){
+  for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+
+# Table Captions from @DeanK on 
+#http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+
+knit_hooks$set(tab.cap = function(before, options, envir) {
+                  if(!before) { 
+                    paste('\n\n:', options$tab.cap, sep='') 
+                  }
+                })
+default_output_hook = knit_hooks$get("output")
+knit_hooks$set(output = function(x, options) {
+  if (is.null(options$tab.cap) == FALSE) {
+    x
+  } else
+    default_output_hook(x,options)
+})
+

### 2014-09-23 inst/doc/manuscript.Rmd
still fixing shit

@@
     template: components/manuscript.latex
-  html_document: null
-  word_document: null
 fontsize: 11pt
@@
 ---
+
 <!--
@@
 library("rgdal")
-library("xtable")
 library("e1071")
@@
 
-\vspace{2mm}\hrule
+\vspace{3mm}\hrule
 \doublespace

### 2014-09-23 R/nlaMap.R

@@ -8,7 +8,7 @@
 #' 
 #' @examples
 #' state<-map_data('state')
-#' lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
+#' lakes_alb<-data.frame(ltmData[["AlbersX"]],ltmData[["AlbersY"]])
 #' p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" 
 #' ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs" 
 #' lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))

### 2014-09-23 R/plotCdf.R

@@ -14,11 +14,11 @@
 #' @examples
 #' devtools::install_github('wesanderson','karthik')
 #' library(wesanderson)
-#' data(hkm2014Data)
-#' plotCdf(hkm2014Data$bvCat,hkm2014Data$CHLA,cdf_colors=wes.palette(3,'Royal1'),y='Percent',
+#' data(LakeTrophicModelling)
+#' plotCdf(ltmData$bvCat,ltmData$CHLA,cdf_colors=wes.palette(3,'Royal1'),y='Percent',
 #'         x=expression(paste('Log10(Chl ', italic("a"),')')),title='CDF for category bvCat',
 #'         color="Biovolume\nCategories")
-#' plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=wes.palette(3,'FantasticFox'),
+#' plotCdf(ltmData$TS_CHLA_3,ltmData$sumBioV+1,cdf_colors=wes.palette(3,'FantasticFox'),
 #'         y='Percent',x='Log10(Cyanobacterial Biovolume)',
 #'         title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
 #'         color="Trophic State\nCategories")

### 2014-09-23 R/scatterPlot.R

@@ -5,8 +5,8 @@
 #' @examples
 #' devtools::install_github("wesanderson","jhollist")
 #' library('wesanderson')
-#' data(hkm2014Data)
-#' mydf<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
+#' data(LakeTrophicModelling)
+#' mydf<-data.frame(chla=ltmData[["CHLA"]],biovp1=ltmData[["sumBioV"]]+1)
 #' zissou2<-wes.palette(5,"Zissou2")
 #' scatterPlot(mydf,xvar="chla",yvar="biovp1",zissou2[3],zissou[2],zissou[1],
 #'             title="Chlorophyll and Cyanobacterai Relationship",

### 2014-09-23 inst/doc/manuscript.R

@@ -8,7 +8,6 @@ library("knitr")
 library("ggplot2")
 library("sp")
 library("rgdal")
-library("xtable")
 library("e1071")
 library("dplyr")
 library("tidyr")
@@ -37,4 +36,231 @@ knit_hooks$set(output = function(x, options) {
   } else
     default_output_hook(x,options)
 })
+
+## ----analysis , eval=FALSE, include=FALSE, echo=FALSE, cache=TRUE--------
+#  ################################################################################
+#  #All analysis in here, that way all bits of paper have access to final objects
+#  #Tables, figures, and numerical results placed at end of paper.
+#  #Included for completeness.  Results already run in another project
+#  #Those results loaded via lazyLoad in setup chunk
+#  #This chunk takes several days to complete - DO NOT EDIT!!!
+#  ################################################################################
+#  #Model 1: All Variables, 4 Trophic State Classes
+#    #Variable Selction
+#    all_ts4<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_4,numCore=4,100,
+#          								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+#    											vars.drop.num=1,outStr="all_ts4",time=TRUE)
+#    #Random Forest
+#    all_ts4_vars<-unique(unlist(all_ts4[1:100]))
+#    all_ts4_rf<-runRandomForest(ltmData[all_ts4_vars],ltmData$TS_CHLA_4,
+#                                ntree=10000,importance=TRUE,proximity=TRUE)
+#  ################################################################################
+#  #Model 2: All Variables, 3 Trophic State Classes
+#    #Variable Selction
+#    all_ts3<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_3,numCore=4,100,
+#        									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+#    											vars.drop.num=1,outStr="all_ts3",time=TRUE)
+#    #Random Forest
+#    all_ts3_vars<-unique(unlist(all_ts3[1:100]))
+#    all_ts3_rf<-runRandomForest(ltmData[all_ts3_vars],ltmData$TS_CHLA_3,
[…]

### 2014-09-24 R/plotCdf.R
update plot font size, updated manuscript

@@ -38,13 +38,13 @@ plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
           theme(text = element_text(family="Times"),
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 
-                plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
+                plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
                 legend.position = c(0.11,0.8), legend.key = element_rect(fill = 'white'),
-                legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=15),
-                axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
-                axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
-                axis.text.x = element_text(family="serif",size = 15),
-                axis.text.y = element_text(family="serif",size = 15)) + 
+                legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=11),
+                axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
+                axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
+                axis.text.x = element_text(family="serif",size = 11),
+                axis.text.y = element_text(family="serif",size = 11)) + 
           geom_hline(linetype = 3, size = 1, colour = "gray", 
                      yintercept = c(0,0.5,1)) +
           labs(...)

### 2014-09-27 R/importancePlot.R
added some text to methods

@@ -45,9 +45,9 @@ importancePlot <- function(rf, type = c("acc", "gini"), ...) {
          theme(text = element_text(family="Times"),
                panel.background = element_blank(), panel.grid = element_blank(), 
                panel.border = element_rect(fill = NA), legend.position = "none", 
-               axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
-               axis.text.x = element_text(family="serif",size = 11),
-               axis.text.y = element_text(family="serif",size = 11)) + 
+               axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 20), 
+               axis.text.x = element_text(family="serif",size = 15),
+               axis.text.y = element_text(family="serif",size = 15)) + 
          ylab("") + 
          xlab(label)
     

### 2014-09-27 R/plotCdf.R

@@ -38,13 +38,13 @@ plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
           theme(text = element_text(family="Times"),
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 
-                plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
+                plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
                 legend.position = c(0.11,0.8), legend.key = element_rect(fill = 'white'),
-                legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=11),
-                axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
-                axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
-                axis.text.x = element_text(family="serif",size = 11),
-                axis.text.y = element_text(family="serif",size = 11)) + 
+                legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=15),
+                axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
+                axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
+                axis.text.x = element_text(family="serif",size = 15),
+                axis.text.y = element_text(family="serif",size = 15)) + 
           geom_hline(linetype = 3, size = 1, colour = "gray", 
                      yintercept = c(0,0.5,1)) +
           labs(...)

### 2014-09-27 R/scatterPlot.R

@@ -21,7 +21,7 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
   xdf<-xdf[complete.cases(xdf),]
 
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=2,colour=pt_col) +
+    geom_point(size=4,colour=pt_col) +
     scale_x_log10() +
     scale_y_log10() +
     stat_smooth(method="lm",fill=cl_col,
@@ -29,13 +29,13 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
     theme(text = element_text(family="serif"),
           panel.background = element_blank(), panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
-          plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
+          plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
           legend.position = "none", legend.key = element_rect(fill = 'white'),
-          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=11),
-          axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
-          axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
-          axis.text.x = element_text(family="serif",size = 11),
-          axis.text.y = element_text(family="serif",size = 11)) + 
+          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=15),
+          axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
+          axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
+          axis.text.x = element_text(family="serif",size = 15),
+          axis.text.y = element_text(family="serif",size = 15)) + 
     labs(y="log10(Cyanobacteria Biovolume + 1)",...)
   return(x)
 }

### 2014-10-20 R/getLakeIDs.R
add getlakes some more in the ms.

@@ -1,0 +1,31 @@
+#' getLakeIDs
+#' 
+#' This function retrieves the lake IDs for each of the rf objects
+#' 
+#' @param indVar independent variables
+#' @param depVar dependent variable
+#' @param id vector of IDS, same length and order as indVar and 
+#' 
+#' @examples
+#' data(LakeTrophicModelling)
+#' x2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,10,4,ntree=10,ntreeIterat=5,
+#' vars.drop.frac=NULL,vars.drop.num=1,time=FALSE)
+#' x2_vars<-unique(unlist(x2))
+#' gis_ts4_rf<-runRandomForest(ltmData[x2_vars],ltmData$TS_CHLA_4)
+#' @export
+#' 
+getLakeIDs <- function(indVar, depVar, id) {
+  
+  # Only Pass Complete Cases
+  comp_case <- complete.cases(indVar, depVar)
+  if (length(indVar) > 1) {
+    indVar <- indVar[comp_case, ]
+  } else {
+    indVar <- indVar[comp_case]
+  }
+  depVar <- depVar[comp_case]
+  
+  
+  return(id[comp_case])
+  
+} 

### 2014-10-31 R/condAccuracy.R
added text on conditional accuracy and start of conditional accuracy function

@@ -1,0 +1,12 @@
+#' Plot conditional accuracy
+#' 
+#' This function plots the total accuracy as a function of maximum probability. 
+#' We determine the maximum probability for each lake for a given random forest
+#' object and then plot the total accuracy for lakes at each value of x or greater
+#' 
+#' @param rf A random forest object
+#' @param ... parameters to be passed to condprob
+#' @export
+condAccuracy<-function(rf,...){
+  
+}

### 2014-10-31 R/condAccuracy.R
added conditional accuracy figure

@@ -7,6 +7,29 @@
 #' @param rf A random forest object
 #' @param ... parameters to be passed to condprob
 #' @export
-condAccuracy<-function(rf,...){
+#' @import condprob2 ggplot2
+#' @examples
+#' condAccuracy(gis_ts4_rf,xImpair=0,R=1,xlab="Maximum Vote Probability")
+condAccuracy<-function(rf,xlab="x",...){
+  max_vote<-apply(rf$votes,1,max)
+  obs<-rf$y
+  pred<-factor(rf$predicted,levels=unique(obs),ordered=is.ordered(obs))
+  correct<-pred==obs
+  cp<-condprob(max_vote,correct,ProbComp="gt",Exceed="gte",...)
+  #plot(cp$max_vote,cp$Raw.Data.Probability)
+  cp<-data.frame(cp$max_vote,cp$Raw.Data.Probability)
   
+  ggplot(cp,aes(x=cp.max_vote,y=cp.Raw.Data.Probability))+
+    geom_point()+
+    theme(text = element_text(family="serif"),
+          panel.background = element_blank(), #panel.grid = element_blank(), 
+          panel.border = element_rect(fill = NA), 
+          plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
+          legend.position = "none", legend.key = element_rect(fill = 'white'),
+          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=15),
+          axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
+          axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
+          axis.text.x = element_text(family="serif",size = 15),
+          axis.text.y = element_text(family="serif",size = 15)) + 
+    labs(y="Conditional Total Accuracy",x=xlab)
 }

### 2014-11-01 R/condAccuracy.R
fixed condprob2 imports in condAccuracy.R

@@ -7,15 +7,16 @@
 #' @param rf A random forest object
 #' @param ... parameters to be passed to condprob
 #' @export
-#' @import condprob2 ggplot2
+#' @import ggplot2
 #' @examples
 #' condAccuracy(gis_ts4_rf,xImpair=0,R=1,xlab="Maximum Vote Probability")
 condAccuracy<-function(rf,xlab="x",...){
+  devtools::install_github("jhollist/condprob2")
   max_vote<-apply(rf$votes,1,max)
   obs<-rf$y
   pred<-factor(rf$predicted,levels=unique(obs),ordered=is.ordered(obs))
   correct<-pred==obs
-  cp<-condprob(max_vote,correct,ProbComp="gt",Exceed="gte",...)
+  cp<-condprob2::condprob(max_vote,correct,ProbComp="gt",Exceed="gte",...)
   #plot(cp$max_vote,cp$Raw.Data.Probability)
   cp<-data.frame(cp$max_vote,cp$Raw.Data.Probability)
   

### 2014-11-01 R/importancePlot.R

@@ -45,9 +45,9 @@ importancePlot <- function(rf, type = c("acc", "gini"), ...) {
          theme(text = element_text(family="Times"),
                panel.background = element_blank(), panel.grid = element_blank(), 
                panel.border = element_rect(fill = NA), legend.position = "none", 
-               axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 20), 
-               axis.text.x = element_text(family="serif",size = 15),
-               axis.text.y = element_text(family="serif",size = 15)) + 
+               axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
+               axis.text.x = element_text(family="serif",size = 11),
+               axis.text.y = element_text(family="serif",size = 11)) + 
          ylab("") + 
          xlab(label)
     

### 2014-11-01 R/scatterPlot.R

@@ -21,7 +21,7 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
   xdf<-xdf[complete.cases(xdf),]
 
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=4,colour=pt_col) +
+    geom_point(size=2,colour=pt_col) +
     scale_x_log10() +
     scale_y_log10() +
     stat_smooth(method="lm",fill=cl_col,
@@ -29,13 +29,13 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
     theme(text = element_text(family="serif"),
           panel.background = element_blank(), panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
-          plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
+          plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
           legend.position = "none", legend.key = element_rect(fill = 'white'),
-          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=15),
-          axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
-          axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
-          axis.text.x = element_text(family="serif",size = 15),
-          axis.text.y = element_text(family="serif",size = 15)) + 
+          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=11),
+          axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="serif",size = 11),
+          axis.text.y = element_text(family="serif",size = 11)) + 
     labs(y="log10(Cyanobacteria Biovolume + 1)",...)
   return(x)
 }

### 2014-11-04 R/condAccuracy.R
more script updates and manuscript updates

@@ -11,7 +11,7 @@
 #' @examples
 #' condAccuracy(gis_ts4_rf,xImpair=0,R=1,xlab="Maximum Vote Probability")
 condAccuracy<-function(rf,xlab="x",...){
-  devtools::install_github("jhollist/condprob2")
+  #devtools::install_github("jhollist/condprob2")
   max_vote<-apply(rf$votes,1,max)
   obs<-rf$y
   pred<-factor(rf$predicted,levels=unique(obs),ordered=is.ordered(obs))

### 2014-11-19 R/importancePlot.R
additional edits and new importance plots

@@ -15,7 +15,7 @@
 #' importancePlot(iris_rf,'acc',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 #' importancePlot(iris_rf,'gini',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 
-importancePlot <- function(rf, type = c("acc", "gini"), ...) {
+importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
     
     idf<-importance(rf)
     # set up data.frame for ggplot
@@ -25,17 +25,18 @@ importancePlot <- function(rf, type = c("acc", "gini"), ...) {
         o <- order(imp_df$Mean_Decrease, decreasing = FALSE)
         imp_df$variables <- factor(imp_df$variables, 
             levels = imp_df$variables[o], ordered = T)
-        label <- "Mean Decrease Accuracy"
+        label <- "Mean Decrease Gini"
     } else if (type == "acc") {
         imp_df <- data.frame(variables = names(idf[,dim(idf)[2]-1]), 
                              Mean_Decrease = idf[,dim(idf)[2]-1])
       o <- order(imp_df$Mean_Decrease, decreasing = FALSE)
       imp_df$variables <- factor(imp_df$variables, 
                                  levels = imp_df$variables[o], ordered = T)
-        label <- "Mean Decrease Gini"
+        label <- "Mean Decrease Accuracy"
     } else {
         stop("Type not specificed correctly")
     }
+    imp_df<-merge(imp_df,sumtbl,by.x="variables",by.y="Variable")
     
     x <- ggplot(imp_df, aes(Mean_Decrease, variables)) + 
          geom_point(...) +   
@@ -47,9 +48,25 @@ importancePlot <- function(rf, type = c("acc", "gini"), ...) {
                panel.border = element_rect(fill = NA), legend.position = "none", 
                axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
                axis.text.x = element_text(family="serif",size = 11),
-               axis.text.y = element_text(family="serif",size = 11)) + 
+               axis.text.y = element_text(family="serif",size = 11),
+               plot.margin = unit(c(0.25,0,0.1,0),"inches")) + 
          ylab("") + 
          xlab(label)
+    y <- ggplot(imp_df, aes(Percent, variables)) + 
+      geom_point(...) +   
+      geom_hline(linetype = 3, size = 1, colour = "gray", 
[…]

### 2014-12-05 R/importancePlot.R
fixed grid dependency in importancePlot, more on draft

@@ -49,7 +49,7 @@ importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
                axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
                axis.text.x = element_text(family="serif",size = 11),
                axis.text.y = element_text(family="serif",size = 11),
-               plot.margin = unit(c(0.25,0,0.1,0),"inches")) + 
+               plot.margin = grid::unit(c(0.25,0,0.1,0),"inches")) + 
          ylab("") + 
          xlab(label)
     y <- ggplot(imp_df, aes(Percent, variables)) + 
@@ -64,7 +64,7 @@ importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
             axis.text.x = element_text(family="serif",size = 11),
             axis.text.y = element_blank(),
             axis.ticks = element_blank(),
-            plot.margin = unit(c(0.25,0.875,0.1,0),"inches")) + 
+            plot.margin = grid::unit(c(0.25,0.875,0.1,0),"inches")) + 
       ylab("") + 
       xlab("Proportion of varSelRF runs selected")
     

### 2014-12-15 R/importancePlot.R
updated figures and finalized all co-author edits, ready for internal review

@@ -15,9 +15,10 @@
 #' importancePlot(iris_rf,'acc',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 #' importancePlot(iris_rf,'gini',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 
-importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
+importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"), ...) {
     
     idf<-importance(rf)
+    
     # set up data.frame for ggplot
     if (type == "gini") {
         imp_df <- data.frame(variables = names(idf[,dim(idf)[2]]), 
@@ -38,7 +39,27 @@ importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
     }
     imp_df<-merge(imp_df,sumtbl,by.x="variables",by.y="Variable")
     
+    if (!is.null(data_def)){
+      imp_df$variables<-data_def$description[match(imp_df$variables,data_def$variable_names)]
+      o <- order(imp_df$Mean_Decrease, decreasing = FALSE)
+      imp_df$variables <- factor(imp_df$variables, 
+                                 levels = imp_df$variables[o], ordered = T)
+    }
+    
     x <- ggplot(imp_df, aes(Mean_Decrease, variables)) + 
+      geom_blank() +  
+      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 25))+
+      theme(text = element_text(family="Times"),
+            panel.background = element_blank(), panel.grid = element_blank(), 
+            panel.border = element_blank(), 
+            axis.title.x = element_blank(),
+            axis.text.x = element_blank(),
+            axis.text.y = element_text(family="serif",size = 11),
+            axis.ticks.x = element_blank(),
+            plot.margin = grid::unit(c(0.25,0,0.3,0.6),"inches")) + 
+      ylab("") + 
+      xlab(label)
+    y <- ggplot(imp_df, aes(Mean_Decrease, variables)) + 
          geom_point(...) +   
          geom_hline(linetype = 3, size = 1, colour = "gray", 
                  yintercept = 1:nlevels(imp_df$variables)) + 
@@ -48,11 +69,12 @@ importancePlot <- function(rf, sumtbl, type = c("acc", "gini"), ...) {
                panel.border = element_rect(fill = NA), legend.position = "none", 
[…]

### 2014-12-15 R/plotCdf.R

@@ -39,7 +39,7 @@ plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 
                 plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
-                legend.position = c(0.11,0.8), legend.key = element_rect(fill = 'white'),
+                legend.position = c(0.15,0.75), legend.key = element_rect(fill = 'white'),
                 legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=11),
                 axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
                 axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),

### 2014-12-15 vignettes/manuscript.R

@@ -1,0 +1,208 @@
+## ----setup, include=FALSE, echo=FALSE------------------------------------
+#devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE,
+#                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
+#devtools::install_github("jhollist/wesanderson")
+#devtools::install_github("jhollist/condprob2")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("e1071")
+library("dplyr")
+library("tidyr")
+library("condprob2")
+opts_chunk$set(dev = 'png', fig.width=7.5)
+data(LakeTrophicModelling)
+
+
+#Checks for existing cache (from another project)
+if(file.exists("prior_cache")){
+  for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+
+
+# Table Captions from @DeanK on 
+#http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
+
+knit_hooks$set(tab.cap = function(before, options, envir) {
+                  if(!before) { 
+                    paste('\n\n:', options$tab.cap, sep='') 
+                  }
+                })
+default_output_hook = knit_hooks$get("output")
+knit_hooks$set(output = function(x, options) {
+  if (is.null(options$tab.cap) == FALSE) {
+    x
+  } else
+    default_output_hook(x,options)
+})
+
+## ----analysis , eval=FALSE, include=FALSE, echo=FALSE, cache=FALSE-------
+#  ################################################################################
+#  #All analysis in here, that way all bits of paper have access to final objects
+#  #Tables, figures, and numerical results placed at end of paper.
+#  #Included for completeness.  Results already run in another project
+#  #Those results loaded via lazyLoad in setup chunk
+#  #This chunk takes several days to complete - DO NOT EDIT!!!
+#  ################################################################################
[…]

### 2015-02-09 R/condAccuracy.R
finishing up Nates edits

@@ -25,12 +25,12 @@ condAccuracy<-function(rf,xlab="x",...){
     theme(text = element_text(family="serif"),
           panel.background = element_blank(), #panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
-          plot.title  = element_text(family="serif",size=22,face="bold",vjust=1.1),
+          plot.title  = element_text(family="serif",size=15,face="bold",vjust=1.1),
           legend.position = "none", legend.key = element_rect(fill = 'white'),
           legend.text = element_text(family="serif",size=15), legend.title = element_text(size=15),
-          axis.title.x = element_text(family="serif",vjust = -0.5, size = 20),
-          axis.title.y = element_text(family="serif",vjust = 1.5, size = 20),
-          axis.text.x = element_text(family="serif",size = 15),
-          axis.text.y = element_text(family="serif",size = 15)) + 
+          axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="serif",size = 11),
+          axis.text.y = element_text(family="serif",size = 11)) + 
     labs(y="Conditional Total Accuracy",x=xlab)
 }

### 2015-02-09 R/scatterPlot.R

@@ -8,7 +8,7 @@
 #' data(LakeTrophicModelling)
 #' mydf<-data.frame(chla=ltmData[["CHLA"]],biovp1=ltmData[["sumBioV"]]+1)
 #' zissou2<-wes.palette(5,"Zissou2")
-#' scatterPlot(mydf,xvar="chla",yvar="biovp1",zissou2[3],zissou[2],zissou[1],
+#' scatterPlot(mydf,xvar="chla",yvar="biovp1",zissou2[3],zissou2[2],zissou2[1],
 #'             title="Chlorophyll and Cyanobacterai Relationship",
 #'             x=expression(paste('Log10(Chl ', italic("a"),')')),
 #'             y="Log10(Cyanobaterial Biovolumes + 1)")
@@ -19,13 +19,22 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
   options(scipen=5)
   xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]])
   xdf<-xdf[complete.cases(xdf),]
-
+  xlm<-summary(lm(log10(yvar+1)~log10(xvar+1),data=xdf))
+  xlm_r<-paste0("Adj. R-squared = ", round(xlm$adj.r.squared,2))
+  xlm_p<-paste0("p-value = ", format(pf(xlm$fstatistic[1],
+                                         xlm$fstatistic[2],
+                                         xlm$fstatistic[3],lower.tail=FALSE),
+                                      scientific=TRUE))              
+  
+  
   x <- ggplot(xdf,aes(xvar,yvar))+
     geom_point(size=2,colour=pt_col) +
     scale_x_log10() +
     scale_y_log10() +
     stat_smooth(method="lm",fill=cl_col,
-                colour=ln_col,size=2) +
+                colour=ln_col,size=2) + 
+    annotate("text", x = 0.25, y = 100000000, label = xlm_r) +
+    annotate("text", x = 0.285, y = 20000000, label = xlm_p) +
     theme(text = element_text(family="serif"),
           panel.background = element_blank(), panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
@@ -37,5 +46,6 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
           axis.text.x = element_text(family="serif",size = 11),
           axis.text.y = element_text(family="serif",size = 11)) + 
[…]

### 2015-04-03 inst/doc/manuscript.Rmd
final clearance version

@@
 ---
-title: 'Modeling Lake Trophic State: A Data Mining Approach'
+title: 'Modeling Lake Trophic State: A Random Forest Approach'
 author:
-- name: Jeffrey W. Hollister
+- affilnum: 1
+  email: hollister.jeff@epa.gov
+  name: Jeffrey W. Hollister
+- affilnum: 1
+  name: W. Bryan Milstead
+- affilnum: 1
+  name: Betty J. Kreakie
+affiliation:
+- affil: US Environmental Protection Agency, Office of Research and Development, National     Health
+    and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell
+    Drive  Narragansett, RI, 02882, USA
   affilnum: 1
-  email: hollister.jeff@epa.gov
-- name: W. Bryan Milstead
-  affilnum: 1
-- name: Betty J. Kreakie
-  affilnum: 1
-affiliation:
-- affilnum: 1
-  affil: US Environmental Protection Agency, Office of Research and Development, National     Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
 output:
@@
     template: components/manuscript.latex
+  word_document: yes
+capsize: normalsize
+csl: components/freshwater-science.csl
+documentclass: article
 fontsize: 11pt
-capsize: normalsize
-csl: components/ecology.csl
-documentclass: article
-spacing: doublespacing
 linenumbers: yes
 bibliography: components/LakeTrophicModelling.bib
+spacing: doublespacing
 abstract: no
@@
 ```{r setup, include=FALSE, echo=FALSE}
-#devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE,
-#                        auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
-#devtools::install_github("jhollist/wesanderson")
+if(!require(LakeTrophicModelling)){
+  devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE)
+}
+if(!require(wesanderson)){
+devtools::install_github("jhollist/wesanderson")
+}
+if(!require(condprob2)){
+devtools::install_github("jhollist/condprob2")
+}
+if(!require(tidyr)){
+  install.packages("tidyr")
+}
 library("wesanderson")
@@
 library("tidyr")
-opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
+library("condprob2")
+opts_chunk$set(dev = 'png', fig.width=7.5)
 data(LakeTrophicModelling)
@@
 #Checks for existing cache (from another project)
+#Else if Repeats running chunk outside of knitr
 if(file.exists("prior_cache")){
   for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+} else if(file.exists("vignettes/prior_cache")){
+  for(i in gsub(".rdb","",list.files("vignettes/prior_cache",".rdb",full.names=T))){
     lazyLoad(i)
@@
 
-```{r analysis , eval=FALSE, include=FALSE, echo=FALSE, cache=TRUE}
+```{r analysis , eval=FALSE, include=FALSE, echo=FALSE, cache=FALSE}
 ################################################################################
@@
   #Variable Selction
-  all_ts4<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_4,numCore=4,100,
-        								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="all_ts4",time=TRUE)
+  #all_ts4<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_4,numCore=4,100,
+  #      								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #  											vars.drop.num=1,outStr="all_ts4",time=TRUE)
+  #all_ts4_rf_full<-runRandomForest(ltmData[predictors_all],ltmData$TS_CHLA_4, 
+  #                          ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
   #Variable Selction
-  all_ts3<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_3,numCore=4,100,
-      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="all_ts3",time=TRUE)
+  #all_ts3<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_3,numCore=4,100,
+  #    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #  											vars.drop.num=1,outStr="all_ts3",time=TRUE)
+  # all_ts3_rf_full<-runRandomForest(ltmData[predictors_all],ltmData$TS_CHLA_3, 
+  #                            ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
   #Variable Selction
-  all_ts2<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_2,numCore=4,100,
-      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="all_ts2",time=TRUE)
-
+  #all_ts2<-iterVarSelRF(ltmData[predictors_all],ltmData$TS_CHLA_2,numCore=4,100,
+  #    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #											vars.drop.num=1,outStr="all_ts2",time=TRUE)
+  #all_ts2_rf_full<-runRandomForest(ltmData[predictors_all],ltmData$TS_CHLA_2, 
+  #                           ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
   #Variable Selction
-  gis_ts4<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,numCore=4,100,
-          							ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="gis_ts4",time=TRUE)
-
+  #gis_ts4<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_4,numCore=4,100,
+  #        							ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #											vars.drop.num=1,outStr="gis_ts4",time=TRUE)
+  #gis_ts4_rf_full<-runRandomForest(ltmData[predictors_gis],ltmData$TS_CHLA_4, 
+  #                            ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
   #Variable Selction
-  gis_ts3<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_3,numCore=4,100,
-        								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="gis_ts3",time=TRUE)
-
+  #gis_ts3<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_3,numCore=4,100,
+  #      								ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #											vars.drop.num=1,outStr="gis_ts3",time=TRUE)
+  #gis_ts3_rf_full<-runRandomForest(ltmData[predictors_gis],ltmData$TS_CHLA_3, 
+  #                            ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
   #Variable Selction
-  gis_ts2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_2,numCore=4,100,
-      									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
-  											vars.drop.num=1,outStr="gis_ts2",time=TRUE)
+  #gis_ts2<-iterVarSelRF(ltmData[predictors_gis],ltmData$TS_CHLA_2,numCore=4,100,
+  #    									ntree=10000,ntreeIterat=5000,vars.drop.frac=NULL,
+  #											vars.drop.num=1,outStr="gis_ts2",time=TRUE)
+  #gis_ts2_rf_full<-runRandomForest(ltmData[predictors_gis],ltmData$TS_CHLA_2, 
+  #                            ntree=10000,importance=TRUE,proximity=TRUE)
   #Random Forest
@@
 
-```{r summaryStats, echo=FALSE, cache=TRUE, warning=FALSE}
+```{r summaryStats, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE}
 acc_results<-data.frame(model_number=c("Model 1","Model 2","Model 3","Model 4","Model 5","Model 6"),
@@
                                          classAgreement(gis_ts2_rf$confusion[,1:2])$kappa))
-var_importance<-rbind_list(varImportance(all_ts4_rf,"Model 1"),
+var_importance<-rbind(varImportance(all_ts4_rf,"Model 1"),
                            varImportance(all_ts3_rf,"Model 2"),
@@
 
+#Example Votes
+lakeVotes<-data.frame(all_ts4_rf$y,all_ts4_rf$predicted,all_ts4_rf$votes)
+lakeVotes<-data.frame(lakeVotes,max_prob=apply(lakeVotes[,3:6],1,max))
+lakeVotes$all_ts4_rf.predicted<-factor(lakeVotes$all_ts4_rf.predicted,ordered=T)
+lakeVotes<-data.frame(lakeVotes,correct_binary=lakeVotes$all_ts4_rf.y==lakeVotes$all_ts4_rf.predicted)
+lakeVotes<-data.frame(COMID=getLakeIDs(ltmData[all_ts4_vars],ltmData$TS_CHLA_4,ltmData$Comid),lakeVotes)
+lakeVotes<-na.omit(lakeVotes)
 ```
@@
 **Abstract**  
-Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from low trophic state (e.g.
+Productivity of lentic ecosystems has been well studied and it is widely accepted that as nutrient inputs increase,  productivity increases and lakes transition from lower trophic state (e.g.
 oligotrophic) to higher trophic states  (e.g.
 eutrophic).
-These broad trophic state classifications are good predictors of ecosystem health and ecosystem services/disservices (e.g.
+These broad trophic state classifications are good predictors of ecosystem condition and ecosystem services and disservices (e.g.
 recreation, aesthetics, fisheries, and harmful algal blooms).
-While the relationship between  nutrients and trophic state provides reliable predictions, it requires *in situ* water quality data in order to paramterize the model.
-This limits the application of these models to lakes with existing and, more importanly, available water quality data.
-To expand our ability to predict in lakes without water quality data, we take advantage of the availability of a large national lakes water quality database, land use/land cover data, lake morphometry data, other universally available data, and modern data mining approaches to build and assess models of lake tropic state that may be more universally applied.
-We use random forests and random forest variable selection to identify variables to be used for predicting trophic state and we compare the performance of two models of trophic state (as determined by chlorophyll *a* concentration).
-The first model estimates trophic state with *in situ* as well as universally available data and the second model uses universally available data only.
-For each of these models we used three separate trophic state categories, for a total of six models.
-Overall accuracy for models buils from *in situ* and universal data ranged from `r round(min(acc_results[1:3,2]),3)`% to `r round(max(acc_results[1:3,2]),3)`%.
-For the universal data only models, Overall accuraccy ranged from `r round(min(acc_results[4:6,2]),3)`% to `r round(max(acc_results[4:6,2]),3)`%.
-Lastly, it is believed that the presence and abundance of cyanobacteria is strongly associated with trophic state.
-To test this we examine the association between estimates of cyanobacteria biovolume and the measured and predicted trophic state and find a positive relationship.
-Expanding these preliminary results to include cyanobacteria taxa indicates that cyanobacteria are significantly more likely to be found in highly eutrophic lakes.
+While the relationship between  nutrients and trophic state provides reliable predictions, it requires *in situ* water quality data in order to parameterize the model.
+This limits the application of these models to lakes with existing and, more importantly, available water quality data.
+To expand our ability to predict trophic state in lakes without water quality data, we take advantage of the availability of a large national lakes water quality database (i.e.
+the National Lakes Assessment), land use/land cover data, lake morphometry data, other universally available data, and modern data mining approaches to build and assess models of lake trophic state that may be more universally applied.
+We use random forests and random forest variable selection to identify variables to be used for predicting trophic state and we compare the performance of two sets of models of trophic state (as determined by chlorophyll *a* concentration).
+The first set of models estimates three different trophic state classification with *in situ* as well as universally available data and the second set of models uses universally available GIS data only.
+Overall accuracy for models built from *in situ* and universal data ranged from `r round(min(acc_results[1:3,2]),2)*100`% to `r round(max(acc_results[1:3,2]),2)*100`%.
+For the universal data only models, overall accuracy ranged from `r round(min(acc_results[4:6,2]),2)*100`% to `r round(max(acc_results[4:6,2]),2)*100`%.
+Lastly, presence and abundance of cyanobacteria is often strongly associated with trophic state.
+To test this we examine the association between estimates of cyanobacteria abundance and measured chlorophyll *a* and find a positive relationship.
 These results suggest that predictive models of lake trophic state may be improved with additional information on the landscape surrounding lakes and that those models provide additional information on the presence of potentially harmful cyanobacteria taxa.
+Lastly, the source code and data for this manuscript are freely available from https://github.com/USEPA/LakeTrophicModelling.
 
@@
 #Introduction
-Productivity in lentic systems is often categorized across a range of tropic states (e.g.
-the tropic continuum) from early succesional (i.e.
-oligotrophic)to late successional lakes (i.e.
-hypereutrophic) [@carlson1977trophic] and lakes naturally occur across this range.
-Naturally oligotrophic lakes occur in nutrient poor areas or have a more recent geologic history.
-These lakes are often found in higher elevations, have clear water, and are often favored for drinking water or direct contact recreation (e.g.
+Productivity in lentic systems is often categorized across a range of trophic states (e.g.
+the trophic continuum) from early successional (i.e.
+oligotrophic) to late successional lakes (i.e.
+hypereutrophic) with lakes naturally occurring across this range [@carlson1977trophic].
+Oligotrophic lakes occur in nutrient poor areas or have a more recent geologic history, are often found in higher elevations, have clear water, and are usually favored for drinking water or direct contact recreation (e.g.
 swimming).
 Lakes with higher productivity (e.g.
-eutrophic lakes) have greater nutrient loads, tend to be less clear, have greater density of aquatic plants, and often support more diverse and abundant fish communities.
-Higher primary productivity is not necessarily a predictor of poor ecological condition.
+mesotrophic and eutrophic lakes) have greater nutrient loads, tend to be less clear, have greater density of aquatic plants, and often support more diverse and abundant fish communities.
+Higher primary productivity is not necessarily a predictor of poor ecological condition as it is natural for lakes to shift from lower to higher trophic states but this is a slow process.
+However, at the highest productivity levels (hypereutrophic lakes) biological integrity is compromised [@hasler1969cultural; @schindler2008algal; @smith1999eutrophication].
 
-It is natural for lakes to shift from lower to higher trophic states but this is a slow process.
-Given this fact, monitoring trophic state allows the identification of rapid shifts in trophic state or locating lakes with unusually high productivity (e.g.
+Monitoring trophic state allows the identification of rapid shifts in trophic state or locating lakes with unusually high productivity (e.g.
 hypereutrophic).
-These cases are indicative of lakes under greater anthropogenic nutrient loads, also known as cultural eutrophication, and are more likely to be at risk of fish kills, fouling, and harmful algal blooms[@smith1998cultural;@smith1999eutrophication;@smith2006eutrophication].
-Given the association between trophic state and many ecosystem services and disservices, being able to model trophic state could allow for estimating trophic state in unmonitored lakes and provide a first cut at identifying lakes with the potential for harmful algal blooms and other problems associated with cultural eutrophication.
+These cases are indicative of lakes under greater anthropogenic nutrient loads, also known as cultural eutrophication, and are more likely to be at risk of fish kills, fouling, and harmful algal blooms [@smith1998cultural;@smith1999eutrophication;@smith2006eutrophication].
+Given the association between trophic state and many ecosystem services and disservices, being able to accurately model trophic state could provide a first cut at identifying lakes with the potential for harmful algal blooms (i.e.
+from cyanobacteria) or other problems associated with cultural eutrophication.
+This type of information could be used for setting priorities for managment and allow for more effecient use of limited resources.
 
-As trophic state can be defined by a number of *in situ* water quality measurements, most models have used this information as predictors.
-This leads to accurate models, but also requires data that is sparse and not always availble, limiting the population of lakes that can be modeled.
-Landscape data is ubiqutious ...
+As trophic state and related indices can be best defined by a number of *in situ* water quality parameters (modeled or measured), most models have used this information as predictors [e.g., @carvalho_cyanobacterial_2011;@milstead2013estimating;@imboden1978dynamic;@salas1991simplified].
+This leads to accurate models, but also requires data that are often sparse and not always available, thus limiting the population of lakes for which we can make predictions.
+A possible solution for this is to build models that use widely available data that are correlated to many of the *in situ* variables.
+For instance, landscape metrics of forests, agriculture, wetlands, and urban land in contributing watersheds have all been shown to explain a significant proportion of the variation (ranging from 50-86%, depending on study) in nutrients in receiving waters [@jones2001predicting; @jones2004importance; @seilheimer2013landscape].
+Building on these previously identified associations might allow us to use only landscape and other universally available data to build models.
+Identifying predictors using this type of ubiquitous data would allow for estimating trophic state in both monitored and unmonitored lakes.
 
-We have three goals for this preliminary research.
-First, we build and assess multiple models of lake trophic state using a full suite of data including *in situ* water quality and universally available data (e.g.
-landscape data).
-Second, we assess the accuracy of predicted trophic state in lakes with only the universally available data.
-Lastly, we explore associations between trophic state and cyanobacteria to explore.
+Many published models of nutrients and trophic state in freshwater systems are based on linear modelling methods such as standard least squares regression or linear mixed models [e.g., @jones2004importance; @jones2001predicting].
+While these methods have proven to be reliable, they have limitations (e.g.
+independence and distribution assumptions, and outlier sensitivity).
+Using data mining approaches, such as random forests, avoids many of the limitations, may reduce bias and often provides better predictions [@cutler2007random; @peters2007random; @breiman2001random].
+For instance, random forests are non-parametric and thus the data do not need to come from a specific distribution (e.g.
+Gaussian) and can contain collinear variables [@cutler2007random].
+Second, random forests work well with very large numbers of predictors [@cutler2007random].
+Lastly, random forests can deal with model selection uncertainty as predictions are based upon a consensus of many models and not just a single model selected with some measure of goodness of fit.
+
+To build on past work, we have identified five areas in which our research contributes.
+First, we update trophic state modelling efforts with the use of random forests.
+Second, we assess the accuracy of predicted trophic state in lakes with the a) *in situ* and universally available GIS data  and then b) with the universally available GIS data only.
+Third, we identify important variables for describing lake trophic state.
+Fourth, we explore associations between trophic state and cyanobacteria, a key taxa in harmful algal blooms, so that we may begin to understand how changes in trophic state may be linked to an important ecosystem disservice.
+Lastly, a key goal of this research is to provide full access to the data and code we used in this research.
+As such, this paper, the code, and the data used in the models is made available as an R package from https://github.com/USEPA/LakeTrophicModeling.
 
@@
 ##Data and Study Area
-We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].
-All datasets are national in scale and provide a unique snapshot view of the condition of lakes in the United States'.
+We utilize three primary sources of data for this study, the National Lakes Assessment (NLA), the National Land Cover Dataset (NLCD), and  lake morphometery modeled from the NHDPlus and National Elevation Data Set [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014].
+All datasets are national in scale and provide a unique snapshot view of the condition of lakes in the conterminous United States during the summer of 2007.
 
-The NLA data were collected during the summer of 2007 and the final data were released in 2009.
+The NLA data were collected during the summer of 2007 and the final data were released in 2009 [@usepa2009national].
 With consistent methods and metrics collected at 1056 locations across the conterminous United States (Figure \ref{fig:nlaMap}), the NLA provides a unique opportunity to examine broad scale patterns in lake productivity.
-The NLA collected data on biophysical measures of lake water quality and habitat.
-For this analysis we primarily examined the water quality measurements from the NLA [@usepa2009national].
-Adding to the monitoring data collected via the NLA, we use the 2006 NLCD data to examine the possible landscape-level drivers of trophic status in lakes.
-The NLCD is a nationally collected land use land cover dataset that also provides estimates of impervious surface.
-We collected total land use land cover and total percent impervious surface within a 3 kilometer buffer surrounding the lake to examine larger landscape-level effect [@homer2004development;@xian2009updating].
-We also used various measures of lake morphometry (i.e.
-depth, volume, fetch, etc.) as they are important in understanding lake productivity, yet many of these data are difficult to obtain for large numbers of lakes over broad regions.
-To add this information we modeled lake morphometry [@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@NLMDInPrep].
-Lastly, to explore associations between trophic state and cyanobacteria, we used estimates of cyanobacterial biovoulme caluclated by Beaulieu *et al.* [-@beaulieu2013nutrients].
-Cyanobacteria biovolumes are a truer measure of cyanobacteria dominance than abundance as there is great variability in the size within and between species.
-We have consolidated the taxa level estimates from Beaulieu *et al.* [-@beaulieu2013nutrients] and  summed that information on a per-lake basis.
+The NLA collected data on biophysical measures of lake water quality and habitat as well as an assessment of the phytoplankton community.
+For this analysis we examined the water quality measurements and total cyanobacteria abundance from the National Lakes Assessment [@usepa2009national].
+
+Adding to the monitoring data collected via the NLA, we use the 2006 NLCD data to examine landscape-level drivers of trophic status in lakes.
+The NLCD is a national land use/land cover dataset that also provides estimates of impervious surface.
+We calculated total proportion of each NLCD land use land cover class and total percent impervious surface within a 3 kilometer buffer surrounding each lake [@homer2004development;@xian2009updating].
+A three kilometer buffer was selected as an intermediate measure of the adjacent neighborhood; the three kilometer buffer size is greater than the immediate parcel but smaller than regional  and whole-basin measures.
+
+To account for unique aspects of each lake and characterize lake productivity, we also used measures of lake morphometry (i.e.
+depth, volume, fetch, etc.).
+As these data are difficult to obtain for large numbers of lakes over broad regions, we used modeled estimates of lake morphometry [@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014].
+These included: surface area, shoreline length, Shoreline Development, Maximum Depth, Mean Depth, Lake Volume, Maximum Lake Length, Mean Lake Width, Maximum Lake Width, and Fetch.
 
@@
 Random forest is a machine learning algorithm that aggregates numerous decision trees in order to obtain a consensus prediction of the response categories [@breiman2001random].
-Bootstrapped sample data is recursively partitioned according to a given random subset of predictor variables and completely grown without pruning.
-With each new tree, both the sample data and predictor variable subset is randomly selected.
+Bootstrapped sample data are recursively partitioned according to a given random subset of predictor variables and a predetermined number of decision trees  are developed.
+With each new tree, the sample data subset is randomly selected and with each new split, the subset of predictor variables are randomly selected.
+A detailed discussion of the benefits of a random forest approach is beyond the scope of this paper.
+To find out more see Breiman [-@breiman2001random] and Cutler et al.
+[-@cutler2007random].
 
-While random forests are able to handle numerous correlated variables without a decrease in prediction accuracy, unusually large numbers of related variables can reduce accuracy and increase the chances of over-fitting the model.
-This is a problem often faced in gene selection and in that field, a variable selection method based on random forest has been succesfully applied [@diaz2006gene].
-We use varselRF in R to initially examine the importance of the water quality and GIS derived variables and select a subset, the reduced model, to then pass to random forest[@diaz-uriarte2010varSelRF].
-
-Using R's randomForest package, we pass the reduced models selected with varSelRF and calculate confusion matrices, overall accuracy and kappa coeffecient [@liaw2002randomForest].
-From the reduced model random forests we collect a consensus prediction and calculate a confusion matrix and summary stats.
+Random forests are able to handle numerous correlated variables without a decrease in prediction accuracy; however, one possible downfall to this approach is that the resulting model may be difficult to interpret.
+This is a problem often faced in gene selection and in that field, a variable selection method based on random forest has been successfully applied and implemented in the R Language as the `varSelRF` package [@diaz2006gene].
+With this method, a minimum set of variables that maximizes model accuracy is provided.
+This allows us to start with a full suite of predictor variables from which to select a minimum, easier to interpret set of variables.
+One issue with the approach in `varSelRF` is that because of the randomization inherent in random forests it is possible to get variation in the minimum selected set of variables.
+To account for this we repeated `varSelRF` 100 times.
+In our case, repeating the procedure 100 times quickly converged on a set of all possible important variables.
 
 ##Model Details
-Using a combination of the `varSelRF` and `randomForest` we ran models for six combinations of variables and trophic state classifications.
-These combinations included different combinations of the Chlorphyll *a* trophic states (Table \ref{tab:trophicStateTable}) along with all variables and the GIS only variables (i.e.
-no *in situ* infromation).
+Using both the `varSelRF` and `randomForest` R packages we ran models for six sets of variables and trophic state classifications [@diaz-uriarte2010varSelRF; @liaw2002randomForest].
+These included three different combinations of the chlorophyll *a* trophic states as the dependent variables and using all variables (*in situ* and GIS variables) or the GIS only variables (i.e.
+no *in situ* information) as the independent variables in the random forest.
+A listing of all considered variables is in Appendix 1.
+Trophic state was defined using the NLA chlorophyll *a* trophic state cut offs and the three combinations of trophic state were used to highlight the possible error caused by misclassification of adjacent classes, such as mesotrophic and eutrophic (Table \ref{tab:trophicStateTable}).
+Our separation of predictors was chosen so that we could highlight the additional predictive performance provided by adding the *in situ* water quality variables on top of the GIS only variables.
+Lastly, we used only complete cases (i.e.
+missing data were removed) so the total number of observations varied between models.
+
 The six model combinations were:
 
-1. Chlorophyll *a* trophic state - 4 class = All variables (*in situ* water quality, lake morphometry, and landscape)
-2. Chlorophyll *a* trophic state - 3 class = All variables (*in situ* water quality, lake morphometry, and landscape)
-3. Chlorophyll *a* trophic state - 2 class = All variables (*in situ* water quality, lake morphometry, and landscape)
-4. Chlorophyll *a* trophic state - 4 class = All variables (lake morphometry, and landscape)
-5. Chlorophyll *a* trophic state - 3 class = All variables (lake morphometry, and landscape)
-6. Chlorophyll *a* trophic state - 2 class = All variables (lake morphometry, and landscape)
+- **Model 1:** Chlorophyll *a* trophic state (4 class) = All variables (*in situ* water quality, lake morphometry, and landscape)
+- **Model 2:** Chlorophyll *a* trophic state (3 class) = All variables (*in situ* water quality, lake morphometry, and landscape)
+- **Model 3:** Chlorophyll *a* trophic state (2 class) = All variables (*in situ* water quality, lake morphometry, and landscape)
+- **Model 4:** Chlorophyll *a* trophic state (4 class) = GIS Only variables (lake morphometry, and landscape)
+- **Model 5:** Chlorophyll *a* trophic state (3 class) = GIS Only variables (lake morphometry, and landscape)
+- **Model 6:** Chlorophyll *a* trophic state (2 class) = GIS Only variables (lake morphometry, and landscape)
 
-#Results and Discussion
+Our modelling work flow was as follows:
 
-##Model 1: 4 Trophic States ~ All Variables
-The selected variables that made up Model 1 were Potassium, Nitrogen:Phosphorus, Total Nitrogen, Total Phosphorus, Total Organic Carbon, Turbidity, Ecoregion, Organic Ions, Dissolved Organic Carbon, and Maximum Lake Depth (Table \ref{tab:VarSel_Model1}).
-Total accuracy for Model 1 is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,3)` (Table \ref{tab:Confusion_Model1}).
+1.
+Use `iterVarSelRF` in the `LakeTrophicModelling` R package to identify a minimal set of variables that maximize accuracy of the random forest algorithm [@diaz-uriarte2010varSelRF; @LakeTrophicModelling].
+This subset of variables, the reduced model, is calculated for each of our 6 models.
+2. Using R's `randomForest` package, we pass the reduced models selected with `iterVarSelRF` and assess model performance [@liaw2002randomForest].
 
-Lastly, tubidity, total phosphorus, total nitrogen, and total organic carbon were the most important predictors of the 4 classes of trophic state (Figure \ref{fig:Importance_Model1}).
+##Measures of Model Performance and Variable Importance
 
-##Model 2: 3 Trophic States ~ All Variables
+We assessed the performance of the random forest models by comparing the total prediction accuracy and the kappa coefficient of the final confusion matrix.
+For each of the models, the final predictions were compared to the original data via a confusion matrix.
+A confusion matrix shows agreement and disagreement with predicted values forming the columns of the matrix and observed values, the rows.
+The total accuracy (i.e.
+percent correctly predicted) was calculated.
+Since some agreement can be expected by chance alone, it is also useful to take this type of error into account.
+For this we calculated the kappa coefficient from the confusion matrix for each model as well [@cohen1960coefficient; @hubert1985comparing].
+The kappa coefficient can range from -1 to 1 with 0 equalling the agreement expected by chance alone.
+Values greater than 0 represent agreement greater than would be expected by chance, with values greater than 0.61 considered "substantial" agreement  [@landis1977measurement].
+Negative values are rare and would indicate no agreement between the predicted and observed values.
+Additonally, random forest builds each tree on bootstrapped, random subsets of the original data, thus, a separate independent validation dataset is not required and random forest error estimates are expected to be unbiased [@breiman2001random].
 
-Total accuracy for Model 2 is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts3_rf$confusion[,1:3])$kappa,3)`.
+Lastly, the random forest algorithm explicitly measures variable importance as mean decrease in Gini.
+The Gini Index is a measure of how well the data are classified into homogeneous groups.
+For every node, the splitting variables are permuted and the change in actual Gini and permuted Gini is recorded.
+The mean decrease Gini is a summed and standardized value for each variable [@breiman2001random].
+Higher values of mean decrease Gini suggest a higher importance for that variable.
 
-##Model 3: 2 Trophic States ~ All Variables
-Total accuracy for Model 3 is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,3)`.
+#Results
 
-##Model 4: 4 Trophic States ~ GIS Only Variables
-Total accuracy for Model 4 is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,3)`.
+Our complete dataset includes `r dim(ltmData)[1]` lakes; however `r sum(is.na(ltmData$TS_CHLA_4))` lakes did not have chlorophyll *a* data.
+Thus, the base dataset for our modelling was conducted on data for `r dim(ltmData[!is.na(ltmData$TS_CHLA_4),])[1]` lakes.
+As chlorophyll *a* is used to create the trophic state classifications, it was necessary to remove these data because no chlorophyll *a* trophic state could be determined for these lakes.
+The lakes were well distributed both across the four trophic state categories (Table \ref{tab:trophicStateTable}) and spatially throughout the United States (Figure \ref{fig:nlaMap}).
 
-##Model 5: 3 Trophic States ~ GIS Only Variables
-Total accuracy for Model 5 is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,3)`.
+##Models
+Accuracy for the models built with all predictors ranged from `r round(min(acc_results[1:3,2]),2)*100`% to `r round(max(acc_results[1:3,2]),2)*100`% and the kappa coefficient had a minimum value of `r round(min(acc_results[1:3,3]),2)` and maximum of `r round(max(acc_results[1:3,3]),2)`.
+The GIS only models had a total accuracy between `r round(min(acc_results[4:6,2]),2)*100`% and `r round(max(acc_results[4:6,2]),2)*100`% and kappa coefficient between `r round(min(acc_results[4:6,3]),2)` and `r round(max(acc_results[4:6,3]),2)`.
+The importance of variables for the models including the *in situ* data were fairly stable while there was considerably more variation in variable importance for the three different GIS only models.
+Additionally, as each model contained a different subset of variables and since we removed observations with missing data, the total number of lakes varied across the models.
+Details for each model are discussed below.
 
-##Model 6: 2 Trophic States ~ GIS Only Variables
-Total accuracy for Model 6 `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,3)`% and the Cohen's Kappa is `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,3)`.
+*Model 1: 4 Trophic States ~ All Variables*
+
+Total number of lakes for Model 1 was `r sum(complete.cases(ltmData[all_ts4_vars],ltmData$TS_CHLA_4))` and the reduced model for Model 1 included turbidity, total phosphorus, total nitrogen, nitrogen:phosphorus ratio, total organic carbon, potassium, ecoregion, maximum depth, organic ions, and dissolved organic carbon.
+Of these, turbidity, total phosphorus, and total nitrogen were the three most important predictors of the four classes of trophic state (Figure \ref{fig:Importance_Model1}).
+Total accuracy for Model 1 was `r round(classAgreement(all_ts4_rf$confusion[,1:4])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(all_ts4_rf$confusion[,1:4])$kappa,2)` (Table \ref{tab:Confusion_Model1}).
+
+*Model 2: 3 Trophic States ~ All Variables*
+
+For Model 2 there were `r sum(complete.cases(ltmData[all_ts3_vars],ltmData$TS_CHLA_3))` observations and the reduced model included turbidity, total phosphorus, total nitrogen, total organic carbon, nitrogen:phosphorus, longitude, pH, estimated organic anions, elevation, maximum depth, dissolved organic carbon, potassium, latitude, ecoregion, chloride, ammonium and percent cropland (Figure \ref{fig:Importance_Model2}).
+The top predictors for 3 trophic state classes were again turbidity, total phosphorus, and total nitrogen (Figure \ref{fig:Importance_Model2}).
+Model 2 accuracy was `r round(classAgreement(all_ts3_rf$confusion[,1:4])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(all_ts3_rf$confusion[,1:4])$kappa,2)` (Table \ref{tab:Confusion_Model2}).
+
+*Model 3: 2 Trophic States ~ All Variables*
+
+Model 3 was built with `r sum(complete.cases(ltmData[all_ts2_vars],ltmData$TS_CHLA_2))` lakes and the reduced model was similar to Model 1 and Model 2.
+This model included turbidity, total phosphorus, total nitrogen, nitrogen:phosphorus, potassium, ecoregion, elevation, total organic carbon, growing degree days, longitude, sodium, maximum depth, estimated organic anions, latitude, and dissolved organic carbon (Figure \ref{fig:Importance_Model3}).
+The top three predictors were the same as Models 1 and 2.
+(Figure \ref{fig:Importance_Model3}).
+Total accuracy for Model 3 was `r round(classAgreement(all_ts2_rf$confusion[,1:2])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(all_ts2_rf$confusion[,1:2])$kappa,2)` (Table \ref{tab:Confusion_Model3}).
+
+*Model 4: 4 Trophic States ~ GIS Only Variables*
+
+Model 4 had a sample size of `r sum(complete.cases(ltmData[gis_ts4_vars],ltmData$TS_CHLA_4))`.
+The selected variables for the Model 4 were longitude, latitude, elevation, estimated mean lake depth, percent evergreen forest, estimated maximum lake depth, percent cropland, and ecoregion (Figure \ref{fig:Importance_Model4}).
+The three most important variables were longitude, latitude, and elevation (Figure \ref{fig:Importance_Model4}); however, all variables except ecoregion were almost equally important.
+Total accuracy for Model 4 was `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(gis_ts4_rf$confusion[,1:4])$kappa,2)` (Table \ref{tab:Confusion_Model4}).
+
+*Model 5: 3 Trophic States ~ GIS Only Variables*
+
+A total of `r sum(complete.cases(ltmData[gis_ts3_vars],ltmData$TS_CHLA_3))` lakes were used for model 5 which included estimated mean lake depth, percent cropland, longitude, latitude, percent evergreen forest, elevation, estimated maximum lake depth, estimated lake volume, percent deciduous forest, percent developed open space, ecoregion, percent woody wetland, and percent shrub/scrub (Figure \ref{fig:Importance_Model5}).
+The most important variables for model 5 were estimated mean lake depth, percent cropland, and longitude.
+(Figure \ref{fig:Importance_Model5}).
+Total accuracy for Model 5 was `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(gis_ts3_rf$confusion[,1:3])$kappa,2)` (Table \ref{tab:Confusion_Model5}).
+
+*Model 6: 2 Trophic States ~ GIS Only Variables*
+
+Model 6 was built using  `r sum(complete.cases(ltmData[gis_ts2_vars],ltmData$TS_CHLA_2))` total observations.
+The variable selection process for Model 6 produced a reduced model with ecoregion, growing degree days, percent evergreen forest, percent cropland, elevation, estimated mean lake depth, longitude, latitude, watershed area, estimated maximum lake depth, percent developed open space, percent deciduous forest, and estimated lake volume (Figure \ref{fig:Importance_Model6}).
+The three most important variables were ecoregion, growing degree days, and percent evergreen.
+(Figure \ref{fig:Importance_Model6}).
+Ecoregion had the highest importance with a slow decline in importance for the remainder of the variables.
+Total accuracy for Model 6 was `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$diag,2)*100`% and the Cohen's Kappa was `r round(classAgreement(gis_ts2_rf$confusion[,1:2])$kappa,2)` (Table \ref{tab:Confusion_Model6}).
+
+#Discussion
+
+##Trophic State Probabilities
+One of the powerful features of random forests is the ability to aggregate a very large number of competing models or trees.
+Each tree provides an independent prediction or vote for a possible outcome.
+In the context of our trophic state models, we have 10,000 votes for each lake.
+These values may be interpreted as the probability that a lake is in a given trophic state.
+For instance, for a single lake (National Lake Assessment ID = NLA06608-0005), the vote probabilities for Model 1 were `r 100*round(lakeVotes[lakeVotes$COMID==23491387,][,4],2)`% for oligotrophic, `r 100*round(lakeVotes[lakeVotes$COMID==23491387,][,5],2)`% for mesotrophic, `r 100*round(lakeVotes[lakeVotes$COMID==23491387,][,6],2)`% for eutrophic, and `r 100*round(lakeVotes[lakeVotes$COMID==23491387,][,7],2)`% for hypereutrophic.
+This suggests little uncertainty in the predicted oligotrophic state.
+
+Further, the maximum probability for each lake can be used as a measure of how certain the random forest model was of the prediction.
+We would expect higher total accuracy for lakes that had more certain predictions.
+Some lakes may have many votes for a single trophic state and few votes for other trophic states and these would thus have a large maximum probability and the random forest predicitions would be more certain.
+Alternatively, the 10,000 votes could have been spread more equally across the trophic state classes for a lake and that lake would have a small maximum probability and the final predictions would be less certain.
+This should be evident by looking at the total classification accuracy of lakes given their maximum probability is above a certain point.
+To test this we can examine the accuracy of trophic state predictions across the full range of trophic state probabilities, similar to an approach outlined by Paul and MacDonald [-@paul2005development] and implemented by Hollister et al.
+[-@hollister2008cprob].
+We utilize this approach and examine the change in total accuracy as a function of the maximum probability for each lake.
+As expected, lakes with higher maximum vote probabilities were more accurately predicted (Figure \ref{fig:condProbFig}).
+The increasing trend suggests that even for models with lower overall accuracy there can also be a large number of individual cases that are predicted with high accuracy.
+
+##Variable Selection and Importance
+There was a great deal of agreement on the important variables for each set of models.
+In line with past predictive modeling of cyanobacteria abundance and not surprisingly, the *in situ* models consistently select the water quality variables (turbidity, total nitrogen, total phosphorus, and N:P ratios) as important variables [@downing2001predicting].
+While there is variation in the response of cyanobacteria to changes in relative nutrient concentrations, the general pattern suggests that limiting nutrients have considerable impact once amounts increase beyond expected levels.
+
+The mechanistic role of turbidity on lake trophic state is more complex.
+Light availability in turbid waters is lower than in clear waters.
+This would suggest a negative relationship between turbidity and chlorophyll *a*.
+Second, chlorophyll *a* can also be a component of turbidity and lakes with higher chlorophyll *a* concentrations will also be more turbid.
+Last, chlorophyll *a* is not the only component of of turbidity and turbid waters can be caused by, for example, increased sediment loads or tannin.
+This would be a cause for concern with linear models; however, linearity is not an assumption of tree-based modelling approaches such as random forest.
+
+Our models with the GIS-only variables captured the large scale spatial pattern of the trophic status gradient of lakes across the United States.
+We reliably saw latitude and longitude and ecoregion selected as important variables.
+It is also possible that other variables selected as important are also capturing a portion of this trend.
+For instance, elevation and growing degree days both have obvious spatial components, but may also be accounting for variation in temperature.
+
+The land use/land cover variables were also important in describing trophic state patterns.
+Like elevation and growing degree days, broad scale spatial patterns are inherent in the data.
+For instance, the relative continental position of mountains in the United States is the spatial inverse of the distribution of agricultural lands.
+However, it is known that forests are positively associated with lower nutrient loads where as agricultural land shows a negative association.
+These more local scale relationships with land use/land cover  likely provide additional predictive power to the information in the broader scale data.
+
+Lastly, morphometry (e.g.
+depth and volume) also proved to be important in the prediction of lake trophic state.
+As morphometry shows little to no broad scale spatial pattern and is unique to a given lake, these data are likely illuminating the local, lake scale drivers of trophic state.
+As only depth and volume were selected, this likely shows the importance of in-lake nutrient processing and residence time.
 
 ##Associating Trophic State and Cyanobacteria
-Arm waving goes here.
+
+Cyanobacteria biomass should be closely associated with trophic state as cyanobacteria contribute to the chlorophyll concentration in a lake.
+If these associations are strong enough we may be able to expand models such as those reported here to also predict probability of cyanobacteria blooms.
+To test if trophic state can be used to differentiate cyanobacteria abundance, we examine distribution of cyanobacteria abundance for each trophic state and also explored linear associations between chlorohyll *a* and cyanobacteria abundance.
+
+The distribution of cyanobacteria abundance showed separation between all of the trophic state classifications (Figures \ref{fig:ts_4_cyano_cdf}, \ref{fig:ts_3_cyano_cdf}, and \ref{fig:ts_2_cyano_cdf}) and there was a significant linear relationship (r^2^=`r round(summary(lm(log10(ltmData$CHLA)~log10(ltmData$cyanoCellsPerML)))$r.squared,2)`) between chlorophyll *a* and cyanobacteria abundance (Figure \ref{fig:scatterplot}).
+Furthermore, Yuan et al.
+[-@yuan2014managing] used the 2007 NLA to demonstrate that total nitrogen and chlorophyll *a* concentrations were good predictors of World Health Organization microcystin (a toxin produced by some cyanobacteria) criteria exceedences.
+These results suggest that trophic state is indeed an acceptable proxy for cyanobacteria abundance and that in lakes with higher trophic state it is also reasonable to expect higher cyanobacteria.
+
+#Conclusions
+
+Our research goals were to explore the utility of a widely used data mining algorithm, random forests, in the modelling of lake trophic state.
+Further, we hoped to examine the utility of these models when built with only ubiquitous GIS data, which allows estimation of trophic state for all lakes in the United States.
+We were able to successfully predict a variety of trophic state classes.
+With the GIS only models our total accuracy ranged from `r 100*round(min(acc_results$total_accuracy[4:6]),2)`% to `r 100*round(max(acc_results$total_accuracy[4:6]),2)`%, and with the full suite of data our model accuracy had a minimum accuracy of `r 100*round(min(acc_results$total_accuracy[1:3]),2)`% and maximum accuracy of `r 100*round(max(acc_results$total_accuracy[1:3]),2)`%.
+
+While some of the models (i.e.
+Model 4) showed relatively low prediction accuracies, another feature of the random forest, votes, can provide additional information.
+In addition to providing a single estimate of trophic state for each lake, our models also indicated the probability that a lake was classified in any of the categories.
+These probabilities may be mapped directly to show the uncertainty of a given predicted class.
+Furthermore, as the certainty of prediction increases, so does overall trophic state classification accuracy (Figure \ref{fig:condProbFig}).
+These results suggest that our models will provide reasonable estimates of trophic state across the United States.
+
+There was great deal of agreement on the important variables for each set of models.
+For the combined *in situ* and GIS models, the *in situ* water quality variables drove the predictions.
+This is expected.
+For the GIS only models,  the results were more nuanced with three broad categories routinely being selected as important: broad scale spatial patterns in trophic state, land use/land cover controls of trophic state, and local, lake-scale control driven by lake morphometry.
+Lastly, associations between trophic state and cyanobacteria showed that, at the broad scale of the 2007 NLA, there is a linear relationship between chlorophyll *a* and cyanobacteria abundance and that using trophic state as a proxy for cyanobacteria has potential.
+
+These broad categories and the association between trophic state and total cyanobacteria abundance raise three important considerations related to managing eutrophication.
+First, the broad scale patterning suggests regional trends.
+This is important because it suggests that efforts to monitor, model and manage eutrophication and cyanobacteria should be undertaken at both national and regional levels.
+Second, while direct control of water quality in lakes would have a large impact, the land use/land cover drivers (i.e.
+non-point sources) of water quality are also important, and better management of the spatial distribution of important classes such as forest and agriculture can provide some level of control on trophic state and amount of cyanobacteria present.
+Third, in-lake processes (i.e.
+residence time, nutrient cycling, etc.) are, as expected, very important and need to be part of any management strategy.
+Building on these efforts through updated models,  direct prediction of cyanobacteria, and additional information on the regional differences will help us get a better handle on the broad scale dynamics of productivity in lakes and the potential risk to human health from cyanobacteria blooms.
+
+#Acknowledgements
+We would like to thank Farnaz Nojavan, Nathan Schmucker, John Kiddon, Joe LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this paper.
+This paper has not been subjected to Agency review.
+Therefore, it does not necessary reflect the views of the Agency.
+Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
+This contribution is identified by the tracking number ORD-011075 of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
 
@@
 
-```{r figSetup,echo=FALSE,eval=TRUE,cache=TRUE}
+```{r figSetup,echo=FALSE,eval=TRUE,cache=FALSE}
 #create list of all selected variables and assign a color
@@
 all_movies_palette<-unique(all_movies_palette)
-zissou<-c(wes.palette(5,"Zissou"),wes.palette(5,"Zissou2"))
+zissou<-c(wes.palette(5,"Zissou"),wes.palette(5,"Rushmore"))
 col_lu<-data.frame(variables=all_vars,hexcode=sample(all_movies_palette,length(all_vars)))
@@
 
-```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}",cache=TRUE}
+```{r nlaMap, echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations\\label{fig:nlaMap}",cache=FALSE}
 state<-map_data('state')
@@
 lakes_dd<-data.frame(coordinates(lakes_dd))
-mycolor<-wes.palette(5,"Zissou2")
+mycolor<-wes.palette(5,"Rushmore")
 mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
@@
 
-```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1\\label{fig:Importance_Model1}",cache=TRUE}
+```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model1}",cache=FALSE}
 all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
-importancePlot(all_ts4_rf,'gini',size=10,aes(colour=all_ts4_color))
+importancePlot(all_ts4_rf,sumTable(all_ts4[1:100]), data_def=data_def,'gini',size=5,aes(colour=all_ts4_color))
 ```
@@
 
-```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2",cache=TRUE}
+```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model2}",cache=FALSE}
 all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
-importancePlot(all_ts3_rf,'gini',size=10,aes(colour=all_ts3_color))
+importancePlot(all_ts3_rf,sumTable(all_ts3[1:100]), data_def=data_def,'gini',size=5,aes(colour=all_ts3_color))
 ```
@@
 
-```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3",cache=TRUE}
+```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model3}",cache=FALSE}
 all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
-importancePlot(all_ts2_rf,'gini',size=10,aes(colour=all_ts2_color))
+importancePlot(all_ts2_rf,sumTable(all_ts2[1:100]), data_def=data_def,'gini',size=5,aes(colour=all_ts2_color))
 ```
@@
 
-```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 4",cache=TRUE}
+```{r Importance_Model4,results="asis",echo=FALSE, fig.cap="Importance plot for Model 4, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model4}",cache=FALSE}
 gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
-importancePlot(gis_ts4_rf,'gini',size=10,aes(colour=gis_ts4_color))
+importancePlot(gis_ts4_rf,sumTable(gis_ts4[1:100]), data_def=data_def,'gini',size=5,aes(colour=gis_ts4_color))
 ```
@@
 
-```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5",cache=TRUE}
+```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model5}",cache=FALSE}
 gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
-importancePlot(gis_ts3_rf,'gini',size=10,aes(colour=gis_ts3_color))
+importancePlot(gis_ts3_rf,sumTable(gis_ts3[1:100]), data_def=data_def,'gini',size=5,aes(colour=gis_ts3_color))
 ```
@@
 
-```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6",cache=TRUE}
+```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6, shows mean descrease gini and proportion of times a variable was selected for inclusion in the model.
+Higher values of mean decrease gini indicate higher importance.
+\\label{fig:Importance_Model6}",cache=FALSE}
 gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
-importancePlot(gis_ts2_rf,'gini',size=10,aes(colour=gis_ts2_color))
+importancePlot(gis_ts2_rf,sumTable(gis_ts2[1:100]), data_def=data_def,'gini',size=5,aes(colour=gis_ts2_color))
 ```
@@
 
-```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes",cache=TRUE}
-plotCdf(ltmData$TS_CHLA_4,ltmData$sumBioV+1,cdf_colors=zissou[6:9],
-        y='Percent',x='Log10(Cyanobacteria Biovolume)',
-        title=expression(paste('CDF for Chlorophyll ', 
-                                italic("a"),' Trophic States (4 Categories)')),
+```{r condProbFig,echo=FALSE, fig.width=7, fig.cap="Comparison of certainity of trophic state prediction and total accuracy\\label{fig:condProbFig}",cache=FALSE}
+condAccuracy(gis_ts4_rf,xImpair=0,R=1,xlab="Maximum Vote Probability")
+```
+
+\newpage
+
+```{r ts_4_cyano_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria abundance for 4 trophic state classes\\label{fig:ts_4_cyano_cdf}",cache=FALSE}
+plotCdf(ltmData$TS_CHLA_4,ltmData$cyanoCellsPerML+1,cdf_colors=zissou[6:9],
+        y='Percent',x='Log10(Cyanobacteria Abundance (cells/ml) + 1)',
         color="Trophic State\nCategories")
@@
 
-```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes",cache=TRUE }
-plotCdf(ltmData$TS_CHLA_3,ltmData$sumBioV+1,cdf_colors=zissou[6:8],
-       y='Percent',x='Log10(Cyanobacteria Biovolume)', 
-       title=expression(paste('CDF for Chlorophyll ', 
-                              italic("a"),' Trophic States (3 Categories)')),
+```{r ts_3_cyano_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria abundance for 3 trophic state classes\\label{fig:ts_3_cyano_cdf}",cache=FALSE }
+plotCdf(ltmData$TS_CHLA_3,ltmData$cyanoCellsPerML+1,cdf_colors=zissou[6:8],
+       y='Percent',x='Log10(Cyanobacteria Abundance (cells/ml) + 1)', 
        color="Trophic State\nCategories")
@@
 
-```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes",cache=TRUE}
-plotCdf(ltmData$TS_CHLA_2,ltmData$sumBioV+1,cdf_colors=zissou[6:7],
-       y='Percent',x='Log10(Cyanobacteria Biovolume)',
-       title=expression(paste('CDF for Chlorophyll ', 
-                              italic("a"),' Trophic States (2 Categories)')),
+```{r ts_2_cyano_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria abundance for 2 trophic state classes\\label{fig:ts_2_cyano_cdf}",cache=FALSE}
+plotCdf(ltmData$TS_CHLA_2,ltmData$cyanoCellsPerML+1,cdf_colors=zissou[6:7],
+       y='Percent',x='Log10(Cyanobacteria Abundance (cells/ml) + 1)',
        color="Trophic State\nCategories")
@@
 
-```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot",cache=TRUE}
-scp_df<-data.frame(chla=ltmData[["CHLA"]],biovp1=ltmData[["sumBioV"]]+1)
-scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
-             title=expression(paste("Chlorophyll ", 
-                                    italic("a")," and Cyanobacteria Relationship")),
-             x=expression(paste('Log10(Chl ', italic("a"),')')),
-             y="Log10(Cyanobaterial Biovolumes + 1)")
+```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria abundance scatterplot\\label{fig:scatterplot}",cache=FALSE}
+scp_df<-data.frame(chla=ltmData[["CHLA"]],abundp1=ltmData[["cyanoCellsPerML"]]+1)
+scatterPlot(scp_df,xvar="chla",yvar="abundp1",zissou[8],zissou[7],zissou[6],
+             x=expression(paste('Log10(Chl ', italic("a"),'(',mu,'g/L))')),
+             y="Log10(Cyanobaterial Abundance (cells/ml) + 1)")
 ```
@@
 
-```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs\\label{tab:trophicStateTable}",cache=TRUE}
-ts_4<-c("oligo","meso","eu","hyper")
-ts_3<-c("oligo","meso/eu","meso/eu","hyper")
-ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
+```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorophyll a based trophic state cut-offs with total number of possible observations.\\label{tab:trophicStateTable}",cache=FALSE}
+ts_4<-c("oligotrophic","mesotrophic","eutrophic","hypereutrophic")
+ts_3<-c("oligotrophic","mesotrophic/eutrophic","mesotrophic/eutrophic","hypereutrophic")
+ts_2<-c("oligotrophic/mesotrophic","oligotrophic/mesotrophic","eutrophic/hypereutrophic","eutrophic/hypereutrophic")
 co<-c("<= 0.2",">2-7",">7-30",">30")
 xdf<-data.frame(ts_4,ts_3,ts_2,co)
-names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
+names(xdf)<-c("Trophic State (4 class)","Trophic State (3 class)","Trophic State (2 class)",
+              "Concentration Cut-off")
 kable(xdf)
@@
 
-```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1\\label{tab:VarSel_Model1}",cache=TRUE}
-kable(sumTable(all_ts4[1:100]),row.names=F)
+```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in 'Class Accuracy' column.
+\\label{tab:Confusion_Model1}",cache=FALSE}
+cm1<-round(all_ts4_rf$confusion,2)
+cm1[,5]<-paste0((1-cm1[,5])*100,"%")
+dimnames(cm1)[[2]][5]<-'Class Accuracy'
+kable(cm1,row.names=T)
 ```
@@
 
-```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1\\label{tab:Confusion_Model1}",cache=TRUE}
-kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
+```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in 'Class Accuracy' column.
+\\label{tab:Confusion_Model2}",cache=FALSE}
+cm2<-round(all_ts3_rf$confusion,2)
+cm2[,4]<-paste0((1-cm2[,4])*100,"%")
+dimnames(cm2)[[2]][4]<-'Class Accuracy'
+kable(cm2,row.names=T)
 ```
@@
 
-```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2",cache=TRUE}
-kable(sumTable(all_ts3[1:100]),row.names=F)
+```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in 'Class Accuracy' column.
+\\label{tab:Confusion_Model3}",cache=FALSE}
+cm3<-round(all_ts2_rf$confusion,2)
+cm3[,3]<-paste0((1-cm3[,3])*100,"%")
+dimnames(cm3)[[2]][3]<-'Class Accuracy'
+kable(cm3,row.names=T)
 ```
@@
 
-```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2",cache=TRUE}
-kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
+```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in 'Class Accuracy' column.
+\\label{tab:Confusion_Model4}",cache=FALSE }
+cm4<-round(gis_ts4_rf$confusion,2)
+cm4[,5]<-paste0((1-cm4[,5])*100,"%")
+dimnames(cm4)[[2]][5]<-'Class Accuracy'
+kable(cm4,row.names=T)
 ```
@@
 
-```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3",cache=TRUE}
-kable(sumTable(all_ts2[1:100]),row.names=F)
+```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in 'Class Accuracy' column.
+\\label{tab:Confusion_Model5}",cache=FALSE}
+cm5<-round(gis_ts3_rf$confusion,2)
+cm5[,4]<-paste0((1-cm5[,4])*100,"%")
+dimnames(cm5)[[2]][4]<-'Class Accuracy'
+kable(cm5,row.names=T)
 ```
@@
 
-```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3",cache=TRUE}
-kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
+```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6.
+Columns show predicted values and rows show observed values.
+Agreement indicated on diagonal and accuracy for each trophic state indicated in Class Accuracy' column.
+\\label{tab:Confusion_Model6}",cache=FALSE}
+cm6<-round(gis_ts2_rf$confusion,2)
+cm6[,3]<-paste0((1-cm6[,3])*100,"%")
+dimnames(cm6)[[2]][3]<-'Class Accuracy'
+kable(cm6,row.names=T)
 ```
@@
 
-```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4",cache=TRUE }
-kable(sumTable(gis_ts4[1:100]),row.names=F)
-```
+#Appendix 1. Variable Definitions
 
-\newpage
-
-```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4",cache=TRUE }
-kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
-```
-
-\newpage
-
-```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5",cache=TRUE}
-kable(sumTable(gis_ts3[1:100]),row.names=F)
-```
-
-\newpage
-
-```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5",cache=TRUE}
-kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
-```
-
-\newpage
-
-```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6",cache=TRUE}
-kable(sumTable(gis_ts2[1:100]),row.names=F)
-```
-
-\newpage
-
-```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6",cache=TRUE}
-kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
+```{r data_def, results="asis",echo=FALSE,cache=FALSE}
+kable(data_def,row.names=FALSE)
 ```

### 2015-04-04 vignettes/submission/prepare_submission.R
working on remake to build submission

@@ -1,0 +1,3 @@
+copy<-function(){
+  file.copy("../manuscript.Rmd","manuscript.Rmd")
+}

### 2015-04-05 vignettes/submission/prepare_submission.R
first bit of remake working

@@ -1,3 +1,3 @@
 copy<-function(){
-  file.copy("../manuscript.Rmd","manuscript.Rmd")
+  file.copy("../manuscript.Rmd","manuscript.Rmd",overwrite=T)
 }

### 2015-04-06 R/condAccuracy.R
submission almost ready

@@ -22,15 +22,15 @@ condAccuracy<-function(rf,xlab="x",...){
   
   ggplot(cp,aes(x=cp.max_vote,y=cp.Raw.Data.Probability))+
     geom_point()+
-    theme(text = element_text(family="serif"),
+    theme(text = element_text(family="sans"),
           panel.background = element_blank(), #panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
-          plot.title  = element_text(family="serif",size=15,face="bold",vjust=1.1),
+          plot.title  = element_text(family="sans",size=15,face="bold",vjust=1.1),
           legend.position = "none", legend.key = element_rect(fill = 'white'),
-          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=15),
-          axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
-          axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
-          axis.text.x = element_text(family="serif",size = 11),
-          axis.text.y = element_text(family="serif",size = 11)) + 
+          legend.text = element_text(family="sans",size=15), legend.title = element_text(size=15),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="sans",size = 11),
+          axis.text.y = element_text(family="sans",size = 11)) + 
     labs(y="Conditional Total Accuracy",x=xlab)
 }

### 2015-04-06 R/importancePlot.R

@@ -54,7 +54,7 @@ importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"),
             panel.border = element_blank(), 
             axis.title.x = element_blank(),
             axis.text.x = element_blank(),
-            axis.text.y = element_text(family="serif",size = 11),
+            axis.text.y = element_text(family="sans",size = 11),
             axis.ticks.x = element_blank(),
             plot.margin = grid::unit(c(0.25,0,0.3,0.6),"inches")) + 
       ylab("") + 
@@ -67,8 +67,8 @@ importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"),
          theme(text = element_text(family="Times"),
                panel.background = element_blank(), panel.grid = element_blank(), 
                panel.border = element_rect(fill = NA), legend.position = "none", 
-               axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
-               axis.text.x = element_text(family="serif",size = 11),
+               axis.title.x = element_text(family="sans",face = "bold", vjust = -0.5, size = 12), 
+               axis.text.x = element_text(family="sans",size = 11),
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                plot.margin = grid::unit(c(0.25,0.1,0.1,0),"inches")) + 
@@ -82,8 +82,8 @@ importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"),
       theme(text = element_text(family="Times"),
             panel.background = element_blank(), panel.grid = element_blank(), 
             panel.border = element_rect(fill = NA), legend.position = "none", 
-            axis.title.x = element_text(family="serif",face = "bold", vjust = -0.5, size = 12), 
-            axis.text.x = element_text(family="serif",size = 11),
+            axis.title.x = element_text(family="sans",face = "bold", vjust = -0.5, size = 12), 
+            axis.text.x = element_text(family="sans",size = 11),
             axis.text.y = element_blank(),
             axis.ticks = element_blank(),
             plot.margin = grid::unit(c(0.25,0.1,0.1,0),"inches")) + 

### 2015-04-06 R/plotCdf.R

@@ -38,13 +38,13 @@ plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
           theme(text = element_text(family="Times"),
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 
-                plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
+                plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
                 legend.position = c(0.15,0.75), legend.key = element_rect(fill = 'white'),
-                legend.text = element_text(family="serif",size=15), legend.title = element_text(family="serif",size=11),
-                axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
-                axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
-                axis.text.x = element_text(family="serif",size = 11),
-                axis.text.y = element_text(family="serif",size = 11)) + 
+                legend.text = element_text(family="sans",size=15), legend.title = element_text(family="sans",size=11),
+                axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+                axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+                axis.text.x = element_text(family="sans",size = 11),
+                axis.text.y = element_text(family="sans",size = 11)) + 
           geom_hline(linetype = 3, size = 1, colour = "gray", 
                      yintercept = c(0,0.5,1)) +
           labs(...)

### 2015-04-06 R/scatterPlot.R

@@ -35,16 +35,16 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
     #            colour=ln_col,size=2) + 
     #annotate("text", x = 0.25, y = 100000000, label = xlm_r) +
     #annotate("text", x = 0.285, y = 20000000, label = xlm_p) +
-    theme(text = element_text(family="serif"),
+    theme(text = element_text(family="sans"),
           panel.background = element_blank(), panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
-          plot.title  = element_text(family="serif",size=12,face="bold",vjust=1.1),
+          plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
           legend.position = "none", legend.key = element_rect(fill = 'white'),
-          legend.text = element_text(family="serif",size=15), legend.title = element_text(size=11),
-          axis.title.x = element_text(family="serif",vjust = -0.5, size = 12),
-          axis.title.y = element_text(family="serif",vjust = 1.5, size = 12),
-          axis.text.x = element_text(family="serif",size = 11),
-          axis.text.y = element_text(family="serif",size = 11)) + 
+          legend.text = element_text(family="sans",size=15), legend.title = element_text(size=11),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="sans",size = 11),
+          axis.text.y = element_text(family="sans",size = 11)) + 
     labs(...)
     
   return(x)

### 2015-04-06 vignettes/submission/manuscript.R

@@ -21,10 +21,10 @@ library("e1071")
 library("dplyr")
 library("tidyr")
 library("condprob2")
-opts_chunk$set(dev = 'png', fig.width=7.5)
+opts_chunk$set(dev = 'jpeg', dpi=300, fig.width=7.5, knitr.table.format="html")
 data(LakeTrophicModelling)
 
-
+#
 #Checks for existing cache (from another project)
 #Else if Repeats running chunk outside of knitr
 if(file.exists("prior_cache")){
@@ -33,6 +33,10 @@ if(file.exists("prior_cache")){
   }
 } else if(file.exists("vignettes/prior_cache")){
   for(i in gsub(".rdb","",list.files("vignettes/prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+} else if(file.exists("../prior_cache")){
+  for(i in gsub(".rdb","",list.files("../prior_cache",".rdb",full.names=T))){
     lazyLoad(i)
   }
 }
@@ -157,7 +161,7 @@ var_importance<-rbind(varImportance(all_ts4_rf,"Model 1"),
                            varImportance(gis_ts4_rf,"Model 4"),
                            varImportance(gis_ts3_rf,"Model 5"),
                            varImportance(gis_ts2_rf,"Model 6"))
-var_importance$variable_names<-factor(var_importance$variable_names)
+#var_importance$variable_names<-factor(var_importance$variable_names)
 #var_importance$model_id<-factor(var_importance$model_id)
 var_importance<-inner_join(var_importance,data_def[c("variable_names","description")],"variable_names")
 var_importance<-var_importance[order(var_importance$model_id,-var_importance$mean_decrease_gini, decreasing=FALSE),]                     
@@ -179,14 +183,14 @@ for(i in vars){
 }
 all_vars<-unique(all_vars)
 
-all_movies_palette<-vector()
-for(i in 1:dim(namelist)[1]){
-  all_movies_palette<-c(all_movies_palette,wes.palette(namelist[i,2],
-                                                       namelist[i,1]))
-}
-all_movies_palette<-unique(all_movies_palette)
-zissou<-c(wes.palette(5,"Zissou"),wes.palette(5,"Rushmore"))
[…]

### 2015-04-06 vignettes/submission/prepare_submission.R

@@ -1,3 +1,10 @@
-copy<-function(){
-  file.copy("../manuscript.Rmd","manuscript.Rmd",overwrite=T)
+# Function to copy manuscript.Rmd from vignette into submission
+copy<-function(file){
+  file.copy(file,"manuscript.Rmd",overwrite=TRUE)
+  #system("cp ../manuscript.Rmd .")
+  system("cp -r ../components .")
+}
+
+render_doc<-function(file){
+  rmarkdown::render(file,output_file = "manuscript.docx", output_format = "all")
 }

### 2015-06-09 R/crossval_rf.R
new crossval and some other changes

@@ -1,0 +1,53 @@
+#' Cross Validate Random Forests to get Categorical Accuracy
+#'
+#' Funtion to cross-validate categorical prediction accuracy from 
+#' resultant chl a.  A Random forest of forests...  Used to get categorical
+#' accuracy.  Also returns a cross-validated accuracy of % var explained.
+#' 
+#' @param y response a vector 
+#' @param x predictors a data.frame
+#' @param breaks numeric vector of cut points
+#' @param cat categories of response
+#' @param split proportion to include in training datset.  test set to inverse.
+#' @param n number of iterations
+#' @import broom party
+#' @export
+#' @examples
+#' data(LakeTrophicModelling)
+#' predictors_all <- predictors_all[predictors_all!="DATE_COL"]
+#' all_cf_dat <- data.frame(ltmData[predictors_all],LogCHLA=log10(ltmData$CHLA))
+#' all_cf_dat <- all_cf_dat[complete.cases(all_cf_dat),]
+#' ts_brks <- c(min(all_cf_dat$LogCHLA),log10(2),log10(7),log10(30),max(all_cf_dat$LogCHLA))
+#' ts_cats <- c("oligo","meso","eu","hyper")
+#' x<-crossval_rf(all_cf_dat$LogCHLA,all_cf_dat[,names(all_cf_dat)!="LogCHLA"],
+#'             ts_brks,ts_cats,0.8,100,100)
+crossval_rf <- function(y,x, breaks, cat=NULL, split, n, ntree){
+  out <- list(summary=NULL,acc=NULL,all=list())
+  dat <- data.frame(y=y,x)
+  for(i in 1:n){
+    train_idx <- sample(1:nrow(dat),split*nrow(dat))
+    test_idx <-!1:nrow(dat)%in%train_idx
+    dat_train <-dat[train_idx,] 
+    dat_test <-dat[test_idx,]
+    cf_train <- cforest(y~., dat=dat_train,
+                        control=cforest_unbiased(ntree=ntree,
+                                                 mtry=ncol(dat_train)/3))
+    cf_pred_test <- as.numeric(predict(cf_train,newdata=dat_test))
+    cf_orig_test <- dat_test$y
+    cf_pred_test_cat <- cut(cf_pred_test,breaks,cat)
+    cf_orig_test_cat <- cut(dat_test$y,breaks,cat)
[…]

### 2015-06-09 R/crossval_rf.R
new udpates

@@ -20,7 +20,7 @@
 #' ts_brks <- c(min(all_cf_dat$LogCHLA),log10(2),log10(7),log10(30),max(all_cf_dat$LogCHLA))
 #' ts_cats <- c("oligo","meso","eu","hyper")
 #' x<-crossval_rf(all_cf_dat$LogCHLA,all_cf_dat[,names(all_cf_dat)!="LogCHLA"],
-#'             ts_brks,ts_cats,0.8,100,100)
+#'             ts_brks,ts_cats,0.8,1000,1000)
 crossval_rf <- function(y,x, breaks, cat=NULL, split, n, ntree){
   out <- list(summary=NULL,acc=NULL,all=list())
   dat <- data.frame(y=y,x)

### 2015-06-10 R/varsel_regression_rf.R
finishing touches on varsel

@@ -12,14 +12,30 @@
 #' data(LakeTrophicModelling)
 #' predictors_all <- predictors_all[predictors_all!="DATE_COL"]
 #' all_dat <- data.frame(ltmData[predictors_all],LogCHLA=log10(ltmData$CHLA))
-#' all_dat <- all_cf_dat[complete.cases(all_cf_dat),]
-#' x<-varsel_regression_rf(all_dat$LogCHLA,all_cf_dat[,names(all_cf_dat)!="LogCHLA"])
+#' all_dat <- all_dat[complete.cases(all_dat),]
+#' x<-varsel_regression_rf(all_dat$LogCHLA,all_dat[,names(all_dat)!="LogCHLA"],
+#'                         ntree=100)
 varsel_regression_rf <- function(y,x,...){
-  out <- list()
+  out <- list(mse=NULL,rsq=NULL,num_var=NULL,vars=NULL)
   dat <- data.frame(y=y,x)
-  init_rf <- randomForest(y~.,data=dat,...)
+  init_rf <- randomForest(y=dat$y,x=dat[,names(x)],...)
   init_imp <- importance(init_rf)
-  browser()
-  for(i in 1:n){
+  var_sort <- rownames(init_imp)[order(init_imp[,1],decreasing = TRUE)]
+  vars <- NULL
+  for(i in var_sort){
+    if(is.null(vars)){ 
+      vars <- c(vars,i) 
+      idx <- 1
+    } else {
+      vars <- c(vars,i)
+      vars_rf <- randomForest(y=dat$y,x=dat[,vars],...)
+      out$mse[[idx]] <- vars_rf$mse[length(vars_rf$mse)]
+      out$rsq[[idx]] <- vars_rf$rsq[length(vars_rf$rsq)]
+      out$num_var[[idx]] <- length(vars)
+      out$vars[[idx]] <- vars
+      idx <- idx + 1
+    }
+    print(paste(i/length(var_sort),"% completed"))
   }
+  return(out)
 }

### 2015-06-10 R/varsel_plot.R
varsel plot

@@ -1,0 +1,10 @@
+#' Plot varsel_regression output
+#' 
+#' @param vs_reg_obj Object from output of \link{varsel_regression_rf}.
+#' @export
+#' @import ggplot2
+varsel_plot <- function(vs_reg_obj){
+  dat <- data.frame(num = vs_reg_obj$num_var, mse = vs_reg_obj$mse)
+  plot_out<-ggplot(dat, aes(x=num,y=mse))+geom_point()
+  return(plot_out)
+}

### 2015-06-10 R/varsel_regression_rf.R

@@ -35,7 +35,7 @@ varsel_regression_rf <- function(y,x,...){
       out$vars[[idx]] <- vars
       idx <- idx + 1
     }
-    print(paste(i/length(var_sort),"% completed"))
+    print(paste(idx/length(var_sort),"% completed"))
   }
   return(out)
 }

### 2015-06-11 R/varsel_regression_rf.R
varsel_regression and removed prior cache

@@ -14,8 +14,8 @@
 #' all_dat <- data.frame(ltmData[predictors_all],LogCHLA=log10(ltmData$CHLA))
 #' all_dat <- all_dat[complete.cases(all_dat),]
 #' x<-varsel_regression_rf(all_dat$LogCHLA,all_dat[,names(all_dat)!="LogCHLA"],
-#'                         ntree=100)
-varsel_regression_rf <- function(y,x,...){
+#'                         ntree=100,prog=T)
+varsel_regression_rf <- function(y,x,prog=F,...){
   out <- list(mse=NULL,rsq=NULL,num_var=NULL,vars=NULL)
   dat <- data.frame(y=y,x)
   init_rf <- randomForest(y=dat$y,x=dat[,names(x)],...)
@@ -35,7 +35,9 @@ varsel_regression_rf <- function(y,x,...){
       out$vars[[idx]] <- vars
       idx <- idx + 1
     }
-    print(paste(idx/length(var_sort),"% completed"))
+    if(prog & idx%%3==0){
+      print(paste0(round(idx/length(var_sort)*100,1),"% completed"))
+    }
   }
   return(out)
 }

### 2015-06-11 R/varImportance.R
more changes

@@ -1,19 +1,19 @@
 #' Variable Importance Summary
 #' 
-#' @param rf Random Forest Object
+#' @param rf_imp a Random Forest Importance Object
 #' @param label A label for the model
 #' @export
-varImportance<-function(rf,label=NULL){
+varImportance<-function(rf_imp,label=NULL){
   if(!is.null(label)){
-    xdf<-data.frame(model_id=rep(label,dim(rf$importance)[1]),
-             variable_names=row.names(rf$importance),
-             mean_decrease_acc=data.frame(rf$importance)$MeanDecreaseAccuracy,
-             mean_decrease_gini=data.frame(rf$importance)$MeanDecreaseGini)
+    xdf<-data.frame(model_id=rep(label,nrow(rf_imp)),
+             variable_names=row.names(rf_imp),
+             mean_decrease_acc=rf_imp[,1],
+             mean_decrease_gini=rf_imp[,2])
   } else {
-    xdf<-data.frame(model_id=rep(NA,dim(rf$importance)[1]),
-                    variable_names=row.names(rf$importance),
-                    mean_decrease_acc=data.frame(rf$importance)$MeanDecreaseAccuracy,
-                    mean_decrease_gini=data.frame(rf$importance)$MeanDecreaseGini)
+    xdf<-data.frame(model_id=rep(NA,nrow(rf_imp),
+                                 variable_names=row.names(rf_imp)),
+                                 mean_decrease_acc=rf_imp[,1],
+                                 mean_decrease_gini=rf_imp[,2])
   }   
   return(xdf)
 }

### 2015-06-18 R/class_prob_rf.R
whole bunch a stuff

@@ -1,0 +1,35 @@
+#' Get Class Prediction Probabilities
+#' 
+#' This function takes a regression randomForest and returns class probabilities
+#' for a provided classifcation
+#' 
+#' @param rf_obj a randomForest object of type regression
+#' @param newdata data frame used to make predictions. Can be same as original
+#'                data.  
+#' @param breaks a numeric vector of values used to turn predictions into classes
+#' @param labels a character vector of labels for the classes
+#' @param ordered logical indicating if categories are ordered (in order specified 
+#'                in labels)
+#' @param type a character vector indicating whether total votes or probabilites
+#'              should be returned
+#' @return returns a dataframe of probabilites or number of votes for each class
+#' @export
+class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
+                             type=c("probs","votes")){
+  if(rf_obj$type!="regression"){
+    stop("rf_obj is not a regression randomForest object.")
+  }
+  type <- match.arg(type)
+  preds <- predict(all_rf,newdata=all_dat,predict.all=TRUE)
+  class_prob <- apply(preds$individual,2,
+                          function(x) cut(x,breaks,labels))
+  if(type=="probs"){
+    class_prob <- apply(class_prob,1, function(x)
+      table(factor(x,levels=labels, ordered=ordered))/rf_obj$ntree)
+  } else {
+    class_prob <- apply(class_prob,1, function(x)
+      table(factor(x,levels=labels, ordered=ordered)))
+  }
+  
+  return(data.frame(t(class_prob)))  
+}  

### 2015-06-18 R/class_prob_rf.R
more abstract fiddlining

@@ -20,7 +20,7 @@ class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
     stop("rf_obj is not a regression randomForest object.")
   }
   type <- match.arg(type)
-  preds <- predict(all_rf,newdata=all_dat,predict.all=TRUE)
+  preds <- predict(rf_obj,newdata=newdata,predict.all=TRUE)
   class_prob <- apply(preds$individual,2,
                           function(x) cut(x,breaks,labels))
   if(type=="probs"){

### 2015-07-09 R/varImportance.R
rendered and edits up to line 283

@@ -4,14 +4,15 @@
 #' @param label A label for the model
 #' @export
 varImportance<-function(rf_imp,label=NULL){
+
   if(!is.null(label)){
     xdf<-data.frame(model_id=rep(label,nrow(rf_imp)),
              variable_names=row.names(rf_imp),
              mean_decrease_acc=rf_imp[,1],
              mean_decrease_gini=rf_imp[,2])
   } else {
-    xdf<-data.frame(model_id=rep(NA,nrow(rf_imp),
-                                 variable_names=row.names(rf_imp)),
+    xdf<-data.frame(model_id=rep(NA,nrow(rf_imp)),
+                                 variable_names=row.names(rf_imp),
                                  mean_decrease_acc=rf_imp[,1],
                                  mean_decrease_gini=rf_imp[,2])
   }   

### 2015-07-12 R/importancePlot.R
updated variable importance plot.

@@ -15,10 +15,9 @@
 #' importancePlot(iris_rf,'acc',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 #' importancePlot(iris_rf,'gini',size=10,aes(colour=wes.palette(4, 'GrandBudapest')))
 
-importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"), ...) {
+importancePlot <- function(rf, sumtbl=NULL, data_def = NULL, type = c("acc", "gini"), ...) {
     
     idf<-importance(rf)
-    
     # set up data.frame for ggplot
     if (type == "gini") {
         imp_df <- data.frame(variables = names(idf[,dim(idf)[2]]), 
@@ -33,11 +32,13 @@ importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"),
       o <- order(imp_df$Mean_Decrease, decreasing = FALSE)
       imp_df$variables <- factor(imp_df$variables, 
                                  levels = imp_df$variables[o], ordered = T)
-        label <- "Mean Decrease Accuracy"
+        label <- "Percent Increase in Mean Square Error"
     } else {
         stop("Type not specificed correctly")
     }
-    imp_df<-merge(imp_df,sumtbl,by.x="variables",by.y="Variable")
+    if(!is.null(sumtbl)){
+      imp_df<-merge(imp_df,sumtbl,by.x="variables",by.y="Variable")
+    } 
     
     if (!is.null(data_def)){
       imp_df$variables<-data_def$description[match(imp_df$variables,data_def$variable_names)]
@@ -45,21 +46,7 @@ importancePlot <- function(rf, sumtbl, data_def = NULL, type = c("acc", "gini"),
       imp_df$variables <- factor(imp_df$variables, 
                                  levels = imp_df$variables[o], ordered = T)
     }
-    
     x <- ggplot(imp_df, aes(Mean_Decrease, variables)) + 
-      geom_blank() +  
-      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 25))+
-      theme(text = element_text(family="Times"),
-            panel.background = element_blank(), panel.grid = element_blank(), 
-            panel.border = element_blank(), 
-            axis.title.x = element_blank(),
-            axis.text.x = element_blank(),
-            axis.text.y = element_text(family="sans",size = 11),
-            axis.ticks.x = element_blank(),
[…]

### 2015-07-18 R/table_to_df.R
updated confusion matrix tables

@@ -1,0 +1,13 @@
+#' convert table to df
+#' 
+#' @param cm_tab a table class with confusion matrix
+#' 
+#' @export
+table_to_df <- function(cm_tab){
+  nc <- ncol(cm_tab)
+  nr <- nrow(cm_tab)
+  df<-data.frame(matrix(cm_tab[1:(nc*nr)],ncol=nc,nrow=nr))
+  names(df)<-attributes(cm_tab)[[2]][[1]]
+  row.names(df)<-names(df)
+  return(df)
+}

### 2015-07-20 R/ts_prob_map.R
updated render and new save options on prob maps

@@ -21,12 +21,18 @@
 #' ts_prob_map(state,lakes_dd,gis_rf_ts_prob)
 #' @export
 #' @import ggplot2
-ts_prob_map<-function(states,lakes,probs){
+ts_prob_map<-function(states,lakes,probs,save_sep=FALSE){
   lakes<-merge(lakes,probs,by="nla_id")
   gmap_o<-make_ts_map(states,lakes,"Oligotrophic",low="#bdd7e7",high="#08519c")
   gmap_m<-make_ts_map(states,lakes,"Mesotrophic", low="#efedf5",high="#756bb1")
   gmap_e<-make_ts_map(states,lakes,"Eutrophic", low="#fee6ce",high="#e6550d")
   gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", low="#fee0d2",high="#de2d26")
+  if(save_sep){
+    ggsave("oligo_prob.jpg",gmap_o,width=8.5)
+    ggsave("meso_prob.jpg",gmap_m,width=8.5)
+    ggsave("eu_prob.jpg",gmap_e,width=8.5)
+    ggsave("hyper_prob.jpg",gmap_h,width=8.5)
+  }
   return(multiplot(gmap_o,gmap_e,gmap_m,gmap_h,cols=2))
 }
 
@@ -46,7 +52,7 @@ make_ts_map <- function(states,lake_prob,title,low,high){
   if(title=="Hypereutrophic"){value<<-"hyper"}
   gmap<-ggplot(states,aes(x=long,y=lat))+
     geom_polygon(aes(group=group),fill="white",colour="grey")+
-    geom_point(data=lakes,aes_string(x="long",y="lat",colour=value),size=2.5)+
+    geom_point(data=lake_prob,aes_string(x="long",y="lat",colour=value),size=2.5)+
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
     theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
           panel.border = element_blank(), 

### 2015-07-21 R/prob_cdf.R
working on cdf of lake predicion probabilites.

@@ -1,0 +1,43 @@
+#' Function to create CDFs of probability for each trophic state prediction
+#' 
+#' 
+#' This functions produces CDF plots from predicted trophic state probabilities   
+#' 
+#' 
+#' @param probs the continuous Variable such as-'sumBioV','CHLA', 'NTL', 'PTL', etc.  Of same 
+#'                length as contVar 
+#' @param cdf_colors colors for CDF plots and median line -should equal # of categories in catVar + 1
+#' @param ... pass additional parameters to labs (e.g. set titles, axis labels, etc.)
+#' 
+#' @examples
+#' devtools::install_github('wesanderson','karthik')
+#' library(wesanderson)
+#' data(LakeTrophicModelling)
+#' 
+#' @export
+#' @import ggplot2
+
+prob_cdf <- function(probs cdf_colors = 1:length(catVar), ...) {
+
+    options(scipen = 5)  #tell r not to use scientific notation on axis labels
+    
+    keep<-complete.cases(catVar,contVar)
+    cdf_df<-data.frame(catVar,contVar)[keep,]
+    x <- ggplot(cdf_df,aes(contVar,colour=catVar)) +
+          stat_ecdf(size=2)+scale_color_manual(values=cdf_colors)+
+          scale_x_log10()+
+          theme(text = element_text(family="Times"),
+                panel.background = element_blank(), panel.grid = element_blank(), 
+                panel.border = element_rect(fill = NA), 
+                plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
+                legend.position = c(0.15,0.75), legend.key = element_rect(fill = 'white'),
+                legend.text = element_text(family="sans",size=15), legend.title = element_text(family="sans",size=11),
+                axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+                axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+                axis.text.x = element_text(family="sans",size = 11),
+                axis.text.y = element_text(family="sans",size = 11)) + 
+          geom_hline(linetype = 3, size = 1, colour = "gray", 
+                     yintercept = c(0,0.5,1)) +
+          labs(...)
+    return(x)        
+} 

### 2015-07-22 R/prob_cdf.R
updated prod cdf finished All Variables results

@@ -4,28 +4,27 @@
 #' This functions produces CDF plots from predicted trophic state probabilities   
 #' 
 #' 
-#' @param probs the continuous Variable such as-'sumBioV','CHLA', 'NTL', 'PTL', etc.  Of same 
-#'                length as contVar 
+#' @param probs1 prediction probabilities for one of the models
+#' @param probs2 prediction probabilities for the other model
 #' @param ... pass additional parameters to labs (e.g. set titles, axis labels, etc.)
 #' 
 #' @examples
 #' devtools::install_github('wesanderson','karthik')
 #' library(wesanderson)
 #' data(LakeTrophicModelling)
-#' prob_cdf(gis_rf_ts_prob)
+#' prob_cdf(all_rf_ts_prob, gis_rf_ts_prob)
 #' @export
 #' @import ggplot2
 
-prob_cdf <- function(probs, ...) {
-    browser()
+prob_cdf <- function(probs1, probs2, ...) {
+    
     options(scipen = 5)  #tell r not to use scientific notation on axis labels
-    probs$max <- apply(probs[,1:4],1,max)
-    x <- ggplot(probs) +
-          stat_ecdf(color = "black", size = 3 , aes(max)) +
-          stat_ecdf(color = "blue", size = 3 , aes(oligo)) +
-          stat_ecdf(color = "yellow", size = 3 , aes(meso)) +
-          stat_ecdf(color = "orange", size = 3 , aes(eu)) +
-          stat_ecdf(color = "red", size = 3 , aes(hyper)) +
+    probs1$max <- apply(probs1[,1:4],1,max)
+    probs2$max <- apply(probs2[,1:4],1,max)
+    
+    x <- ggplot(data=probs1) +
+          stat_ecdf(color = "black", size = 3 , aes(x=max)) +
+          stat_ecdf(data=probs2, color = "red", size = 3 , aes(x=max)) +
           theme(text = element_text(family="Times"),
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 

### 2015-07-23 R/ecdf_ks_ci.R
updates plus first shot of cdf CI.  Need to work on prob_cdf to get working with CI

@@ -1,0 +1,17 @@
+#' Kolmogrov-Smirnoff CIs for a ecdf
+#' 
+#' Code from sfsmisc::ecdf.KSci. Ammended here to simply output the xy for both upper and 
+#' lower for use as input to ggplot
+#' @export
+ecdf_ks_ci <- function(x){
+  n <- length(x)
+  ec <- ecdf(x)
+  xx <- get("x", envir = environment(ec))
+  yy <- get("y", envir = environment(ec))
+  D <- sfsmisc::KSd(n)
+  yyu <- pmin(yy + D, 1)
+  yyl <- pmax(yy - D, 0)
+  ecu <- stepfun(xx, c(yyu, 1))
+  ecl <- stepfun(xx, c(yyl, yyl[n]))
+  return(list(upper=ecu(xx),lower=ecl(xx)))
+}

### 2015-07-24 R/class_prob_rf.R
almost done with results

@@ -33,6 +33,9 @@ class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
   }
   class_prob<-data.frame(t(class_prob))
   class_prob$nla_id<-row.names(class_prob)
+  class_prob$max <- apply(class_prob[,1:4],1,max)
+  class_prob$pred_class <- cut(preds$aggregate,breaks,labels)
+  class_prob$obs_class <- cut(rf_obj$y,breaks,labels)
   row.names(class_prob)<-as.character(1:nrow(class_prob))
   return(class_prob)  
 }  

### 2015-07-24 R/ecdf_ks_ci.R

@@ -3,6 +3,8 @@
 #' Code from sfsmisc::ecdf.KSci. Ammended here to simply output the xy for both upper and 
 #' lower for use as input to ggplot
 #' @export
+#' @examples
+#' ecdf_ks_ci(all_pred_prob)
 ecdf_ks_ci <- function(x){
   n <- length(x)
   ec <- ecdf(x)
@@ -13,5 +15,5 @@ ecdf_ks_ci <- function(x){
   yyl <- pmax(yy - D, 0)
   ecu <- stepfun(xx, c(yyu, 1))
   ecl <- stepfun(xx, c(yyl, yyl[n]))
-  return(list(upper=ecu(xx),lower=ecl(xx)))
+  return(list(x=x, upper=ecu(x),lower=ecl(x)))
 }

### 2015-07-24 R/prob_cdf.R

@@ -12,25 +12,24 @@
 #' devtools::install_github('wesanderson','karthik')
 #' library(wesanderson)
 #' data(LakeTrophicModelling)
-#' prob_cdf(all_rf_ts_prob, gis_rf_ts_prob)
+#' prob_cdf(all_rf_ts_prob, gis_rf_ts_prob,x="Prediction Probability",y="Proportion of Samples")
 #' @export
 #' @import ggplot2
 
 prob_cdf <- function(probs1, probs2, ...) {
-    #FIX ME
     options(scipen = 5)  #tell r not to use scientific notation on axis labels
     probs1$max <- apply(probs1[,1:4],1,max)
-    probs1$maxul <- ecdf_ks_ci(probs1$max)$upper
-    probs1$maxll <- ecdf_ks_ci(probs1$max)$lower  
+    ci1 <- na.omit(data.frame(ecdf_ks_ci(probs1$max))) 
     probs2$max <- apply(probs2[,1:4],1,max)
-    probs2$maxul <- ecdf_ks_ci(probs2$max)$upper
-    probs2$maxll <- ecdf_ks_ci(probs2$max)$lower  
+    ci2 <- na.omit(data.frame(ecdf_ks_ci(probs2$max)))
     
     x <- ggplot(data=probs1) +
-          stat_ecdf(color = "black", size = 3 , aes(x=max)) +
-          geom_ribbon(aes(ymin = maxul,ymax = maxll),alpha = 0.2) +
-          stat_ecdf(data=probs2, color = "red", size = 3 , aes(x=max)) +
-          geom_ribbon(data=probs2, aes(ymin = maxul,ymax = maxll),alpha = 0.2) +
+          stat_ecdf(color = "black", size = 2 , aes(x=max)) +
+          geom_ribbon(data=ci1, aes(x=x, ymin = lower,ymax = upper),fill="black",
+                      alpha = 0.4) +
+          stat_ecdf(data=probs2, color = "red", size = 2 , aes(x=max)) +
+          geom_ribbon(data=ci2, aes(x=x, ymin = lower,ymax = upper),fill="red",
+                      alpha = 0.4) +
           theme(text = element_text(family="Times"),
                 panel.background = element_blank(), panel.grid = element_blank(), 
                 panel.border = element_rect(fill = NA), 

### 2015-08-10 R/condAccuracy.R
summary table of prob and acc

@@ -25,6 +25,8 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
   #plot(cp$max_vote,cp$Raw.Data.Probability)
   cp1<-data.frame(max_vote=cp1$max_vote,Raw.Data.Probability=cp1$Raw.Data.Probability)
   cp2<-data.frame(max_vote=cp2$max_vote,Raw.Data.Probability=cp2$Raw.Data.Probability)
+  cp1<<-cp1
+  cp2<<-cp2
   ggplot(cp1,aes(x=max_vote,y=Raw.Data.Probability))+
     geom_point(size=2,color="red")+
     geom_point(data=cp2,aes(x=max_vote,y=Raw.Data.Probability),size=2,color="black") +

### 2015-08-18 R/pred_prob_map.R
prob section draft done, new map, fixed fig refs in pd section

@@ -1,0 +1,48 @@
+#' Create map of discrete predicted probabilities
+#' 
+#' This function takes the predicted class probabilites and maps them for the
+#' predicted class
+#' 
+#' @param states data for states, as data.frame
+#' @param lakes point locations for lake samples as data.frame with NLA ID
+#' @param probs class probabilites with NLA ID
+#' 
+#' @examples
+#' state<-map_data('state')
+#' lakes_alb<-data.frame(ltmData[["AlbersX"]],ltmData[["AlbersY"]],
+#'                       ltmData[["NLA_ID"]])
+#' p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" 
+#' ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs" 
+#' lakes_alb_sp<-SpatialPointsDataFrame(coordinates(lakes_alb[,1:2]),proj4string=CRS(p4s),
+#'                                      data=data.frame(nla_id = as.character(lakes_alb[,3])))
+#' lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
+#' lakes_dd<-data.frame(coordinates(lakes_dd),lakes_dd$nla_id)
+#' names(lakes_dd)<-c("long","lat","nla_id")
+#' pred_prob_map(state,lakes_dd,gis_rf_ts_prob)
+#' @export
+#' @import ggplot2
+pred_prob_map<-function(states,lakes,probs,save_sep=FALSE){
+  lakes<-merge(lakes,probs,by="nla_id")
+  value<-"max"
+  title<-"Predicted Probability"
+  low<-"#E0EEEE"
+  high<-"#00688B"
+  mid_pt<-mean(range(lakes$max))
+  gmap<-ggplot(states,aes(x=long,y=lat))+
+    geom_polygon(aes(group=group),fill="white",colour="grey")+
+    geom_point(data=lakes,aes_string(x="long",y="lat",colour=value),size=3.5)+
+    coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
+    theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
+          panel.border = element_blank(), 
+          axis.text = element_blank(),axis.ticks = element_blank(),
[…]

### 2015-08-19 R/scatterPlot.R
first revised draft!

@@ -28,7 +28,7 @@ scatterPlot<-function(df,xvar,yvar,pt_col,ln_col,cl_col,...){
   
   
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=2,colour=pt_col) +
+    geom_point(size=3,colour=pt_col) +
     scale_x_log10() +
     scale_y_log10() +
     #stat_smooth(method="lm",fill=cl_col,

### 2015-08-20 R/density_plot.R
draft to send to B&B

@@ -1,0 +1,28 @@
+#' density plot for lake trophic modeling paper
+#' 
+#' Produces a density plot via ggplot
+#' 
+#' @examples
+#' density_plot(gis_rf_ts_prob,xvar="max",x="Predicted Probability")
+#' @export
+#' @import ggplot2
+
+density_plot<-function(df,xvar,...){
+  options(scipen=5)
+  xdf<-data.frame(xvar=df[[xvar]])
+  x <- ggplot(xdf,aes(x=xvar))+
+    geom_density() +
+    theme(text = element_text(family="sans"),
+          panel.background = element_blank(),
+          panel.border = element_rect(fill = NA), 
+          plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
+          legend.position = "none", legend.key = element_rect(fill = 'white'),
+          legend.text = element_text(family="sans",size=15), legend.title = element_text(size=11),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="sans",size = 11),
+          axis.text.y = element_text(family="sans",size = 11)) + 
+    labs(...)
+    
+  return(x)
+}

### 2015-08-31 R/ecor_map.R
working on final figs

@@ -18,9 +18,9 @@ ecor_map<-function(ecor){
   ecor_f <- fortify(ecor, region="WSA_9")
   names(ecor_f)[which(names(ecor_f)=="id")] = "WSA_9"
   ecor_df <- plyr::join(ecor_f, ecor@data, by="WSA_9")
-  gmap<-ggplot(ecor_df, aes(long,lat,group=group,fill=id)) +
-    geom_polygon(data=subset(ecor_df,WSA_9!="WMT")) +
-    geom_polygon(data=subset(ecor_df,WSA_9=="WMT")) +
+  gmap<-ggplot(ecor_df, aes(long,lat,group=group,fill=WSA_9)) +
+    geom_polygon(data=subset(ecor_df,WSA_9!="Western Mountains (WMT)")) +
+    geom_polygon(data=subset(ecor_df,WSA_9=="Western Mountains (WMT)")) +
     geom_path(color="white") +
     coord_equal() +
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
@@ -28,6 +28,8 @@ ecor_map<-function(ecor){
           panel.border = element_blank(), #legend.position = "none",
           axis.text = element_blank(),axis.ticks = element_blank()) +
     ylab("") +
-    xlab("")
+    xlab("") +
+    scale_fill_manual(values=viridis(10))
+    
   return(gmap)
 }

### 2015-08-31 R/nlaMap.R

@@ -4,6 +4,7 @@
 #' 
 #' @param states data for states, as data.frame
 #' @param lakes point locations for lake samples as data.frame
+#' @param cats plot ts categories?
 #' @param myColor vector of length 3 with colors for fill, lines, and points in that order
 #' 
 #' @examples
@@ -18,17 +19,32 @@
 #' nlaMap(state,lakes_dd)
 #' @export
 #' @import ggplot2
-nlaMap<-function(states,lakes,myColor=1:3){
-  #browser()
-
+nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
+  #Color Blind Friendly Palette
+  if(cats){
+    gmap<-ggplot(states,aes(x=long,y=lat))+
+      geom_polygon(aes(group=group),fill=myColor[2],colour=myColor[1])+
+      geom_point(data=lakes,aes(x=long,y=lat,color=factor(data)),size=2.5)+
+      scale_color_manual(values=catColor,
+                         name="Trophic\nState",
+                         breaks=c("Oligo", "Meso", "Eu","Hyper"),
+                         labels=c("Oligotrophc", "Mesotrophic", "Eutrophic","Hypereutrophic"))+
+      coord_map("albers", lat2 = 45.5, lat1 = 29.5) +
+      theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
+            panel.border = element_blank(),
+            axis.text = element_blank(),axis.ticks = element_blank()) + 
+      ylab("") + 
+      xlab("")
+  } else {
   gmap<-ggplot(states,aes(x=long,y=lat))+
                geom_polygon(aes(group=group),fill=myColor[2],colour=myColor[1])+
-               geom_point(data=lakes,aes(x=long,y=lat),size=2,colour=myColor[3])+
+               geom_point(data=lakes,aes(x=long,y=lat),size=2.5,colour=myColor[3])+
                coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
                theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
                      panel.border = element_blank(), legend.position = "none", 
                      axis.text = element_blank(),axis.ticks = element_blank()) + 
                ylab("") + 
                xlab("")
+  }
   return(gmap)
 }

### 2015-08-31 R/partial_plot.R

@@ -1,0 +1,50 @@
+#' Create parital dependence plot
+#' 
+#' This function takes the output of randomForest::paritalPlot() and
+#' generates a plot via ggplot.
+#' 
+#' @param pd partial dependence from randomForest::partialPlot()
+#' @export
+partial_plot <- function(pd,xl,yl,...){
+  y_scale<-c(min(pd$y),min(pd$y)+diff(range(pd$y))/2,max(pd$y))
+  if(!is.factor(pd$x)){
+    ggp<-ggplot(data.frame(pd),aes(x=x,y=y))+
+      geom_line()+
+      theme(text = element_text(family="sans"),
+                        panel.background = element_blank(), panel.grid = element_blank(), 
+                        panel.border = element_rect(fill = NA), 
+                        plot.title  = element_text(family="sans",size=12,face="bold",
+                                                   vjust=1.1),
+                        legend.position = "none", legend.key = element_rect(fill = 'white'),
+                        legend.text = element_text(family="sans",size=15), 
+                        legend.title = element_text(size=11),
+                        axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+                        axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+                        axis.text.x = element_text(family="sans",size = 11),
+                        axis.text.y = element_text(family="sans",size = 11),
+                        ...) + 
+      labs(x=xl,y=yl)+
+      scale_y_continuous(breaks=y_scale,
+                         label=round(y_scale,2))
+  } else {
+    ggp<-ggplot(data.frame(pd),aes(x=x,y=y))+
+      geom_bar(stat = "identity")+
+      theme(text = element_text(family="sans"),
+            panel.background = element_blank(), panel.grid = element_blank(), 
+            panel.border = element_rect(fill = NA), 
+            plot.title  = element_text(family="sans",size=12,face="bold",
+                                       vjust=1.1),
+            legend.position = "none", legend.key = element_rect(fill = 'white'),
+            legend.text = element_text(family="sans",size=15), 
+            legend.title = element_text(size=11),
+            axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+            axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+            axis.text.x = element_text(angle=50, vjust=0.5,family="sans",size = 11),
+            axis.text.y = element_text(family="sans",size = 11),
+            ...) + 
+      labs(x=xl,y=yl)
+    
+  }
+  return(ggp) 
+}  
+

### 2015-09-01 R/condAccuracy.R
closer to final

@@ -23,22 +23,26 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
   correct<-pred==obs
   cp2<-condprob2::condprob(max_vote,correct,ProbComp="gt",Exceed="gte",...)
   #plot(cp$max_vote,cp$Raw.Data.Probability)
-  cp1<-data.frame(max_vote=cp1$max_vote,Raw.Data.Probability=cp1$Raw.Data.Probability)
-  cp2<-data.frame(max_vote=cp2$max_vote,Raw.Data.Probability=cp2$Raw.Data.Probability)
+  cp1<-data.frame(max_vote=cp1$max_vote,Raw.Data.Probability=cp1$Raw.Data.Probability,
+                  model=rep(pred_prob1$model[1],length(cp1[[1]])))
+  cp2<-data.frame(max_vote=cp2$max_vote,Raw.Data.Probability=cp2$Raw.Data.Probability,
+                  model=rep(pred_prob2$model[1],length(cp2[[1]])))
+  df<-rbind(cp1,cp2)
   cp1<<-cp1
   cp2<<-cp2
-  ggplot(cp1,aes(x=max_vote,y=Raw.Data.Probability))+
-    geom_point(size=2,color="red")+
-    geom_point(data=cp2,aes(x=max_vote,y=Raw.Data.Probability),size=2,color="black") +
+  ggp<-ggplot(df,aes(x=max_vote,y=Raw.Data.Probability,colour=model))+
+    geom_point(size=2)+
     theme(text = element_text(family="sans"),
           panel.background = element_blank(), #panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
           plot.title  = element_text(family="sans",size=15,face="bold",vjust=1.1),
-          legend.position = "none", legend.key = element_rect(fill = 'white'),
+          legend.key = element_rect(fill = 'white'),
           legend.text = element_text(family="sans",size=15), legend.title = element_text(size=15),
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
           axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
           axis.text.x = element_text(family="sans",size = 11),
           axis.text.y = element_text(family="sans",size = 11)) + 
-    labs(y="Conditional Total Accuracy",x=xlab)
+    labs(y="Total Accuracy",x=xlab) +
+    scale_colour_manual(name='',values = viridis(2))
+  return(ggp)
 }

### 2015-09-01 R/ecor_map.R

@@ -26,10 +26,12 @@ ecor_map<-function(ecor){
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
     theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(),
           panel.border = element_blank(), #legend.position = "none",
-          axis.text = element_blank(),axis.ticks = element_blank()) +
+          axis.text = element_blank(),axis.ticks = element_blank(),
+          legend.text = element_text(size=10),
+          legend.key.width = unit(0.5, "line")) +
     ylab("") +
     xlab("") +
-    scale_fill_manual(values=viridis(10))
+    scale_fill_manual(name="",values=viridis(10))
     
   return(gmap)
 }

### 2015-09-01 R/nlaMap.R

@@ -32,7 +32,10 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
       coord_map("albers", lat2 = 45.5, lat1 = 29.5) +
       theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
             panel.border = element_blank(),
-            axis.text = element_blank(),axis.ticks = element_blank()) + 
+            axis.text = element_blank(),axis.ticks = element_blank(),
+            legend.key.width=unit(2, "line"), 
+            legend.key.height=unit(0.5, "line"),
+            legend.position="bottom", legend.direction="horizontal") + 
       ylab("") + 
       xlab("")
   } else {

### 2015-09-01 R/pred_prob_map.R

@@ -24,25 +24,27 @@
 pred_prob_map<-function(states,lakes,probs,save_sep=FALSE){
   lakes<-merge(lakes,probs,by="nla_id")
   value<-"max"
-  title<-"Predicted Probability"
-  low<-"#E0EEEE"
-  high<-"#00688B"
+  title<-""
+  low<-viridis(3)[2]
+  high<-viridis(3)[1]
   mid_pt<-mean(range(lakes$max))
   gmap<-ggplot(states,aes(x=long,y=lat))+
     geom_polygon(aes(group=group),fill="white",colour="grey")+
-    geom_point(data=lakes,aes_string(x="long",y="lat",colour=value),size=3.5)+
+    geom_point(data=lakes,aes_string(x="long",y="lat",colour=value),size=2.5)+
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
-    theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
-          panel.border = element_blank(), 
-          axis.text = element_blank(),axis.ticks = element_blank(),
-          plot.margin = grid::unit(c(0,0,0,0),"inches")) + 
+    theme(text = element_text(family="sans"),
+          panel.background = element_blank(), #panel.grid = element_blank(), 
+          panel.border = element_rect(fill = NA), 
+          plot.title  = element_text(family="sans",size=15,face="bold",vjust=1.1),
+          legend.key = element_rect(fill = 'white'),
+          legend.text = element_text(family="sans",size=15), legend.title = element_text(size=15),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+          axis.text.x = element_text(family="sans",size = 11),
+          axis.text.y = element_text(family="sans",size = 11)) + 
     ggtitle(title)+
     ylab("") + 
     xlab("") +
-    scale_colour_gradient2("Probability",midpoint = mid_pt) +
-    guides(fill = guide_legend(
-      title.theme = element_text(size=10),
-      keywidth = 0.5, keyheight = 0.5
-    ))
+    scale_colour_gradient2("Probability",low=low,midpoint = mid_pt,high=high)
   return(gmap)
 }

### 2015-09-01 R/prob_cdf.R

@@ -17,31 +17,26 @@
 #' @import ggplot2
 
 prob_cdf <- function(probs1, probs2, ...) {
-    options(scipen = 5)  #tell r not to use scientific notation on axis labels
-    probs1$max <- apply(probs1[,1:4],1,max)
-    ci1 <- na.omit(data.frame(ecdf_ks_ci(probs1$max))) 
-    probs2$max <- apply(probs2[,1:4],1,max)
-    ci2 <- na.omit(data.frame(ecdf_ks_ci(probs2$max)))
-    
-    x <- ggplot(data=probs1) +
-          stat_ecdf(color = "black", size = 2 , aes(x=max)) +
-          geom_ribbon(data=ci1, aes(x=x, ymin = lower,ymax = upper),fill="black",
-                      alpha = 0.4) +
-          stat_ecdf(data=probs2, color = "red", size = 2 , aes(x=max)) +
-          geom_ribbon(data=ci2, aes(x=x, ymin = lower,ymax = upper),fill="red",
-                      alpha = 0.4) +
-          theme(text = element_text(family="Times"),
-                panel.background = element_blank(), panel.grid = element_blank(), 
-                panel.border = element_rect(fill = NA), 
-                plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
-                legend.position = c(0.15,0.75), legend.key = element_rect(fill = 'white'),
-                legend.text = element_text(family="sans",size=15), legend.title = element_text(family="sans",size=11),
-                axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
-                axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
-                axis.text.x = element_text(family="sans",size = 11),
-                axis.text.y = element_text(family="sans",size = 11)) + 
-          geom_hline(linetype = 3, size = 1, colour = "gray", 
-                     yintercept = c(0,0.5,1)) +
-          labs(...)
+  df<-rbind(probs1,probs2)
+  df <- na.omit(data.frame(df))
+  options(scipen = 5)  #tell r not to use scientific notation on axis labels
+  x <- ggplot(df,aes(x=max,colour=model)) +
+    stat_ecdf(size=2) +
+    geom_ribbon(aes(ymin = lower,ymax = upper,fill=model),alpha = 0.4) +
+    theme(text = element_text(family="sans"),
+          panel.background = element_blank(), #panel.grid = element_blank(), 
+          panel.border = element_rect(fill = NA), 
[…]

### 2015-09-01 R/ts_prob_map.R

@@ -23,10 +23,10 @@
 #' @import ggplot2
 ts_prob_map<-function(states,lakes,probs,save_sep=FALSE){
   lakes<-merge(lakes,probs,by="nla_id")
-  gmap_o<-make_ts_map(states,lakes,"Oligotrophic",low="#bdd7e7",high="#08519c")
-  gmap_m<-make_ts_map(states,lakes,"Mesotrophic", low="#efedf5",high="#756bb1")
-  gmap_e<-make_ts_map(states,lakes,"Eutrophic", low="#fee6ce",high="#e6550d")
-  gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", low="#fee0d2",high="#de2d26")
+  gmap_o<-make_ts_map(states,lakes,"Oligotrophic",high=viridis(4)[1])#"#08519c")
+  gmap_m<-make_ts_map(states,lakes,"Mesotrophic",high=viridis(4)[2])#"#756bb1")
+  gmap_e<-make_ts_map(states,lakes,"Eutrophic", high=viridis(4)[3])#"#e6550d")
+  gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", high=viridis(4)[4])#"#de2d26")
   if(save_sep){
     ggsave("oligo_prob.jpg",gmap_o,width=8.5)
     ggsave("meso_prob.jpg",gmap_m,width=8.5)
@@ -51,20 +51,19 @@ make_ts_map <- function(states,lake_prob,title,low,high){
   if(title=="Eutrophic"){value<<-"eu"}
   if(title=="Hypereutrophic"){value<<-"hyper"}
   gmap<-ggplot(states,aes(x=long,y=lat))+
-    geom_polygon(aes(group=group),fill="white",colour="grey")+
     geom_point(data=lake_prob,aes_string(x="long",y="lat",colour=value),size=2.5)+
+    geom_polygon(aes(group=group),alpha=0,colour="grey")+
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
     theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
           panel.border = element_blank(), 
           axis.text = element_blank(),axis.ticks = element_blank(),
-          plot.margin = grid::unit(c(0,0,0,0),"inches")) + 
+          plot.margin = grid::unit(c(0,0,0,0),"inches"),
+          legend.key.width=unit(2, "line"), 
+          legend.key.height=unit(0.5, "line"),
[…]

### 2015-09-01 R/varsel_plot.R

@@ -5,6 +5,17 @@
 #' @import ggplot2
 varsel_plot <- function(vs_reg_obj){
   dat <- data.frame(num = vs_reg_obj$num_var, mse = vs_reg_obj$mse)
-  plot_out<-ggplot(dat, aes(x=num,y=mse))+geom_point()
+  plot_out<-ggplot(dat, aes(x=num,y=mse))+geom_point(size=2.5)+
+  theme(text = element_text(family="sans"),
+        panel.background = element_blank(), #panel.grid = element_blank(), 
+        panel.border = element_rect(fill = NA), 
+        plot.title  = element_text(family="sans",size=15,face="bold",vjust=1.1),
+        legend.key = element_rect(fill = 'white'),
+        legend.text = element_text(family="sans",size=15), legend.title = element_text(size=15),
+        axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
+        axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
+        axis.text.x = element_text(family="sans",size = 11),
+        axis.text.y = element_text(family="sans",size = 11))+
+    labs(x="Number of variables in model", y="Root Mean Square Error")
   return(plot_out)
 }

### 2015-11-18 R/condAccuracy.R
added kappa to output for tables and figures, need to add to graph and tbales

@@ -27,9 +27,22 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
                   model=rep(pred_prob1$model[1],length(cp1[[1]])))
   cp2<-data.frame(max_vote=cp2$max_vote,Raw.Data.Probability=cp2$Raw.Data.Probability,
                   model=rep(pred_prob2$model[1],length(cp2[[1]])))
+  
+  #browser()
+  cp1_k <- vector("numeric",nrow(cp1))
+  for(i in 1:nrow(cp1)){
+    x<-dplyr::filter(pred_prob1,max>=cp1$max_vote[i])
+    cp1_k[i] <- classAgreement(table(x$pred_class,x$obs_class))$kappa
+ 
+  }
+  cp2_k <- vector("numeric",nrow(cp2))
+  for(i in 1:nrow(cp2)){
+    x<-dplyr::filter(pred_prob2,max>=cp2$max_vote[i])
+    cp2_k[i] <- classAgreement(table(x$pred_class,x$obs_class))$kappa 
+  }
+  cp1<<-data.frame(cp1,kappa=cp1_k)
+  cp2<<-data.frame(cp2,kappa=cp2_k)
   df<-rbind(cp1,cp2)
-  cp1<<-cp1
-  cp2<<-cp2
   ggp<-ggplot(df,aes(x=max_vote,y=Raw.Data.Probability,colour=model))+
     geom_point(size=2)+
     theme(text = element_text(family="sans"),

### 2015-11-23 R/condAccuracy.R
more work on ecosphere revisions

@@ -42,8 +42,8 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
   }
   cp1<<-data.frame(cp1,kappa=cp1_k)
   cp2<<-data.frame(cp2,kappa=cp2_k)
-  df<-rbind(cp1,cp2)
-  ggp<-ggplot(df,aes(x=max_vote,y=Raw.Data.Probability,colour=model))+
+  df<-rbind(data.frame(cp1,kappa=cp1_k),data.frame(cp2,kappa=cp2_k))
+  ggp<-ggplot(df,aes(x=max_vote,y=kappa,colour=model))+
     geom_point(size=2)+
     theme(text = element_text(family="sans"),
           panel.background = element_blank(), #panel.grid = element_blank(), 
@@ -55,7 +55,7 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
           axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
           axis.text.x = element_text(family="sans",size = 11),
           axis.text.y = element_text(family="sans",size = 11)) + 
-    labs(y="Total Accuracy",x=xlab) +
+    labs(y="Kappa Coefficient",x=xlab) +
     scale_colour_manual(name='',values = viridis(2))
   return(ggp)
 }

### 2015-12-01 R/getCyanoAbund.R
fixed bum link to data plus resubmit commit.

@@ -14,7 +14,7 @@
 
   getCyanoAbund<-function(){
     #read the raw data from the EPA website
-      Raw<-read.csv('http://water.epa.gov/type/lakes/assessmonitor/lakessurvey/upload/NLA2007_Phytoplankton_SoftAlgaeCount_20091023.csv')
+      Raw<-read.csv('http://www2.epa.gov/sites/production/files/2014-10/nla2007_phytoplankton_softalgaecount_20091023.csv')
     #subset data-VISIT_NO==1 and Taxonomic Division=='Cyanophyta'
       Cyano<-subset(Raw[,c('SITE_ID','DIVISION','OTU','ABUND')],Raw$VISIT_NO==1 & Raw$DIVISION=='Cyanophyta')
     #Sum cyano abundance across taxa by NLA SITE_ID

### 2015-12-18 R/nlaMap.R
started on re-sizing figures...

@@ -35,7 +35,8 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
             axis.text = element_blank(),axis.ticks = element_blank(),
             legend.key.width=unit(2, "line"), 
             legend.key.height=unit(0.5, "line"),
-            legend.position="bottom", legend.direction="horizontal") + 
+            legend.position="bottom", legend.direction="horizontal",
+            plot.margin = unit(c(0,0,0,0),"in")) + 
       ylab("") + 
       xlab("")
   } else {
@@ -45,7 +46,8 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
                coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
                theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
                      panel.border = element_blank(), legend.position = "none", 
-                     axis.text = element_blank(),axis.ticks = element_blank()) + 
+                     axis.text = element_blank(),axis.ticks = element_blank(),
+                     plot.margin = unit(c(0,0,0,0),"in")) +
                ylab("") + 
                xlab("")
   }

### 2015-12-21 R/nlaMap.R
Working on autocrop

@@ -30,15 +30,16 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
                          breaks=c("Oligo", "Meso", "Eu","Hyper"),
                          labels=c("Oligotrophc", "Mesotrophic", "Eutrophic","Hypereutrophic"))+
       coord_map("albers", lat2 = 45.5, lat1 = 29.5) +
-      theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
-            panel.border = element_blank(),
-            axis.text = element_blank(),axis.ticks = element_blank(),
-            legend.key.width=unit(2, "line"), 
-            legend.key.height=unit(0.5, "line"),
-            legend.position="bottom", legend.direction="horizontal",
-            plot.margin = unit(c(0,0,0,0),"in")) + 
-      ylab("") + 
-      xlab("")
+      theme(panel.background = element_blank(),
+            panel.grid = elelment_blank(),
+            panel.margin=unit(c(0,0,0,0),"in"),
+            plot.margin=unit(c(0,0,0,0),"in"),
+            plot.background = element_blank(),
+            legend.position = "bottom",
+            legend.direction="horizontal", 
+            axis.text=element_blank(),
+            axis.ticks = element_blank(),
+            axis.title=element_blank())
   } else {
   gmap<-ggplot(states,aes(x=long,y=lat))+
                geom_polygon(aes(group=group),fill=myColor[2],colour=myColor[1])+
@@ -46,8 +47,7 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
                coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
                theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
                      panel.border = element_blank(), legend.position = "none", 
-                     axis.text = element_blank(),axis.ticks = element_blank(),
-                     plot.margin = unit(c(0,0,0,0),"in")) +
+                     axis.text = element_blank(),axis.ticks = element_blank()) + 
                ylab("") + 
                xlab("")
   }

### 2015-12-28 R/nlaMap.R
committing local changes

@@ -31,7 +31,7 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
                          labels=c("Oligotrophc", "Mesotrophic", "Eutrophic","Hypereutrophic"))+
       coord_map("albers", lat2 = 45.5, lat1 = 29.5) +
       theme(panel.background = element_blank(),
-            panel.grid = elelment_blank(),
+            panel.grid = element_blank(),
             panel.margin=unit(c(0,0,0,0),"in"),
             plot.margin=unit(c(0,0,0,0),"in"),
             plot.background = element_blank(),

### 2015-12-28 R/condAccuracy.R
closes #24

@@ -28,7 +28,6 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
   cp2<-data.frame(max_vote=cp2$max_vote,Raw.Data.Probability=cp2$Raw.Data.Probability,
                   model=rep(pred_prob2$model[1],length(cp2[[1]])))
   
-  #browser()
   cp1_k <- vector("numeric",nrow(cp1))
   for(i in 1:nrow(cp1)){
     x<-dplyr::filter(pred_prob1,max>=cp1$max_vote[i])
@@ -54,7 +53,9 @@ condAccuracy<-function(pred_prob1,pred_prob2,xlab="x",...){
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
           axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
           axis.text.x = element_text(family="sans",size = 11),
-          axis.text.y = element_text(family="sans",size = 11)) + 
+          axis.text.y = element_text(family="sans",size = 11),
+          legend.position = c(0, 0.975), 
+          legend.justification = c(0, 1)) + 
     labs(y="Kappa Coefficient",x=xlab) +
     scale_colour_manual(name='',values = viridis(2))
   return(ggp)

### 2015-12-28 R/ecor_map.R

@@ -12,23 +12,28 @@
 #' @export
 #' @import ggplot2
 ecor_map<-function(ecor){
-  #browser()
   #borrowed from: https://github.com/hadley/ggplot2/wiki/plotting-polygon-shapefiles#workaround-forced-layering-of-geom_plot
   ecor@data$id <- rownames(ecor@data)
   ecor_f <- fortify(ecor, region="WSA_9")
   names(ecor_f)[which(names(ecor_f)=="id")] = "WSA_9"
   ecor_df <- plyr::join(ecor_f, ecor@data, by="WSA_9")
   gmap<-ggplot(ecor_df, aes(long,lat,group=group,fill=WSA_9)) +
-    geom_polygon(data=subset(ecor_df,WSA_9!="Western Mountains (WMT)")) +
-    geom_polygon(data=subset(ecor_df,WSA_9=="Western Mountains (WMT)")) +
-    geom_path(color="white") +
+    geom_polygon(data=subset(ecor_df,WSA_9!="Western\nMountains (WMT)")) +
+    geom_polygon(data=subset(ecor_df,WSA_9=="Western\nMountains (WMT)")) +
+    geom_path(color="lightgrey") +
     coord_equal() +
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
     theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(),
-          panel.border = element_blank(), #legend.position = "none",
+          panel.border = element_blank(), 
           axis.text = element_blank(),axis.ticks = element_blank(),
           legend.text = element_text(size=10),
-          legend.key.width = unit(0.5, "line")) +
+          legend.position = c(0.96,0.5),
+          #legend.justification = c(0,0),
+          legend.key.width = unit(0.5, "line"),
+          legend.key.height = unit(1.6, "line"),
+          legend.margin = unit(0,"line"),
+          plot.margin = unit(c(1,3,0.5,0.5),"line")
+          ) +
     ylab("") +
     xlab("") +
     scale_fill_manual(name="",values=viridis(10))

### 2015-12-28 R/nlaMap.R

@@ -35,7 +35,7 @@ nlaMap<-function(states,lakes,cats=F,myColor=1:3,catColor=NULL){
             panel.margin=unit(c(0,0,0,0),"in"),
             plot.margin=unit(c(0,0,0,0),"in"),
             plot.background = element_blank(),
-            legend.position = "bottom",
+            legend.position = c(0.5,0),
             legend.direction="horizontal", 
             axis.text=element_blank(),
             axis.ticks = element_blank(),

### 2015-12-28 R/prob_cdf.R

@@ -17,12 +17,13 @@
 #' @import ggplot2
 
 prob_cdf <- function(probs1, probs2, ...) {
+  
   df<-rbind(probs1,probs2)
   df <- na.omit(data.frame(df))
   options(scipen = 5)  #tell r not to use scientific notation on axis labels
   x <- ggplot(df,aes(x=max,colour=model)) +
     stat_ecdf(size=2) +
-    geom_ribbon(aes(ymin = lower,ymax = upper,fill=model),alpha = 0.4) +
+    geom_ribbon(aes(ymin = lower,ymax = upper,fill=model,show_guide=FALSE),alpha = 0.4) +
     theme(text = element_text(family="sans"),
           panel.background = element_blank(), #panel.grid = element_blank(), 
           panel.border = element_rect(fill = NA), 
@@ -32,7 +33,9 @@ prob_cdf <- function(probs1, probs2, ...) {
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 12),
           axis.title.y = element_text(family="sans",vjust = 1.5, size = 12),
           axis.text.x = element_text(family="sans",size = 11),
-          axis.text.y = element_text(family="sans",size = 11)) + 
+          axis.text.y = element_text(family="sans",size = 11),
+          legend.position = c(0, 0.975), 
+          legend.justification = c(0, 1)) + 
     geom_hline(linetype = 3, size = 1, colour = "gray", 
                yintercept = c(0,0.5,1)) +
     labs(...) + 

### 2015-12-28 R/ts_prob_map.R

@@ -27,12 +27,17 @@ ts_prob_map<-function(states,lakes,probs,save_sep=FALSE){
   gmap_m<-make_ts_map(states,lakes,"Mesotrophic",high=viridis(4)[2])#"#756bb1")
   gmap_e<-make_ts_map(states,lakes,"Eutrophic", high=viridis(4)[3])#"#e6550d")
   gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", high=viridis(4)[4])#"#de2d26")
+  gmap_m$theme$plot.margin <- grid::unit(c(0,0,0,-2),"line")
+  gmap_h$theme$plot.margin <- grid::unit(c(0,0,0,-2),"line")
+  gmap_o$theme$plot.margin <- grid::unit(c(0,-2,0,0),"line")
+  gmap_e$theme$plot.margin <- grid::unit(c(0,-2,0,0),"line")
   if(save_sep){
     ggsave("oligo_prob.jpg",gmap_o,width=8.5)
     ggsave("meso_prob.jpg",gmap_m,width=8.5)
     ggsave("eu_prob.jpg",gmap_e,width=8.5)
     ggsave("hyper_prob.jpg",gmap_h,width=8.5)
   }
+  
   return(multiplot(gmap_o,gmap_e,gmap_m,gmap_h,cols=2))
 }
 
@@ -55,15 +60,16 @@ make_ts_map <- function(states,lake_prob,title,low,high){
     geom_polygon(aes(group=group),alpha=0,colour="grey")+
     coord_map("albers", lat2 = 45.5, lat1 = 29.5)+
     theme(panel.background = element_rect(fill="white"), panel.grid = element_blank(), 
-          panel.border = element_blank(), 
+          panel.border = element_blank(),
           axis.text = element_blank(),axis.ticks = element_blank(),
-          plot.margin = grid::unit(c(0,0,0,0),"inches"),
+          plot.margin = grid::unit(c(0,0,0,0),"line"),
           legend.key.width=unit(2, "line"), 
           legend.key.height=unit(0.5, "line"),
-          legend.position="bottom", legend.direction="horizontal") + 
+          legend.position = c(0.5,-0.1), legend.direction="horizontal") + 
     ggtitle(title)+
     ylab("") + 
     xlab("") +
-    scale_colour_gradient("Probability",low="white",high=high)
+    scale_colour_gradient("Probability",low="white",high=high,
+                          breaks=c(0.2,0.4,0.6,0.8))
   return(gmap)
 }

### 2015-12-28 R/importancePlot.R
still need to get micrograms in plot...

@@ -45,7 +45,12 @@ importancePlot <- function(rf, sumtbl=NULL, data_def = NULL, type = c("acc", "gi
       o <- order(imp_df$Mean_Decrease, decreasing = FALSE)
       imp_df$variables <- factor(imp_df$variables, 
                                  levels = imp_df$variables[o], ordered = T)
+      imp_df$variables[37] <- expression(paste("Total Nitrogen", mu, "g/L)",
+                                                   sep=""))
+      imp_df$variables[38] <- expression(paste("Total Phosphorus", mu, "g/L)",
+                                                   sep=""))
     }
+    browser()
     x <- ggplot(imp_df, aes(Mean_Decrease, variables)) + 
          geom_point(...) +   
          geom_hline(linetype = 3, size = 1, colour = "gray", 

### 2020-10-05 R/class_prob_rf.R
yanked cache, added part to descript, and created some output csv containing predictions.

@@ -37,7 +37,8 @@ class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
   class_prob<-data.frame(t(class_prob))
   class_prob$nla_id<-row.names(class_prob)
   class_prob$max <- apply(class_prob[,1:4],1,max)
-  class_prob$pred_class <- cut(rf_obj$predicted,breaks,labels)
+  class_prob$pred_class_oob <- cut(rf_obj$predicted,breaks,labels)
+  class_prob$pred_class <- cut(predict(rf_obj,newdata=newdata), breaks, labels)
   class_prob$obs_class <- cut(rf_obj$y,breaks,labels)
   row.names(class_prob)<-as.character(1:nrow(class_prob))
   return(class_prob)  

### 2020-10-05 R/importancePlot.R

@@ -18,6 +18,7 @@
 importancePlot <- function(rf, sumtbl=NULL, data_def = NULL, type = c("acc", "gini"), ...) {
     
     idf<-importance(rf)
+    browser()
     # set up data.frame for ggplot
     if (type == "gini") {
         imp_df <- data.frame(variables = names(idf[,dim(idf)[2]]), 

### 2020-10-05 vignettes/manuscript_peerj.R

@@ -1,0 +1,127 @@
+## ----setup, include=FALSE, echo=FALSE-----------------------------------------
+options(repos="https://cran.rstudio.com")
+if(!require(wesanderson)){
+devtools::install_github("jhollist/wesanderson")
+}
+if(!require(edarf)){
+devtools::install_github("zmjones/edarf/pkg")
+}
+if(!require(condprob2)){
+devtools::install_github("jhollist/condprob2")
+}
+if(!require(pander)){
+  install.packages("pander")
+}
+if(!require(tidyr)){
+  install.packages("tidyr")
+}
+if(!require(party)){
+  install.packages("party")
+}
+
+if(!require(broom)){
+  install.packages("broom")
+}
+
+if(!require(interpretR)){
+  install.packages("interpretR")
+}
+
+if(!require(doParallel)){
+  install.packages("doParallel")
+}
+if(!require(viridis)){
+  devtools::install_github("CRAN/viridis")
+}
+
+if(!require(LakeTrophicModelling)){
+  devtools::install_github("USEPA/LakeTrophicModelling",quick=TRUE)
+}
+
+library("viridis")
+library("wesanderson")
+library("LakeTrophicModelling")
+library("knitr")
+library("ggplot2")
+library("sp")
+library("rgdal")
+library("e1071")
+library("dplyr")
+library("tidyr")
+library("condprob2")
+library("party")
+library("broom")
+library("edarf")
+#library("caret")
+library("randomForest")
+library("doParallel")
+opts_chunk$set(dev = 'jpeg', dpi=300, fig.width=7.5, knitr.table.format="html")
+data(LakeTrophicModelling)
+
+#
+#Checks for existing cache (from another project)
+#Else if Repeats running chunk outside of knitr
+if(file.exists("prior_cache")){
+  for(i in gsub(".rdb","",list.files("prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+} else if(file.exists("vignettes/prior_cache")){
+  for(i in gsub(".rdb","",list.files("vignettes/prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+} else if(file.exists("../prior_cache")){
+  for(i in gsub(".rdb","",list.files("../prior_cache",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
+#multiple caches... silly Jeff
+#
[…]

### 2020-10-06 R/class_prob_rf.R
oob probs added

@@ -22,6 +22,14 @@ class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
   if(rf_obj$type!="regression"){
     stop("rf_obj is not a regression randomForest object.")
   }
+  
+  get_oob_predictions <- function(rf_obj, newdata){
+    if(!"inbag" %in% names(rf_obj)){stop("The in bag matrix is not present.  Try re-running random forest with keep.inbag = T.")}
+    rf_pred <- predict(rf_obj, newdata = newdata, predict.all = TRUE)
+    rf_inbag <- rf_obj$inbag
+    rf_inbag[rf_inbag != 0] <- NA
+    rf_pred$individual + rf_inbag
+  } 
   type <- match.arg(type)
   preds <- predict(rf_obj,newdata=newdata,predict.all=TRUE)
   preds_df <- data.frame(preds$individual)
@@ -41,5 +49,20 @@ class_prob_rf <- function(rf_obj,newdata,breaks,labels,ordered=FALSE,
   class_prob$pred_class <- cut(predict(rf_obj,newdata=newdata), breaks, labels)
   class_prob$obs_class <- cut(rf_obj$y,breaks,labels)
   row.names(class_prob)<-as.character(1:nrow(class_prob))
+  oob_pred <- data.frame(get_oob_predictions(rf_obj,newdata))
+  oob_class_prob <- apply(oob_pred,2, function(x) cut(x,breaks,labels))
+  row.names(oob_class_prob)<-row.names(oob_pred)
+  if(type=="probs"){
+    oob_class_prob <- apply(oob_class_prob,1, function(x)
+      table(factor(x,levels=labels, ordered=ordered))/length(na.omit(oob_class_prob[1,])))
+  } else {
+    oob_class_prob <- apply(oob_class_prob,1, function(x)
+      table(factor(x,levels=labels, ordered=ordered)))
+  }
+  oob_class_prob<-data.frame(t(oob_class_prob))
+  oob_class_prob$nla_id<-row.names(oob_class_prob)
+  names(oob_class_prob)[1:4] <- c("oligo_oob", "meso_oob", "eu_oob", 
+                                  "hyper_oob")
+  class_prob <- left_join(class_prob, oob_class_prob)
   return(class_prob)  
 }  

### 2020-10-06 R/ts_prob_map.R
update ts_prob_map to do oob or all data

@@ -21,12 +21,12 @@
 #' ts_prob_map(state,lakes_dd,gis_rf_ts_prob)
 #' @export
 #' @import ggplot2
-ts_prob_map<-function(states,lakes,probs,save_sep=FALSE){
+ts_prob_map<-function(states,lakes,probs,save_sep=FALSE, original = TRUE){
   lakes<-merge(lakes,probs,by="nla_id")
-  gmap_o<-make_ts_map(states,lakes,"Oligotrophic",high=viridis(4)[1])#"#08519c")
-  gmap_m<-make_ts_map(states,lakes,"Mesotrophic",high=viridis(4)[2])#"#756bb1")
-  gmap_e<-make_ts_map(states,lakes,"Eutrophic", high=viridis(4)[3])#"#e6550d")
-  gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", high=viridis(4)[4])#"#de2d26")
+  gmap_o<-make_ts_map(states,lakes,"Oligotrophic",high=viridis(4)[1], original)#"#08519c")
+  gmap_m<-make_ts_map(states,lakes,"Mesotrophic",high=viridis(4)[2], original)#"#756bb1")
+  gmap_e<-make_ts_map(states,lakes,"Eutrophic", high=viridis(4)[3], original)#"#e6550d")
+  gmap_h<-make_ts_map(states,lakes,"Hypereutrophic", high=viridis(4)[4], original)#"#de2d26")
   gmap_m$theme$plot.margin <- grid::unit(c(0,0,0,-2),"line")
   gmap_h$theme$plot.margin <- grid::unit(c(0,0,0,-2),"line")
   gmap_o$theme$plot.margin <- grid::unit(c(0,-2,0,0),"line")
@@ -49,12 +49,20 @@ ts_prob_map<-function(states,lakes,probs,save_sep=FALSE){
 #' @param low lower color
 #' @param high higher color
 #' @export
-make_ts_map <- function(states,lake_prob,title,low,high){
-
-  if(title=="Oligotrophic"){value<<-"oligo"}
-  if(title=="Mesotrophic"){value<<-"meso"}
-  if(title=="Eutrophic"){value<<-"eu"}
-  if(title=="Hypereutrophic"){value<<-"hyper"}
+make_ts_map <- function(states,lake_prob,title,low,high, original = TRUE){
+  browser()
+  if(original){
+    if(title=="Oligotrophic"){value<<-"oligo"}
[…]
